---
title: "Comparison of HCMR Case Rare-Variant Carriers on Adverse Clinical OUtcomes "
author: "Jonathan Chan"
date: "2026-01-23"
output: 
  html_document:
    code_folding: hide   
---

This performs pairwise comparisons of adverse clinical outcomes among HCMR cases with rare-variants with subgrouping by different approaches.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(tictoc)
library(kableExtra)
library(readxl)
library(ggrepel)

rm(list=ls())

theme_set(theme_classic(base_size=14))
source('genetic_subset_analysis/common/hcmr_outcomes_functions.R')
```

# Import

This imports in the HCMR data after preprocessing.

```{r}
primary_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_primary_outcome_df_915.rds') #Uses the 915 sarcomere positive dataset (w/o filtering for common genetics)
secondary_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_secondary_outcome_df_915.rds')

hf_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_hf_outcome_df_915.rds')
af_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_af_outcome_df_915.rds')
va_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_va_outcome_df_915.rds')

#Add in the rare variant class as P_LP or VUS or sarcomere-negative
hcmr_rarevar_class <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_rarevar_class_plp_vusfpvus.tsv') %>%
    select(HCR_IDs=HCR_ID, rarevar_class=class) %>%
    mutate(sarcomere_binned = case_when(
      str_detect(rarevar_class,'_P_LP$') ~ 'P_LP', #So multiple variant individuals are assigned to P_LP if they have one
      str_detect(rarevar_class, 'VUS_FP_VUS') ~ 'VUS',
      TRUE ~ 'Sarcomere-negative'
    ))

covars <- c('sex','bmi')
outcome_vars <- c('hcmdt','event','event_date','event_indicator','event_or_censoring_date','time_to_event_or_censoring_years','age_at_event_or_censoring')
```

```{r}
#Merge into single sarcpos tibble
sarcpos <- map(list(primary_outcomes_df, secondary_outcomes_df, hf_outcomes_df, af_outcomes_df, va_outcomes_df), ~ data_preprocessor(., c(covars, outcome_vars), sarcomere_status=T,hcmr_rarevar_class=hcmr_rarevar_class))

rm(primary_outcomes_df, secondary_outcomes_df, hf_outcomes_df, af_outcomes_df, va_outcomes_df, hcmr_rarevar_class)

#Check for unique values of sarcomere_binned
map(sarcpos, ~ unique(.x$sarcomere_binned))
```

# Comparison Tibble Generation

This builds the comparison tibbles for each of the comparisons in the rare variant subset analysis.
```{r}
pergene_rarevar_classification <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_rarevar_class.tsv')
pergene_perpatho_rarevar_classification <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_rarevar_class_plp_vusfpvus.tsv')

#Build the phenotype tibble, combining the two rare variant classifications into the same tibble
sarcpos <- map(sarcpos, function(x) x %>%
  left_join(select(pergene_rarevar_classification,HCR_ID, class_primary=class), by=c('HCR_IDs'='HCR_ID')) %>%
  left_join(select(pergene_perpatho_rarevar_classification,HCR_ID, class_secondary=class), by=c('HCR_IDs'='HCR_ID'))
)
```

```{r}
#This generates the comparison groups for the analyses by collating all the variant classes in each comparison group

#For primary classification comparisons
#For thick vs. thin (excluding ACTC1 from Thin btw)
thick_class <- unique(sarcpos[[1]]$class_primary)[!str_detect(unique(sarcpos[[1]]$class_primary),'ACTC1|TPM1|TNNI3|TNNT2') & str_detect(unique(sarcpos[[1]]$class_primary),'MYBPC3|MYH7|MYL2|MYL3')]
thin_class <- unique(sarcpos[[1]]$class_primary)[str_detect(unique(sarcpos[[1]]$class_primary),'TPM1|TNNI3|TNNT2') & !str_detect(unique(sarcpos[[1]]$class_primary),'ACTC1|MYBPC3|MYH7|MYL2|MYL3')]

#For mybpc3_pLOF vs. mybpc3_nonlof
mybpc3_plof_class <- unique(sarcpos[[1]]$class_primary)[str_detect(unique(sarcpos[[1]]$class_primary),'MYBPC3_pLOF') & !str_detect(unique(sarcpos[[1]]$class_primary),'MYBPC3_nonLOF')]
mybpc3_nonlof_class <- unique(sarcpos[[1]]$class_primary)[str_detect(unique(sarcpos[[1]]$class_primary),'MYBPC3_nonLOF') & !str_detect(unique(sarcpos[[1]]$class_primary),'MYBPC3_pLOF')]

#For mybpc3 vs. MYH7
mybpc3_class <- unique(sarcpos[[1]]$class_primary)[str_detect(unique(sarcpos[[1]]$class_primary),'MYBPC3') & !str_detect(unique(sarcpos[[1]]$class_primary),'MYH7')]
myh7_class <- unique(sarcpos[[1]]$class_primary)[str_detect(unique(sarcpos[[1]]$class_primary),'MYH7') & !str_detect(unique(sarcpos[[1]]$class_primary),'MYBPC3')]

#For secondary classification comparison
#For P/LP vs. VUS_VUS_FP
p_lp_class <- unique(sarcpos[[1]]$class_secondary)[str_detect(unique(sarcpos[[1]]$class_secondary),'P_LP') & !str_detect(unique(sarcpos[[1]]$class_secondary), 'VUS')] #I use this because I wish to exclude all the individuals who have a combination across P_LP and VUS_VUS_FP
vus_vusfp_class <- unique(sarcpos[[1]]$class_secondary)[str_detect(unique(sarcpos[[1]]$class_secondary),'VUS') & !str_detect(unique(sarcpos[[1]]$class_secondary), 'P_LP')] 

#---------------------------------------------------------------------------------------
#Parameter lists for each comparison
thick_thin_param_list <- list(thick_class, 'Thick',thin_class, 'Thin')
mybpc3_plof_nonlof_param_list <- list(mybpc3_plof_class, 'MYBPC3_pLOF', mybpc3_nonlof_class, 'MYBPC3_nonLOF')
mybpc3_myh7_param_list <- list(mybpc3_class, 'MYBPC3',myh7_class,'MYH7')
p_lp_vus_vusfp_param_list <- list(p_lp_class, 'P/LP', vus_vusfp_class, 'VUS')

rm(list=ls(pattern='class$'))
```

This now generates the actual tibble to run the comparison on, generating one tibble per comparison.

```{r comparison_tb}
#This function generates each tibble for each comparison
comparison_pheno_tb_gen <- function(phenotype_tb,param_list,rarevar_class_tb, var_class='primary', only_singles=T){
  #This defines the comparison groups based on the param_list input e.g. for thick vs. thin, G1 = Thick, G2 = Thin
  group1_classes <- param_list[[1]]
  group1_name <- param_list[[2]]
  group2_classes<- param_list[[3]]
  group2_name<- param_list[[4]]
  
  if(var_class=='primary'){ # When using primary classification based on gene-based
    
    output_tb <- phenotype_tb %>% 
      mutate(group=case_when(class_primary %in% group1_classes ~ group1_name,
                           class_primary %in% group2_classes ~ group2_name,
                           T~NA))
  }
  else if (var_class == 'secondary'){ # When using secondary classification based on pathogenicity
    
    output_tb <- phenotype_tb %>%
      mutate(group=case_when(class_secondary %in% group1_classes ~ group1_name,
                             class_secondary %in% group2_classes ~ group2_name,
                             T~NA))
                              
  }

  output_tb <- mutate(output_tb, group=factor(group, levels=c(group2_name, group1_name))) %>%
    filter(!is.na(group)) #Filter to only individuals in either group
  
  #This filters down to single dosage individuals, 
   if (isTRUE(only_singles)){
     if(var_class=='primary')
    
    single_dosage_individuals <- rarevar_class_tb %>%
      mutate(total_dosage = select(.,MYBPC3_pLOF:FHOD3) %>% rowSums(na.rm=T))%>%
      filter(total_dosage==1)

     else if (var_class == 'secondary'){
      
      single_dosage_individuals <- rarevar_class_tb %>%
        mutate(total_dosage = select(.,MYBPC3_P_LP:FHOD3_VUS_FP_VUS) %>% rowSums(na.rm=T)) %>%
        filter(total_dosage==1)
       
     }
     output_tb <- filter(output_tb, HCR_IDs %in% single_dosage_individuals$HCR_ID)

  } 
  return(output_tb)
}
```

These generate the comparison_tbs for each comparison - either transformed or non-transformed and all for single-dosage individuals.

```{r}
#These are for the primary composite
temp <- map(list(thick_thin_param_list,mybpc3_plof_nonlof_param_list,mybpc3_myh7_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[1]], param_list=.,rarevar_class_tb = pergene_rarevar_classification, var_class='primary'))
names(temp) <- c('thick_thin_tb','mybpc3_plof_nonlof_tb','mybc3_myh7_tb')

temp2 <- map(list(p_lp_vus_vusfp_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[1]], param_list=.,rarevar_class_tb = pergene_perpatho_rarevar_classification, var_class='secondary'))
names(temp2) <- c('p_lp_vus_vusfp_tb')

comparison_tbs_primarycomposite <- c(temp, temp2)
rm(temp, temp2)


#These are for the secondary composite
temp <- map(list(thick_thin_param_list,mybpc3_plof_nonlof_param_list,mybpc3_myh7_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[2]], param_list=.,rarevar_class_tb = pergene_rarevar_classification, var_class='primary'))
names(temp) <- c('thick_thin_tb','mybpc3_plof_nonlof_tb','mybc3_myh7_tb')

temp2<- map(list(p_lp_vus_vusfp_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[2]], param_list=.,rarevar_class_tb = pergene_perpatho_rarevar_classification, var_class='secondary'))
names(temp2) <- c('p_lp_vus_vusfp_tb')

comparison_tbs_secondarycomposite <- c(temp, temp2)
rm(temp, temp2)
```

This repeats for the focused clinical composites.

```{r}
#Repeat for the HF composite
temp <- map(list(thick_thin_param_list,mybpc3_plof_nonlof_param_list,mybpc3_myh7_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[3]], param_list=.,rarevar_class_tb = pergene_rarevar_classification, var_class='primary'))
names(temp) <- c('thick_thin_tb','mybpc3_plof_nonlof_tb','mybc3_myh7_tb')

temp2<- map(list(p_lp_vus_vusfp_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[3]], param_list=.,rarevar_class_tb = pergene_perpatho_rarevar_classification, var_class='secondary'))
names(temp2) <- c('p_lp_vus_vusfp_tb')

comparison_tbs_hfcomposite <- c(temp, temp2)

#Repeat for the AF composite
temp <- map(list(thick_thin_param_list,mybpc3_plof_nonlof_param_list,mybpc3_myh7_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[4]], param_list=.,rarevar_class_tb = pergene_rarevar_classification, var_class='primary'))
names(temp) <- c('thick_thin_tb','mybpc3_plof_nonlof_tb','mybc3_myh7_tb')

temp2<- map(list(p_lp_vus_vusfp_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[4]], param_list=.,rarevar_class_tb = pergene_perpatho_rarevar_classification, var_class='secondary'))
names(temp2) <- c('p_lp_vus_vusfp_tb')

comparison_tbs_afcomposite <- c(temp, temp2)

#Repeat for the VA composite

temp <- map(list(thick_thin_param_list,mybpc3_plof_nonlof_param_list,mybpc3_myh7_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[5]], param_list=.,rarevar_class_tb = pergene_rarevar_classification, var_class='primary'))
names(temp) <- c('thick_thin_tb','mybpc3_plof_nonlof_tb','mybc3_myh7_tb')

temp2<- map(list(p_lp_vus_vusfp_param_list),~comparison_pheno_tb_gen(phenotype_tb = sarcpos[[5]], param_list=.,rarevar_class_tb = pergene_perpatho_rarevar_classification, var_class='secondary'))
names(temp2) <- c('p_lp_vus_vusfp_tb')

comparison_tbs_vacomposite <- c(temp, temp2)

rm(thick_thin_param_list,mybpc3_myh7_param_list,mybpc3_plof_nonlof_param_list, p_lp_vus_vusfp_param_list)
```
```{r}
#This provides a series of str_glue statements which outline the number of each group in each of the comparison_tbs_primarycomposite for use in the flowchart diagram.

walk(comparison_tbs_primarycomposite, ~ {
  n_group1 <- nrow(filter(.x, group == levels(.x$group)[2]))
  n_group2 <- nrow(filter(.x, group == levels(.x$group)[1]))
  print(str_glue("Comparison: {unique(.x$group)} -- N Group 1 ({levels(.x$group)[2]}): {n_group1}, N Group 2 ({levels(.x$group)[1]}): {n_group2}"))
})
```

# Cox Regression

This performs Cox regression to compare the adverse clinical outcomes between each pairwise comparison using the comparison tibbles generated above.

```{r}
primary_composite_cox_results <- map2(comparison_tbs_primarycomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'),~cox_function(.x, 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/primary_composite/',suffix=.y)) %>% bind_rows()

secondary_composite_cox_results <- map2(comparison_tbs_secondarycomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'),~cox_function(.x, 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/secondary_composite/',suffix=.y)) %>% bind_rows()

#Plot the forest plot for sarcomere status with both sarcomere and sarcomere_binned
sarcomere_cox_plot_tb <- bind_rows(
    primary_composite_cox_results %>% mutate(composite='Primary'), 
    secondary_composite_cox_results %>% mutate(composite='Secondary')
) %>%
  filter(!str_detect(term, 'sex|bmi')) %>%
  mutate(term = case_when(
    term == 'groupThick' ~ 'Thick\nvs.\nThin',
    term == 'groupMYBPC3_pLOF' ~ 'MYBPC3_pLOF\nvs.\nMYBPC3_nonLOF',
    term == 'groupMYBPC3' ~ 'MYBPC3\nvs.\nMYH7',
    term == 'groupP/LP' ~ 'Sarcomere-P/LP\nvs.\nSarcomere-VUS',
  )) %>%
    mutate(term = factor(term, levels=c('Thick\nvs.\nThin','MYBPC3\nvs.\nMYH7','MYBPC3_pLOF\nvs.\nMYBPC3_nonLOF','Sarcomere-P/LP\nvs.\nSarcomere-VUS'),ordered=T)) %>%
    mutate(Group = term) %>%
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr')) %>%
    mutate(across(c(hr, hr_low, hr_high, p.value, adj_pvalue), ~ signif(., digits=3)))

#Plot the forest plot
forest_plot_cox(
  sarcomere_cox_plot_tb,
  output_dir = 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/',
  suffix='_rarevar_subset_analysis', group_palette='Dark2'
)

write_rds(sarcomere_cox_plot_tb, 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hcmr_rarevar_subset_cox_age_timescale.rds')
write_tsv(sarcomere_cox_plot_tb, 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hcmr_rarevar_subset_cox_age_timescale.tsv')

rm(sarcomere_cox_plot_tb, primary_composite_cox_results, secondary_composite_cox_results)
```

Also run for some more specific composites e.g HF or AF specific as per Ho et al, 2018 but only for the MYBPC3 vs. MYH7 comparison.

```{r}
hf_composite_cox_results <- map2(comparison_tbs_hfcomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'),~cox_function(.x, 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hf_composite/',suffix=.y)) %>% bind_rows()

af_composite_cox_results <- map2(comparison_tbs_afcomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'),~cox_function(.x, 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/af_composite/',suffix=.y)) %>% bind_rows()

va_composite_cox_results <- map2(comparison_tbs_vacomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'),~cox_function(.x, 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/va_composite/',suffix=.y)) %>% bind_rows()

#Plot these together as a forest plot
hf_af_va_cox_plot_tb <- bind_rows(
    hf_composite_cox_results %>% mutate(composite='HF'), 
    af_composite_cox_results %>% mutate(composite='AF'),
    va_composite_cox_results %>% mutate(composite='VA')
) %>%
  filter(!str_detect(term, 'sex|bmi')) %>%
  mutate(term = case_when(
    term == 'groupThick' ~ 'Thick\nvs.\nThin',
    term == 'groupMYBPC3_pLOF' ~ 'MYBPC3_pLOF\nvs.\nMYBPC3_nonLOF',
    term == 'groupMYBPC3' ~ 'MYBPC3\nvs.\nMYH7',
    term == 'groupP/LP' ~ 'Sarcomere-P/LP\nvs.\nSarcomere-VUS',
  )) %>%
    mutate(term = factor(term, levels=c('Thick\nvs.\nThin','MYBPC3\nvs.\nMYH7','MYBPC3_pLOF\nvs.\nMYBPC3_nonLOF','Sarcomere-P/LP\nvs.\nSarcomere-VUS'),ordered=T)) %>%
    mutate(Group = term) %>%
    filter(term == 'MYBPC3\nvs.\nMYH7') %>% #Filter to just the MYBPC3 vs. MYH7 comparison
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr'))%>%
    mutate(across(c(hr, hr_low, hr_high, p.value, adj_pvalue), ~ signif(., digits=3)))

#Plot the forest plot
forest_plot_cox(
  hf_af_va_cox_plot_tb,
  output_dir = 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/',
  suffix='_hf_af_va_composite', group_palette='Set3'
)

write_rds(hf_af_va_cox_plot_tb, 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hcmr_rarevar_subset_cox_age_timescale_hf_af_va_composite.rds')
write_tsv(hf_af_va_cox_plot_tb, 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hcmr_rarevar_subset_cox_age_timescale_hf_af_va_composite.tsv')
```

# Kaplan Meier Analysis

```{r}
#Plot for PRS categorical groups for both composite outcomes and for age timescale

#For primary composite
walk2(comparison_tbs_primarycomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'), ~km_survival_curv_plotter(.x,'group','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/2_KaplanMeier/1_AgeAsTimescale/primary_composite/', entry_time_varname='age', output_suffix=.y, ylimits=list(c(40,100),c(0.0,0.4),c(0.0,0.4))) 
)

#For secondary composite
walk2(comparison_tbs_secondarycomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'), ~km_survival_curv_plotter(.x,'group','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/2_KaplanMeier/1_AgeAsTimescale/secondary_composite/', entry_time_varname='age', output_suffix=.y, ylimits=list(c(40,100),c(0.0,0.4),c(0.0,0.4))) 
)
```

# Cox Assumptions Check

```{r}
#This runs the assumptions check on the models

primary_composite_cox_models <- map2(comparison_tbs_primarycomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'),~cox_function(.x, 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/primary_composite/',suffix=.y, return_model=T)) 

secondary_composite_cox_models <- map2(comparison_tbs_secondarycomposite, c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'),~cox_function(.x, 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/secondary_composite/',suffix=.y, return_model=T))

#Run the assumptions check
#Primary composite
primary_composite_cox_zph <- map2(primary_composite_cox_models,c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/primary_composite/',.y)) %>% mutate(comparison=.y)) %>% bind_rows() %>% filter(term != 'GLOBAL') %>% mutate(adj_p = p.adjust(p.value, method='fdr'))

#Secondary composite
secondary_composite_cox_zph <- map2(secondary_composite_cox_models,c('thick_thin','mybpc3_plof_nonlof','mybc3_myh7', 'p_lp_vus_vusfp'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/secondary_composite/',.y)) %>% mutate(comparison=.y)) %>% bind_rows() %>% filter(term != 'GLOBAL') %>% mutate(adj_p = p.adjust(p.value, method='fdr'))

```

# Single vs. Multi Variant Carrier Analysis

```{r}
all_merged_overlapvars_tb <- read_tsv('genetic_subset_analysis/RareVar/3_output/individual_class/overlap_variants/hcmr_KT_overlapvariants.tsv') %>%
  filter(HCR_ID %in% sarcpos[[1]]$HCR_IDs) #Filter only to the 915 sarcomere-positive individuals

#Identify the number of homozygotes and the number of multiple heterozygotes
homozygotes <- all_merged_overlapvars_tb %>% 
  filter(dosage1==2 | dosage2==2) #There are 2 homozygotes 
multi_het <- all_merged_overlapvars_tb %>%
  filter(!is.na(dosage2))

print(str_c('There are', nrow(homozygotes), 'homozygotes and',length(unique(multi_het$HCR_ID)),'multiple heterozygotes with ', length(intersect(homozygotes$HCR_ID, multi_het$HCR_ID)),'individual being both so technically',nrow(homozygotes)+length(unique(multi_het$HCR_ID))-length(intersect(homozygotes$HCR_ID, multi_het$HCR_ID)), 'individuals with >1 dose of a rare variant', sep=' '))

#These are all the individuals with >1 dosage of rare variant
multi_dosage_ids <- unique(c(homozygotes$HCR_ID,multi_het$HCR_ID))
```

```{r}
#Generate the phenotype tb for comparison for both primary and secondary composites
multi_vs_single_tb <- map(sarcpos, function(x) x %>%
  mutate(group = ifelse(HCR_IDs %in% multi_dosage_ids, 'Multi','Single')) %>%
  mutate(group=factor(group, levels=c('Single','Multi'))) 
)

primary_composite_cox_results_multi <- cox_function(multi_vs_single_tb[[1]], 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/primary_composite/multi/',suffix='_multi')

secondary_composite_cox_results_multi <- cox_function(multi_vs_single_tb[[2]], 
             predictor_name='group',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             entry_time_varname='age', #Left-truncation at baseline age (delayed entry)
             output_dir='genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/secondary_composite/multi/',suffix='_multi')

#Plot the forest plot for sarcomere status with both sarcomere and sarcomere_binned
sarcomere_cox_plot_tb_multi_vs_single <- bind_rows(
    primary_composite_cox_results_multi %>% mutate(composite='Primary'), 
    secondary_composite_cox_results_multi %>% mutate(composite='Secondary')
) %>%
  filter(!str_detect(term, 'sex|bmi')) %>%
  mutate(term = case_when(
    term == 'groupMulti' ~ 'Multi\nvs.\nSingle',
  )) %>%
    mutate(Group = term) %>%
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr'))%>%
    mutate(across(c(hr, hr_low, hr_high, p.value, adj_pvalue), ~ signif(., digits=3)))

#Plot the forest plot
forest_plot_cox(
  sarcomere_cox_plot_tb_multi_vs_single,
  output_dir = 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/',
  suffix='_rarevar_subset_analysis_multi', group_palette='Accent'
)

write_rds(sarcomere_cox_plot_tb_multi_vs_single, 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hcmr_rarevar_subset_cox_age_timescale_multi.rds')
write_tsv(sarcomere_cox_plot_tb_multi_vs_single, 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hcmr_rarevar_subset_cox_age_timescale_multi.tsv')

rm(sarcomere_cox_plot_tb_multi_vs_single, primary_composite_cox_results_multi, secondary_composite_cox_results_multi)
```

This generates the Kaplan-Meier plots for the single vs. multi-variant carrier analysis.

```{r}
#Plot for PRS categorical groups for both composite outcomes and for age timescale

#For primary composite
km_survival_curv_plotter(multi_vs_single_tb[[1]],'group','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/2_KaplanMeier/1_AgeAsTimescale/primary_composite/', entry_time_varname='age', output_suffix='_multi', ylimits=list(c(40,100),c(0.0,0.4),c(0.0,0.4)))

km_survival_curv_plotter(multi_vs_single_tb[[2]],'group','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/2_KaplanMeier/1_AgeAsTimescale/secondary_composite/', entry_time_varname='age', output_suffix='_multi', ylimits=list(c(40,100),c(0.0,0.4),c(0.0,0.4)))

```

## Summary Counts Output

This outputs for each comparison, the number of individuals per group.


```{r}
summary_counts_tb <- map_dfr(comparison_tbs_primarycomposite, ~.x %>%
  group_by(group) %>%
  summarise(n=n()) %>%
  pivot_wider(names_from=group, values_from=n), .id='comparison')

print(summary_counts_tb)
```

This outputs for each composite and each comparison, the breakdown of the individual event types (underlying each composite) per group.

```{r event_breakdown}
#Clearer labels for the raw event codes (from hcmr_outcomes_preprocessor.Rmd)
event_labels <- c(cardiac_death = 'Cardiac Death',
                  vtach = 'Ventricular Tachyarrhythmia',
                  transplant = 'Heart Transplant',
                  lvad = 'LVAD Implantation',
                  non_cardiac_death = 'Non-Cardiac Death',
                  heart_failure_hospitalisation = 'Heart Failure Requiring Hospitalisation',
                  atrial_fibrillation = 'Atrial Fibrillation Requiring Therapy',
                  non_fatal_stroke = 'Non-Fatal Stroke',
                  hf_death = 'Heart Failure Death',
                  hf_hospitalisation = 'Heart Failure Requiring Hospitalisation',
                  va_death = 'Ventricular Arrhythmic Death',
                  unknown_death = 'Unknown Death')

#This generates one event breakdown table per comparison, with columns Event | Count (Group 1) | Count (Group 2)
event_breakdown_tb_gen <- function(comparison_tb){
  group1_name <- levels(comparison_tb$group)[2]
  group2_name <- levels(comparison_tb$group)[1]

  output_tb <- comparison_tb %>%
    filter(event_indicator == 1) %>%
    count(group, event, name = 'n') %>%
    complete(group, event, fill = list(n = 0)) %>% #Keeps zero counts for event types absent in one group
    pivot_wider(names_from = group, values_from = n) %>%
    select(Event = event, all_of(c(group1_name, group2_name))) %>%
    arrange(Event) %>%
    mutate(Event = recode(Event, !!!event_labels)) #Relabel raw event codes to clearer labels

  output_tb <- bind_rows(output_tb, summarise(output_tb, Event = 'Total', across(-Event, sum))) %>%
    rename_with(~ str_c('Count (', ., ')'), -Event)

  return(output_tb)
}

comparison_labels <- c(thick_thin_tb = 'Thick vs. Thin',
                       mybpc3_plof_nonlof_tb = 'MYBPC3_pLOF vs. MYBPC3_nonLOF',
                       mybc3_myh7_tb = 'MYBPC3 vs. MYH7',
                       p_lp_vus_vusfp_tb = 'Sarcomere-P/LP vs. Sarcomere-VUS')

composite_comparison_tbs <- list(Primary = comparison_tbs_primarycomposite,
                                 Secondary = comparison_tbs_secondarycomposite,
                                 HF = comparison_tbs_hfcomposite,
                                 AF = comparison_tbs_afcomposite,
                                 VA = comparison_tbs_vacomposite)

event_breakdown_tbs <- map(composite_comparison_tbs, ~ map(.x, event_breakdown_tb_gen))

#Build all HTML kables into a single page, with one section per composite and one subtable per comparison
event_breakdown_html <- imap(event_breakdown_tbs, function(composite_tbs, composite_name){
  htmltools::tagList(
    htmltools::h3(str_glue('{composite_name} Composite')),
    imap(composite_tbs, function(tb, comparison_name){
      tb %>%
        kbl(caption = str_glue('{composite_name} composite: {comparison_labels[comparison_name]}'), align = c('l', 'r', 'r')) %>%
        kable_styling(bootstrap_options = c('striped', 'condensed'), full_width = F, position = 'left') %>%
        row_spec(nrow(tb), bold = T) %>%
        as.character() %>%
        htmltools::HTML()
    })
  )
}) %>% htmltools::tagList()

#Save the HTML page (open in browser and copy-paste into Excel) and display it in the Viewer pane
htmltools::save_html(event_breakdown_html, 'genetic_subset_analysis/RareVar/3_output/plots/2_Severity/hcmr_rarevar_subset_event_breakdown.html')
print(htmltools::browsable(event_breakdown_html))

#Also write out each subtable as a TSV (pastes directly into Excel)
iwalk(event_breakdown_tbs, function(composite_tbs, composite_name){
  iwalk(composite_tbs, ~ write_tsv(.x, str_glue('genetic_subset_analysis/RareVar/3_output/plots/2_Severity/event_breakdown_{str_to_lower(composite_name)}_composite_{str_remove(.y, "_tb$")}.tsv')))
})
```