---
title: "Comparison of PRS x Severity in HCMR"
author: "Jonathan Chan"
date: "2026-01-22"
output: 
  html_document:
    code_folding: hide   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(tictoc)
library(kableExtra)
library(readxl)
library(ggrepel)

rm(list=ls())

theme_set(theme_classic(base_size=14))
source('genetic_subset_analysis/common/hcmr_outcomes_functions.R')
```

# Import

This imports in the HCMR data after preprocessing.

```{r}
primary_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_primary_outcome_df_2445.rds')
secondary_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_secondary_outcome_df_2445.rds')
hf_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_hf_outcome_df_2445.rds')
va_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_va_outcome_df_2445.rds')
af_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_af_outcome_df_2445.rds')
stroke_outcomes_df <- readRDS('genetic_subset_analysis/common/output/data/hcmr_stroke_outcome_df_2445.rds')

#Add in the HCMR PRS scores (this is with the minimac-r2-filter for INFO 0.3-1.0) and computed with the intersectvars with UKB
hcmr_prs <- read_tsv('genetic_subset_analysis/PRS/1_input/hcmr_imputedvars_prs.sscore')

hcmr_pcs <- read_delim('popgen/2_gwas/data/hcmr/HCMR_HARDCALLS/pcs_hcmr_10pcs')

#Add in the rare variant class as P_LP or VUS or sarcomere-negative
hcmr_rarevar_class <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_rarevar_class_plp_vusfpvus.tsv') %>%
    select(HCR_IDs=HCR_ID, rarevar_class=class) %>%
    mutate(sarcomere_binned = case_when(
      str_detect(rarevar_class,'_P_LP$') ~ 'P_LP', #So multiple variant individuals are assigned to P_LP if they have one
      str_detect(rarevar_class, 'VUS_FP_VUS') ~ 'VUS',
      TRUE ~ 'Sarcomere-negative'
    ))

# covars <- c('bmi','cmrdiabp') #Use neither covar because I later implement these as clinical risk factors in SevScore
covars <- c('sex','race')
outcome_vars <- c('hcmdt','event','event_date','event_indicator','event_or_censoring_date','time_to_event_or_censoring_years','age_at_event_or_censoring')
```

# Data Preprocessing

This preprocesses + filters the data to all-comer or the subgroups of sarcomere-positive and sarcomere-negative individuals.
```{r}
sarcall <- map(list(primary_outcomes_df, secondary_outcomes_df, hf_outcomes_df, va_outcomes_df, af_outcomes_df, stroke_outcomes_df), ~ data_preprocessor(.,  hcmr_rarevar_class, c(covars, outcome_vars), sarcomere_status=c(T,F), hcmr_pcs=hcmr_pcs, hcmr_prs=hcmr_prs))
sarcpos <- map(list(primary_outcomes_df, secondary_outcomes_df, hf_outcomes_df, va_outcomes_df, af_outcomes_df, stroke_outcomes_df), ~ data_preprocessor(., hcmr_rarevar_class, c(covars, outcome_vars), sarcomere_status=T, hcmr_pcs=hcmr_pcs, hcmr_prs=hcmr_prs))
sarcneg <- map(list(primary_outcomes_df, secondary_outcomes_df, hf_outcomes_df, va_outcomes_df, af_outcomes_df, stroke_outcomes_df), ~ data_preprocessor(.,  hcmr_rarevar_class, c(covars, outcome_vars), sarcomere_status=F, hcmr_pcs=hcmr_pcs, hcmr_prs=hcmr_prs))

rm(primary_outcomes_df, secondary_outcomes_df, hcmr_prs, hcmr_rarevar_class)

#Check for unique values of sarcomere_binned
map(sarcall, ~ unique(.x$sarcomere_binned))
```

# Cox Regression Analysis

## PRS as Predictor

```{r}
#Run the age as timescale analyses 
sarcall_cox_agetimescale <- map2(sarcall[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c(covars, 'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=.y
             )) %>% bind_rows()

#Repeat with time-varying covariate for sarcomere for the secondary composite only
sarcall_cox_agetimescale_secondary_sarctt <- map2(sarcall[2], c('_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c(covars,'sarcomere'),
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=.y, time_dependent_predictor='sarcomere', age0=median(.x$age, na.rm=T)
             )) %>% bind_rows()

#Add back to sarcall_cox_agetimescale
sarcall_cox_agetimescale <- bind_rows(
  sarcall_cox_agetimescale %>% filter(composite == 'primary'),
  sarcall_cox_agetimescale_secondary_sarctt %>% filter(composite == 'secondary')
)

rm(sarcall_cox_agetimescale_secondary_sarctt)
```

```{r}
#This converts the PRS into categorical PRS groups based on quantiles where the output column name is the prs_colname_categorical
prs_categoriser <- function(input_list, prs_colname = 'cv_prs', quantile_thresholds=c(0.2,0.8)){
  output_list <- map(input_list, function(x) x %>%
    mutate(!!sym(paste0(prs_colname,'_categorical')) := case_when(
      !!sym(prs_colname) <= quantile(!!sym(prs_colname), quantile_thresholds[1], na.rm = TRUE) ~ 'Low PRS',
      !!sym(prs_colname) >= quantile(!!sym(prs_colname), quantile_thresholds[2], na.rm = TRUE) ~ 'High PRS',
      TRUE ~ 'Intermediate PRS'
    )) %>%
    mutate(!!sym(paste0(prs_colname,'_categorical')) := factor(!!sym(paste0(prs_colname,'_categorical')), levels = c('Low PRS', 'Intermediate PRS', 'High PRS')))
  )
}
sarcall_categoricalprs <- prs_categoriser(sarcall)

#Rerun the age as timescale analyses but with categorical prs
sarcall_catprs_cox_agetimescale <- map2(sarcall_categoricalprs[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs_categorical',
             predictor_type='categorical',
             covars=c(covars, 'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cat/',
             suffix=.y
             )) %>% bind_rows()

#Repeat with time-varying covariate for sarcomere for the secondary composite only
sarcall_catprs_cox_agetimescale_secondary_sarctt <- map2(sarcall_categoricalprs[2], c('_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs_categorical',
             predictor_type='categorical',
             covars=c(covars,'sarcomere'),
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cat/',
             suffix=.y, time_dependent_predictor='sarcomere', age0=median(.x$age, na.rm=T)
             )) %>% bind_rows()

#Add back to sarcall_catprs_cox_agetimescale
sarcall_catprs_cox_agetimescale <- bind_rows(
  sarcall_catprs_cox_agetimescale %>% filter(composite == 'primary'),
  sarcall_catprs_cox_agetimescale_secondary_sarctt %>% filter(composite == 'secondary')
)

rm(sarcall_catprs_cox_agetimescale_secondary_sarctt)
```

This provides an output summary table of the Cox regression analyses for both the continuous and categorical PRS analyses in the sarcomere-allcomer group.

```{r}
sarcall_cox_summary <- bind_rows(
  sarcall_cox_agetimescale %>% mutate(model_type='Continuous PRS'),
  sarcall_catprs_cox_agetimescale %>% mutate(model_type='Categorical PRS')
) %>%
  filter(str_detect(term, 'cv_prs')) %>%
  select(model_type, term, n_events, composite, hr, hr_low, hr_high, p.value)

write_tsv(sarcall_cox_summary, 'genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/summary_sarcomereallcomer_cox_results.tsv')
```

```{r}
#Knitr kable out the sarcall_cox_summary by primary and secondary composites
sarcall_cox_summary %>%
  filter(composite == 'primary') %>%
  mutate(across(is.numeric, ~signif(.,3))) %>%
  kable(format='html', 
        col.names=c('Model Type','Term','Number of Events','Composite Outcome','Hazard Ratio','HR 95% CI Low','HR 95% CI High','P-value')) %>%
  kable_styling(full_width = F)
```

```{r}
#Knitr kable out the sarcall_cox_summary by primary and secondary composites
sarcall_cox_summary %>%
  filter(composite == 'secondary') %>%
    mutate(across(is.numeric, ~signif(.,3))) %>%
  kable(format='html', 
        col.names=c('Model Type','Term','Number of Events','Composite Outcome','Hazard Ratio','HR 95% CI Low','HR 95% CI High','P-value')) %>%
  kable_styling(full_width = F)
```

This plots out an overall forest plot with both PRS (in its continuous and categorical forms) facetted by the composite outcome.

```{r}
prs_cox_plot_tb <- bind_rows(
  sarcall_cox_agetimescale %>% mutate(model_type='Continuous PRS'),
  sarcall_catprs_cox_agetimescale %>% mutate(model_type='Categorical PRS')
) %>%
  filter(str_detect(term, 'cv_prs')) %>%
  mutate(term = case_when(
    term == 'cv_prs' ~ 'PRS (per SD increase)',
    term == 'cv_prs_categoricalHigh PRS' ~ 'High PRS vs Low PRS',
    term == 'cv_prs_categoricalIntermediate PRS' ~ 'Intermediate PRS vs Low PRS'
  )) %>%
    mutate(composite = str_to_title(composite)) %>%
    mutate(term = factor(term, levels=c('PRS (per SD increase)',
                                    'Intermediate PRS vs Low PRS',
                                    'High PRS vs Low PRS'), ordered=T)) %>%
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group = model_type) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Plot the forest plot
forest_plot_cox(
  prs_cox_plot_tb,
  output_dir = 'genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/',
  suffix='_prs', group_palette='Set1'
)

rm(prs_cox_plot_tb, sarcall_catprs_cox_agetimescale)
```

### Subgroup Analyses

This analyses the continuous PRS effect in the sarcomere-positive and sarcomere-negative subgroups.

```{r}
sarcpos_cox_agetimescale <- map2(sarcpos[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcpos/cont/',
             suffix=.y
             )) %>% bind_rows()

#Repeat for the sarcomere-negative subgroup
sarcneg_cox_agetimescale <- map2(sarcneg[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcneg/cont/',
             suffix=.y
             )) %>% bind_rows()

#Also repeat the categorical PRS analyses for both subgroups
sarcpos_catprs_cox_agetimescale <- map2(sarcall_categoricalprs[1:2] %>% map(~ filter(.x, sarcomere==T)), c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs_categorical',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcpos/cat/',
             suffix=.y
             )) %>% bind_rows()

sarcneg_catprs_cox_agetimescale <- map2(sarcall_categoricalprs[1:2] %>% map(~ filter(.x, sarcomere==F)), c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs_categorical',
             predictor_type='categorical',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcneg/cat/',
             suffix=.y
             )) %>% bind_rows()
```

```{r}
# Split into 2 forest plots: continuous PRS and categorical PRS

# 1) Continuous PRS subgroup forest plot
subgroup_cont_cox_plot_tb <- bind_rows(
    sarcpos_cox_agetimescale |> mutate(Group = "Sarcomere-positive"),
    sarcneg_cox_agetimescale |> mutate(Group = "Sarcomere-negative")
) |>
    filter(term == "cv_prs") |>
    mutate(
        term = "PRS (per SD increase)",
        term = str_c(term, "\n", Group),
        composite = str_to_title(composite)
    ) |>
    select(term, n_events, composite, hr, hr_low, hr_high, p.value, Group) |>
    mutate(adj_pvalue = p.adjust(p.value, method = "fdr"))

forest_plot_cox(
    subgroup_cont_cox_plot_tb,
    output_dir = "genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/",
    suffix = "_prs_subgroups_continuous"
)

# 2) Categorical PRS subgroup forest plot
subgroup_cat_cox_plot_tb <- bind_rows(
    sarcpos_catprs_cox_agetimescale |> mutate(Group = "Sarcomere-positive"),
    sarcneg_catprs_cox_agetimescale |> mutate(Group = "Sarcomere-negative")
) |>
    filter(str_detect(term, "cv_prs_categorical")) |>
    mutate(
        term = case_when(
            term == "cv_prs_categoricalIntermediate PRS" ~ "Intermediate PRS vs Low PRS",
            term == "cv_prs_categoricalHigh PRS" ~ "High PRS vs Low PRS",
            TRUE ~ term
        ),
        term = str_c(term, "\n", Group),
        composite = str_to_title(composite)
    ) |>
    select(term, n_events, composite, hr, hr_low, hr_high, p.value, Group) |>
    mutate(adj_pvalue = p.adjust(p.value, method = "fdr"))

forest_plot_cox(
    subgroup_cat_cox_plot_tb,
    output_dir = "genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/",
    suffix = "_prs_subgroups_categorical"
)

rm(sarcpos_catprs_cox_agetimescale, sarcneg_catprs_cox_agetimescale,
   sarcpos_cox_agetimescale, sarcneg_cox_agetimescale,
   subgroup_cat_cox_plot_tb, subgroup_cont_cox_plot_tb)
```

## Sarcomeric Status Focus

This section focuses on the sarcomeric status as predictor with continuous PRS as covariate.

This provides an output summary table for the association of sarcomere status.

```{r}
sarcomere_status_cox_summary <- sarcall_cox_agetimescale %>%
  filter(str_detect(term, 'sarcomere')) %>%
  select(term, n_events, composite, hr, hr_low, hr_high, p.value)

sarcomere_status_cox_summary %>%
    mutate(across(is.numeric, ~signif(.,3))) %>%
  kable(format='html', 
        col.names=c('Term','Number of Events','Composite Outcome','Hazard Ratio','HR 95% CI Low','HR 95% CI High','P-value')) %>%
  kable_styling(full_width = F)
```

This reruns the Cox regression but with a sarcomere_binned variable into P_LP vs. VUS vs. Sarcomere-negative.

```{r}
#Rerun the age as timescale analyses with sarcomere_binned but with PRS as covariate
sarcall_sarcomerebinned_cox_agetimescale <- map2(sarcall[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='sarcomere_binned',
             predictor_type='categorical',
             covars=c(covars, 'cv_prs'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',suffix=.y
             )) %>% bind_rows()

#Fails to adjust it as a time-varying covariate due to only applicability to 2-level categorical variables
```

```{r}
#Plot the forest plot for sarcomere status with both sarcomere and sarcomere_binned
sarcomere_cox_plot_tb <- bind_rows(
    sarcall_cox_agetimescale %>% mutate(Group = 'Sarcomere Status'), 
    sarcall_sarcomerebinned_cox_agetimescale %>% mutate(Group = 'Sarcomere Variant Class')
) %>%
  filter(str_detect(term, 'sarcomere')) %>%
  filter(term != 'tt(sarcomere)') %>% #Remove the time-varying covariate term for sarcomere as it is not plottable in a forest plot and the main sarcomere term is still interpretable as the overall effect across time
  mutate(term = case_when(
    term %in% c('sarcomereTRUE', 'sarcomere') ~ 'Sarcomere-positive\nvs.\nSarcomere-negative', #This is because sarcomereTRUE is the term name when using normal Cox regression and sarcomere is the term-name for the time-varying covariate Cox regression
    term == 'sarcomere_binnedVUS' ~ 'Sarcomere-VUS\nvs.\nSarcomere-negative',
    term == 'sarcomere_binnedP_LP' ~ 'Sarcomere-P/LP\nvs.\nSarcomere-negative'
  )) %>%
    mutate(term = factor(term, levels=c('Sarcomere-positive\nvs.\nSarcomere-negative',
                                    'Sarcomere-P/LP\nvs.\nSarcomere-negative',
                                    'Sarcomere-VUS\nvs.\nSarcomere-negative'),ordered=T)) %>%
    mutate(composite = str_to_title(composite)) %>%
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Plot the forest plot
forest_plot_cox(
  sarcomere_cox_plot_tb,
  output_dir = 'genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall',
  suffix='_sarcomere', group_palette='Set3'
)

#N.B These are what the * means
#         .data$.p_for_stars < 0.001 ~ "***",
#         .data$.p_for_stars < 0.01 ~ "**",
#         .data$.p_for_stars < 0.05 ~ "*"

rm(sarcomere_cox_plot_tb)
```

# Kaplan-Meier Analyses

This plots Kaplan-Meier curves for select categorical comparisons.

This is for PRS categorical groups.

```{r}
#Plot for PRS categorical groups for both composite outcomes and for both age and time-since-baseline timescales
walk2(sarcall_categoricalprs[1:2], c('_primarycomposite','_secondarycomposite'), ~km_survival_curv_plotter(.x,'cv_prs_categorical','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/PRS/3_output/2_Severity/2_KaplanMeier/1_AgeAsTimescale/sarcall/', output_suffix=.y, ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4))) 
)

walk2(sarcall_categoricalprs[1:2], c('_primarycomposite','_secondarycomposite'), ~km_survival_curv_plotter(.x,'cv_prs_categorical','event_indicator', 'time_to_event_or_censoring_years', 'genetic_subset_analysis/PRS/3_output/2_Severity/2_KaplanMeier/2_TimeSinceBaseline/sarcall/', output_suffix=.y, ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4)),xlab='Time Since Baseline (Years)') 
)

```

This is for binary sarcomere status.

```{r}
#Plot for sarcomere status for both composite outcomes and for both age and time-since-baseline timescales
walk2(sarcall[1:2], c('_primarycomposite','_secondarycomposite'), ~km_survival_curv_plotter(.x,'sarcomere','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/PRS/3_output/2_Severity/2_KaplanMeier/1_AgeAsTimescale/sarcall/', output_suffix=.y, ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4))) 
)

#And repeat for the Kaplan Meier plot using Wilcoxon-Breslow test due to sarcomere status invalidating the proportional hazards assumption for secondary composite

km_survival_curv_plotter(sarcall[[2]],'sarcomere','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/PRS/3_output/2_Severity/2_KaplanMeier/1_AgeAsTimescale/sarcall/', output_suffix='_secondarycomposite_brewslow', ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4)), pairwise_test='breslow') 

walk2(sarcall[1:2], c('_primarycomposite','_secondarycomposite'), ~km_survival_curv_plotter(.x,'sarcomere','event_indicator', 'time_to_event_or_censoring_years', 'genetic_subset_analysis/PRS/3_output/2_Severity/2_KaplanMeier/2_TimeSinceBaseline/sarcall/', output_suffix=.y, ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4)),xlab='Time Since Baseline (Years)') 
)

#Also visualise the distribution of age at recruitment for sarcomere positive and negative with a Wilcoxon rank sum test

age_sarcomere_dist <- ggplot(sarcall[[1]])+
    geom_density(aes(x=age, fill=sarcomere), alpha=0.5)+
    scale_fill_viridis(discrete=TRUE, option='D', name='Sarcomere Status', labels=c('Negative','Positive'))+
    labs(x='Age at Recruitment (Years)', y='Density')

print(age_sarcomere_dist)

```

This is for sarcomere_binned status.

```{r}
#Plot for sarcomere_binned status for both composite outcomes and for both age and time-since-baseline timescales

walk2(sarcall[1:2], c('_primarycomposite','_secondarycomposite'), ~km_survival_curv_plotter(.x,'sarcomere_binned','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/PRS/3_output/2_Severity/2_KaplanMeier/1_AgeAsTimescale/sarcall/', output_suffix=.y, ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4))) 
)

walk2(sarcall[1:2], c('_primarycomposite','_secondarycomposite'), ~km_survival_curv_plotter(.x,'sarcomere_binned','event_indicator', 'time_to_event_or_censoring_years', 'genetic_subset_analysis/PRS/3_output/2_Severity/2_KaplanMeier/2_TimeSinceBaseline/sarcall/', output_suffix=.y, ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4)),xlab='Time Since Baseline (Years)') 
)

#Also visualise the distribution of age at recruitment for sarcomere positive and negative with a Wilcoxon rank sum test
age_sarcomere_dist <- ggplot(sarcall[[1]])+
    geom_density(aes(x=age, fill=sarcomere_binned), alpha=0.5)+
    scale_fill_viridis(discrete=TRUE, option='D', name='Sarcomere Variant Class')+
    labs(x='Age at Recruitment (Years)', y='Density')

print(age_sarcomere_dist)
```

# Time-Since-Baseline Analyses

This uses time-since-baseline as the timescale.

```{r}
#Run the time-since-baseline analyses 
sarcall_cox_timesincebaseline <- map2(sarcall[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c('age',covars,'sarcomere'),
             indicator_varname ='event_indicator',
             time_varname='time_to_event_or_censoring_years',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/2_TimeSinceBaseline/sarcall/cont/',
             suffix=.y
             )) %>% bind_rows()

#Rerun the time-since-baseline analyses
sarcall_catprs_cox_timesincebaseline <- map2(sarcall_categoricalprs[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs_categorical',
             predictor_type='categorical',
             covars=c('age',covars,'sarcomere'),
             indicator_varname ='event_indicator',
             time_varname='time_to_event_or_censoring_years',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/2_TimeSinceBaseline/sarcall/cat/',
             suffix=.y
             )) %>% bind_rows()

#Rerun the age as timescale analyses with sarcomere_binned but with PRS as covariate
sarcall_sarcomerebinned_cox_timesincebaseline <- map2(sarcall[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='sarcomere_binned',
             predictor_type='categorical',
             covars=c('age','sex','race'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='time_to_event_or_censoring_years',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/2_TimeSinceBaseline/sarcall/cont/',suffix=.y
             )) %>% bind_rows()

print(sarcall_cox_timesincebaseline%>% select(term, hr, p.value, composite))
print(sarcall_catprs_cox_timesincebaseline%>% select(term, hr, p.value, composite))
print(sarcall_sarcomerebinned_cox_timesincebaseline%>% select(term, hr, p.value, composite))

rm(list=ls(pattern='timesincebaseline'))
```

# Time/Age Since Diagnosis Analyses

This section reruns the Cox regression analyses using time since diagnosis as timescale instead of age or time-since-baseline.

```{r}
sarcall_time_since_diagnosis <- map(sarcall, function(x) x %>%
    filter(!is.na(hcmdt)) %>% #Remove NA values for hcmdt
    mutate(hcmdt=ifelse(hcmdt=='15-JUN-1692','15-Jun-62',hcmdt)) %>% #Fix the one erroneous date
    mutate(hcmdt = case_when(
        str_detect(hcmdt, '^\\d{1,2}-[A-Za-z]{3}-\\d{2}$') ~ dmy(hcmdt), #To parse dates like 1-Feb-15
        str_detect(hcmdt, '^\\d{4}$') ~ ymd(paste0(hcmdt,'-01-01')), #To parse year-only dates like 2015
        TRUE ~ as.Date(NA)
        )) %>% #Diagnosis date
    mutate(time_since_diagnosis_years = as.numeric(difftime(event_or_censoring_date, hcmdt, units='days'))/365.25
    ) %>%
    mutate(age_at_diagnosis = age - time_since_diagnosis_years
    ))

#Run the time-since-diagnosis analyses
sarcall_cox_timesincediagnosis <- map2(sarcall_time_since_diagnosis[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c('age',covars,'sarcomere'),
             indicator_varname ='event_indicator',
             time_varname='time_since_diagnosis_years',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/3_TimeSinceDiagnosis/sarcall/cont/',
             suffix=.y
             )) %>% bind_rows()

print(sarcall_cox_timesincediagnosis %>% select(term, hr, p.value, composite))
rm(sarcall_cox_timesincediagnosis)
```

This also fits a joint model with age as timescale but including age at diagnosis as a covariate.

```{r}
#Run the analyses with age as timescale but with age at diagnosis as covariate
sarcall_cox_cov_ageatdiagnosis <- map2(sarcall_time_since_diagnosis[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c(covars,'sarcomere','age_at_diagnosis'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cov_ageatdiag/',
             suffix=.y
             )) %>% bind_rows()

sarcall_cox_cov_ageatdiagnosis_sarcbinned <- map2(sarcall_time_since_diagnosis[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c(covars,'sarcomere_binned','age_at_diagnosis'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cov_ageatdiag/sarcbinned/',
             suffix=.y
             )) %>% bind_rows()
```

There are PH assumption errors with age_at_diagnosis so I fit it as a time-varying coefficient and sarcomere for the secondary composite as well. I then plot all of these together.

```{r}
rm(sarcall_cox_cov_ageatdiagnosis)

sarcall_cox_cov_ageatdiagnosis_primary <- cox_function(sarcall_time_since_diagnosis[[1]],
             predictor_name='sarcomere',
             predictor_type='categorical',
             covars=c(covars,'cv_prs','age_at_diagnosis'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/primary/',
             suffix='_primary_covageatdiag',time_dependent_predictor='age_at_diagnosis',age0 = median(sarcall_time_since_diagnosis[[1]]$age, na.rm=T)
             )

sarcall_cox_cov_ageatdiagnosis_secondary <- cox_function(sarcall_time_since_diagnosis[[2]],
             predictor_name='sarcomere',
             predictor_type='categorical',
             covars=c(covars,'cv_prs','age_at_diagnosis'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/secondary/sarctt/',
             suffix='_secondary_covageatdiagsarctt',time_dependent_predictor = c('age_at_diagnosis','sarcomere'), age0 = median(sarcall_time_since_diagnosis[[2]]$age, na.rm=T) #i.e. 52
             )

sarcall_cox_cov_ageatdiagnosis_sarcbinned <- map2(sarcall_time_since_diagnosis[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c(covars,'sarcomere_binned','age_at_diagnosis'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cov_ageatdiag/sarcbinned/',time_dependent_predictor='age_at_diagnosis',age0 = median(sarcall_time_since_diagnosis[[2]]$age, na.rm=T),
             suffix=.y
             )) %>% bind_rows()

#Replot the sarcomere comparison forest plot with this
sarcomere_cox_plot_tb <- bind_rows(
    sarcall_cox_cov_ageatdiagnosis_primary %>% mutate(Group= 'Sarcomere Status') %>% mutate(composite='primary'), 
    sarcall_cox_cov_ageatdiagnosis_secondary %>% mutate(Group = 'Sarcomere Status') %>% mutate(composite='secondary'), 
    sarcall_cox_cov_ageatdiagnosis_sarcbinned %>% mutate(Group = 'Sarcomere Variant Class')
) %>%
  filter(str_detect(term, 'sarcomere')) %>%
  mutate(term = case_when(
    term %in% c('sarcomereTRUE', 'sarcomere') ~ 'Sarcomere-positive\nvs.\nSarcomere-negative',
    term == 'tt(sarcomere)' ~ 'Time-varying Effect of\nSarcomere-positive vs. Sarcomere-negative',
    term == 'sarcomere_binnedVUS' ~ 'Sarcomere-VUS\nvs.\nSarcomere-negative',
    term == 'sarcomere_binnedP_LP' ~ 'Sarcomere-P/LP\nvs.\nSarcomere-negative'
  )) %>%
    mutate(term = factor(term, levels=c('Sarcomere-positive\nvs.\nSarcomere-negative',
                                    'Sarcomere-P/LP\nvs.\nSarcomere-negative',
                                    'Sarcomere-VUS\nvs.\nSarcomere-negative'),ordered=T)) %>%
    mutate(composite = str_to_title(composite)) %>%
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group) %>%
    filter(!is.na(term)) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Plot the forest plot
forest_plot_cox(
  sarcomere_cox_plot_tb,
  output_dir = 'genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall',
  suffix='_sarcomere_covageatdiag', group_palette='Set3'
)

rm(sarcomere_cox_plot_tb,sarcall_cox_cov_ageatdiagnosis_primary, sarcall_cox_cov_ageatdiagnosis_secondary, sarcall_cox_cov_ageatdiagnosis_sarcbinned)
```

# Cox Assumptions

This evaluates basic assumptions of Cox regression for the mainline analyses e.g.

1. Scaled Schoeneld residuals for proportional hazards assumption
2. Martingale residuals for linearity of continuous predictors
3. Influence plots for outliers and influential points

```{r}
sarcall_cox_agetimescale_models <- map2(sarcall[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c('sex','sarcomere', 'PC1','PC2','PC3'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=.y, return_model=T
             ))

sarcall_catprs_cox_agetimescale_models <- map2(sarcall_categoricalprs[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs_categorical',
             predictor_type='categorical',
             covars=c('sex','sarcomere', 'PC1','PC2','PC3'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cat/',
             suffix=.y, return_model=T
             )) 

sarcall_cox_cov_ageatdiagnosis_models <- map2(sarcall_time_since_diagnosis[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=c('sex','sarcomere','cv_prs', 'PC1','PC2','PC3', 'age_at_diagnosis'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',suffix=.y, return_model=T
             )) 

# Sarcomere-positive and sarcomere-negative subgroup models
sarcpos_cox_agetimescale_models <- map2(sarcpos[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcpos/cont/',
             suffix=.y, return_model=T
             )) 

sarcneg_cox_agetimescale_models <- map2(sarcneg[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs',
             predictor_type='continuous',
             covars=covars, #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcneg/cont/',
             suffix=.y, return_model=T
             )) 

sarcpos_catprs_cox_agetimescale_models <- map2(sarcall_categoricalprs[1:2] %>% map(~ filter(.x, sarcomere==T)), c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs_categorical',
             predictor_type='categorical',
             covars=c('sex','race'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcpos/cat/',
             suffix=.y, return_model=T
             ))

sarcneg_catprs_cox_agetimescale_models <- map2(sarcall_categoricalprs[1:2] %>% map(~ filter(.x, sarcomere==F)), c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='cv_prs_categorical',
             predictor_type='categorical',
             covars=c('sex','race'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcneg/cat/',
             suffix=.y, return_model=T
             )) 

sarcall_sarcomerebinned_cox_agetimescale_models <- map2(sarcall[1:2], c('_primary','_secondary'),
            ~cox_function(.x, 
             predictor_name='sarcomere_binned',
             predictor_type='categorical',
             covars=c('sex','cv_prs', 'PC1','PC2','PC3'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',suffix=.y, return_model=T
             )) 
```


```{r}
#Run the cox_diagnostics function on all models and then rm the models

#Sarcomere-allcomer: continous PRS and categorical PRS (both with sarcomere)
sarcall_cox_zph <- map2(sarcall_cox_agetimescale_models,c('primary','secondary'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',.y)) %>% mutate(composite=.y)) %>% bind_rows() %>% filter(term!='GLOBAL') %>% mutate(adj_pvalue = p.adjust(p.value, method='fdr'))
sarcall_catprs_cox_zph <- map2(sarcall_catprs_cox_agetimescale_models,c('primary','secondary'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cat/',.y)) %>% mutate(composite=.y)) %>% bind_rows() %>% filter(term!='GLOBAL')%>% mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Sarcomere all-comer: sarcomere binned (with continuous PRS)
sarcall_sarcomerebinned_cox_zph <- map2(sarcall_sarcomerebinned_cox_agetimescale_models,c('primary','secondary'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/sarcbinned/',.y)) %>% mutate(composite=.y)) %>% bind_rows() %>% filter(term!='GLOBAL')%>% mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Sarcomere all-comer: age as timescale but with age at diagnosis as covariate
sarcall_cov_ageatdiag_cox_zph <- map2(sarcall_cox_cov_ageatdiagnosis_models,c('primary','secondary'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cov_ageatdiag/',.y)) %>% mutate(composite=.y)) %>% bind_rows() %>% filter(term!='GLOBAL')%>% mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Sarcomere-positive subgroup
sarcpos_cox_zph <- map2(sarcpos_cox_agetimescale_models,c('primary','secondary'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcpos/cont/',.y)) %>% mutate(composite=.y)) %>% bind_rows()%>% filter(term!='GLOBAL') %>% mutate(adj_pvalue = p.adjust(p.value, method='fdr'))
sarcpos_catprs_cox_zph <- map2(sarcpos_catprs_cox_agetimescale_models,c('primary','secondary'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcpos/cat/',.y)) %>% mutate(composite=.y)) %>% bind_rows() %>% filter(term!='GLOBAL') %>% mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Sarcomere-negative subgroup
sarcneg_cox_zph <- map2(sarcneg_cox_agetimescale_models,c('primary','secondary'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcneg/cont/',.y)) %>% mutate(composite=.y)) %>% bind_rows() %>% filter(term!='GLOBAL')%>% mutate(adj_pvalue = p.adjust(p.value, method='fdr'))
sarcneg_catprs_cox_zph <- map2(sarcneg_catprs_cox_agetimescale_models,c('primary','secondary'), ~ cox_diagnostics(.x, str_c('genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcneg/cat/',.y)) %>% mutate(composite=.y)) %>% bind_rows()%>% filter(term!='GLOBAL') %>% mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

rm(sarcall_cox_agetimescale_models,
   sarcall_catprs_cox_agetimescale_models,
   sarcpos_cox_agetimescale_models,
   sarcneg_cox_agetimescale_models,
   sarcpos_catprs_cox_agetimescale_models,
   sarcneg_catprs_cox_agetimescale_models,
   sarcall_sarcomerebinned_cox_agetimescale_models) 

rm(list=ls(pattern='zph$'))
```

# Alternate PRSs

This looks for association between alternative PRSs (not the HCM risk/susceptibility PRS) with adverse clinical outcomes.
This is based on Yang et al, 2025 which showed that risk PRS did not associate with disease-specific mortality but a PRS that was trained in the general population to model the common genetic risk of the outcomes directly performed better.

```{r}
alt_prs <- map(list.files('genetic_subset_analysis/PRS/3_output/alt_prs/3_hcmr_scores/', pattern='sscore', full.names=T), ~read_tsv(.) %>% select(HCR_IDs = IID, SCORE1_AVG) %>% dplyr::rename(cv_prs = SCORE1_AVG))
names(alt_prs) <- str_match(basename(list.files('genetic_subset_analysis/PRS/3_output/alt_prs/3_hcmr_scores/', pattern='sscore', full.names=T)), '([^_]+)\\.sscore$')[,2]

#Define the PRS of interest
altprs_of_interest <- str_c('cv_prs_',names(alt_prs))

#Rename the columns to be cv_prs_<alt_prs_name>
alt_prs <- map2(
  alt_prs,
  names(alt_prs),
  ~ .x |> rename(!!str_c("cv_prs_", .y) := cv_prs)
)

#Join each of the alt_prs to each of the sarcall datasets
merged_alt_prs <- reduce(alt_prs, function(acc, prs_df) {
    left_join(acc, prs_df, by = "HCR_IDs")
  })
rm(alt_prs)

#Join the merged_alt_prs to each of the sarcall datasets
sarcall_altprs <- map(sarcall, ~ left_join(.x, merged_alt_prs, by='HCR_IDs'))

#Scale each of the alt_prs columns within the sarcall datasets
sarcall_altprs <- map(sarcall_altprs, ~ .x |> mutate(across(starts_with("cv_prs_"), ~scale(.)[,1])))
names(sarcall_altprs) <- c('primary','secondary','hf','va','af','stroke')

rm(merged_alt_prs)
```

This runs the Cox regression over each of the alternate PRSs, for the primary and secondary composites.

```{r}
sarcall_cox_agetimescale_altprs_primarycomp <- map(c('cv_prs',altprs_of_interest),
            ~cox_function(sarcall_altprs[[1]], 
             predictor_name=.,
             predictor_type='continuous',
             covars=c(covars,'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=str_c('_primary_',.)
             )) %>% bind_rows()

sarcall_cox_agetimescale_altprs_secondarycomp <- map(c('cv_prs',altprs_of_interest),
            ~cox_function(sarcall_altprs[[2]], 
             predictor_name=.,
             predictor_type='continuous',
             covars=c(covars,'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=str_c('_secondary_',.)
             )) %>% bind_rows()

#Sarcomere status is time-dependent for secondary composite so refit that model with time-varying coefficient for sarcomere status
sarcall_cox_agetimescale_altprs_secondarycomp <- map(c('cv_prs',altprs_of_interest[altprs_of_interest != 'cv_prs_af']),
            ~cox_function(sarcall_altprs[[2]], 
             predictor_name=.,
             predictor_type='continuous',
             covars=c(covars,'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=str_c('_secondary_',.), time_dependent_predictor = 'sarcomere', age0 = median(sarcall_altprs[[2]]$age, na.rm=T)
             )) %>% bind_rows()

#Both sarcomere status and cv_prs_af are time-dependent for secondary composite so refit that model with time-varying coefficient for both sarcomere status and cv_prs_af
sarcall_cox_agetimescale_altprs_secondarycomp_af <- cox_function(sarcall_altprs[[2]], 
             predictor_name='cv_prs_af',
             predictor_type='continuous',
             covars=c(covars,'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix='_secondary_cv_prs_af', time_dependent_predictor = c('sarcomere','cv_prs_af'), age0 = median(sarcall_altprs[[2]]$age, na.rm=T)
             )

#Merge the three tibble results together

sarcall_cox_agetimescale_altprs <- bind_rows(
    sarcall_cox_agetimescale_altprs_primarycomp %>% mutate(composite='primary'),
    sarcall_cox_agetimescale_altprs_secondarycomp %>% mutate(composite='secondary'),
    sarcall_cox_agetimescale_altprs_secondarycomp_af %>% mutate(composite='secondary')
) %>%
  filter(str_detect(term, 'cv_prs'))

#Plot the normal PRS as a forest plot with the alternate PRSs as well.
sarcomere_cox_plot_tb <- sarcall_cox_agetimescale_altprs %>%
  filter(!str_detect(term, 'tt')) %>% #Remove the time-varying coefficient terms for this visualisation
  mutate(term = case_when(
    term == 'cv_prs' ~ 'HCM',
    str_detect(term, 'cv_prs_') ~ str_match(term, 'cv_prs_([A-Za-z\\d]+)')[,2],
    T ~NA)) %>%
    mutate(composite = str_to_title(composite)) %>%
    mutate(Group = ifelse(term=='HCM','Continuous PRS (HCM)', 'Continuous PRS (Alt.)')) %>%
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group) %>%
    filter(!is.na(term)) %>%
    mutate(term = case_when(
      term == 'af' ~ 'AF',
      term == 'dbp' ~ 'DBP',
      term == 'hf' ~ 'HF',
      term == 'stroke' ~ 'Stroke',
      term == 'arrest' ~ 'Cardiac Arrest (UKB)',
      term == 'va' ~ 'Ventricular\nArrhythmia (UKB/BJJ)',
      T ~ term
    )) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Plot the forest plot
forest_plot_cox(
  sarcomere_cox_plot_tb,
  output_dir = 'genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall',
  suffix='_altprs', group_palette='Set2', ylabel='PRS'
)

rm(sarcall_cox_agetimescale_altprs_primarycomp, sarcall_cox_agetimescale_altprs_secondarycomp, sarcall_cox_agetimescale_altprs_secondarycomp_af, sarcomere_cox_plot_tb)
```

This evaluates each of the specific PRSs with its specific composite e.g. HF with HF-related outcomes and AF + VA PRS with VA-related outcomes and stroke.

```{r}
#HF PRS x HF composite
sarcall_cox_agetimescale_altprs_hfcomp <- cox_function(sarcall_altprs[['hf']], 
             predictor_name='cv_prs_hf',
             predictor_type='continuous',
             covars=c(covars,'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix='hf'
             )

#AF PRS x AF composite (with tt(sarcomere) because of PH assumption violation)
sarcall_cox_agetimescale_altprs_af <- cox_function(sarcall_altprs[['af']], 
             predictor_name='cv_prs_af',
             predictor_type='continuous',
             covars=c(covars,'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix='af'
              ,time_dependent_predictor='sarcomere', age0 = median(sarcall_altprs[['af']]$age, na.rm=T)
             )

#Stroke PRS x stroke composite
sarcall_cox_agetimescale_altprs_strokecomp <- cox_function(sarcall_altprs[['stroke']], 
             predictor_name='cv_prs_stroke',
             predictor_type='continuous',
             covars=c(covars,'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix='stroke'
             ) 

#VA/CardArrest PRS x VA composite which VA needs tt(sex)
sarcall_cox_agetimescale_altprs_vacomp <- map(c('cv_prs_arrest','cv_prs_va'), ~ cox_function(sarcall_altprs[['va']], 
             predictor_name=.,
             predictor_type='continuous',
             covars=c(covars,'sarcomere'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix='va', time_dependent_predictor='sex', age0 = median(sarcall_altprs[['va']]$age, na.rm=T)
             )) %>% bind_rows()

#Also run the HCM PRS x each of these composites but HCM PRS x AF needs tt(sarcomere) and HCM PRS x VA needs tt(sex)
sarcall_cox_agetimescale_hcmprs_customcomp <- map2(sarcall_altprs[c('hf','stroke')],c('hf','stroke'), ~cox_function(.x, 
              predictor_name='cv_prs',
              predictor_type='continuous',
              covars=c(covars,'sarcomere'), #No age because it is used as timescale
              indicator_varname ='event_indicator',
              time_varname='age_at_event_or_censoring',
              output_dir='genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
              suffix=str_glue('{.y}')
              )) %>% bind_rows()

sarcall_cox_agetimescale_hcmprs_customcomp <- bind_rows(sarcall_cox_agetimescale_hcmprs_customcomp,
          cox_function(sarcall_altprs[['af']],
              predictor_name='cv_prs',
              predictor_type='continuous',
              covars=c(covars,'sarcomere'), #No age because it is used as timescale
              indicator_varname ='event_indicator',
              time_varname='age_at_event_or_censoring',
              output_dir='genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
              time_dependent_predictor='sarcomere', age0= median(sarcall_altprs[['af']]$age, na.rm=T),
              suffix='af'))

sarcall_cox_agetimescale_hcmprs_customcomp <- bind_rows(sarcall_cox_agetimescale_hcmprs_customcomp,
          cox_function(sarcall_altprs[['va']],
              predictor_name='cv_prs',
              predictor_type='continuous',
              covars=c(covars,'sarcomere'), #No age because it is used as timescale
              indicator_varname ='event_indicator',
              time_varname='age_at_event_or_censoring',
              output_dir='genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
              time_dependent_predictor='sex', age0= median(sarcall_altprs[['va']]$age, na.rm=T),
              suffix='va'))

#Merge all the results together
sarcall_cox_agetimescale_customcomp <- bind_rows(
  sarcall_cox_agetimescale_hcmprs_customcomp, sarcall_cox_agetimescale_altprs_hfcomp, sarcall_cox_agetimescale_altprs_af, sarcall_cox_agetimescale_altprs_strokecomp,sarcall_cox_agetimescale_altprs_vacomp
) %>%
  filter(str_detect(term, 'cv_prs'))

#Plot the normal PRS as a forest plot with the alternate PRSs as well.
sarcomere_cox_plot_tb <- sarcall_cox_agetimescale_customcomp %>%
  mutate(term = case_when(
    term == 'cv_prs' ~ 'HCM',
    str_detect(term, 'cv_prs_') ~ str_match(term, 'cv_prs_([A-Za-z\\d]+)')[,2],
    T ~NA)) %>%
    mutate(composite = case_when(
      composite == 'hf' ~ 'HF',
      composite == 'va' ~ 'VA',
      composite == 'stroke' ~ 'Stroke',
      composite == 'af' ~'AF',
      T ~ composite
    )) %>%
    mutate(Group = ifelse(term=='HCM','Continuous PRS (HCM)', 'Continuous PRS (Alt.)')) %>%
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group) %>%
    filter(!is.na(term)) %>%
    mutate(term = case_when(
      term == 'af' ~ 'AF',
      term == 'hf' ~ 'HF',
      term == 'stroke' ~ 'Stroke',
      term == 'arrest' ~ 'Cardiac Arrest (UKB)',
      term == 'va' ~ 'Ventricular\nArrhythmia (UKB/BJJ)',
      T ~ term
    )) %>%
    mutate(term = factor(term, levels=c('HCM','HF','VA','AF','Stroke','Cardiac Arrest (UKB)','Ventricular\nArrhythmia (UKB/BJJ)'), ordered=T)) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Plot the forest plot
forest_plot_cox(
  sarcomere_cox_plot_tb,
  output_dir = 'genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall',
  suffix='_altprs_customcomp', group_palette='Set2', ylabel = 'PRS'
)

rm(sarcall_cox_agetimescale_hcmprs_customcomp,
   sarcall_cox_agetimescale_altprs_hfcomp,
   sarcall_cox_agetimescale_altprs_vacomp,
   sarcall_cox_agetimescale_altprs_strokecomp)
```

This adds Kaplan-Meier curves for categorical versions of these PRS x custom composite signals which are significant.

```{r}
sarcall_afprs_categorical <- prs_categoriser(sarcall_altprs, 'cv_prs_af', quantile_thresholds=c(0.2,0.8))
sarcall_hfprs_categorical <- prs_categoriser(sarcall_altprs, 'cv_prs_hf', quantile_thresholds=c(0.2,0.8))

km_survival_curv_plotter(sarcall_afprs_categorical[['af']],'cv_prs_af_categorical','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/2_KaplanMeier/1_AgeAsTimescale/sarcall/', output_suffix='_afcomp_afprs', ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4)))

km_survival_curv_plotter(sarcall_hfprs_categorical[['hf']],'cv_prs_hf_categorical','event_indicator', 'age_at_event_or_censoring', 'genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/2_KaplanMeier/1_AgeAsTimescale/sarcall/', output_suffix='_hfcomp_hfprs', ylimits=list(c(60,100),c(0.0,0.4),c(0.0,0.4)))

rm(sarcall_cox_agetimescale_altprs, sarcall_cox_agetimescale_altprs_af, sarcall_cox_agetimescale_altprs_vacomp, sarcall_cox_agetimescale_customcomp)
```

Repeat for a quick look at whether Ig PRS associates with composites.

```{r}
alt_prs <- map(list.files('genetic_subset_analysis/PRS/3_output/alt_prs/3_hcmr_scores/Ig/', pattern='sscore', full.names=T), ~read_tsv(.) %>% select(HCR_IDs = IID, SCORE1_AVG) %>% dplyr::rename(cv_prs = SCORE1_AVG))
names(alt_prs) <- str_match(basename(list.files('genetic_subset_analysis/PRS/3_output/alt_prs/3_hcmr_scores/Ig', pattern='sscore', full.names=T)), '([^_]+)\\.sscore$')[,2]

#Define the PRS of interest
altprs_of_interest <- str_c('cv_prs_',names(alt_prs))

#Rename the columns to be cv_prs_<alt_prs_name>
alt_prs <- map2(
  alt_prs,
  names(alt_prs),
  ~ .x |> rename(!!str_c("cv_prs_", .y) := cv_prs)
)

#Join each of the alt_prs to each of the sarcall datasets
merged_alt_prs <- reduce(alt_prs, function(acc, prs_df) {
    left_join(acc, prs_df, by = "HCR_IDs")
  })
rm(alt_prs)

#Join the merged_alt_prs to each of the sarcall datasets
sarcall_altprs <- map(sarcall, ~ left_join(.x, merged_alt_prs, by='HCR_IDs'))

#Scale each of the alt_prs columns within the sarcall datasets
sarcall_altprs <- map(sarcall_altprs, ~ .x |> mutate(across(starts_with("cv_prs_"), ~scale(.)[,1])))
names(sarcall_altprs) <- c('primary','secondary','hf','va','af','stroke')

rm(merged_alt_prs)
```

```{r}
sarcall_cox_agetimescale_altprs_primarycomp <- map(c('cv_prs',altprs_of_interest),
            ~cox_function(sarcall_altprs[[1]], 
             predictor_name=.,
             predictor_type='continuous',
             covars=c(covars,'sarcomere', 'PC1','PC2','PC3'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=str_c('_primary_',.)
             )) %>% bind_rows()

sarcall_cox_agetimescale_altprs_secondarycomp <- map(c('cv_prs',altprs_of_interest),
            ~cox_function(sarcall_altprs[[2]], 
             predictor_name=.,
             predictor_type='continuous',
             covars=c(covars,'sarcomere', 'PC1','PC2','PC3'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=str_c('_secondary_',.)
             )) %>% bind_rows()

#Sarcomere status is time-dependent for secondary composite so refit that model with time-varying coefficient for sarcomere status
sarcall_cox_agetimescale_altprs_secondarycomp <- map(c('cv_prs',altprs_of_interest),
            ~cox_function(sarcall_altprs[[2]], 
             predictor_name=.,
             predictor_type='continuous',
             covars=c(covars,'sarcomere', 'PC1','PC2','PC3'), #No age because it is used as timescale
             indicator_varname ='event_indicator',
             time_varname='age_at_event_or_censoring',
             output_dir='genetic_subset_analysis/PRS/3_output/2_Severity/1_Cox/1_AgeAsTimescale/sarcall/cont/',
             suffix=str_c('_secondary_',.), time_dependent_predictor = 'sarcomere', age0 = median(sarcall_altprs[[2]]$age, na.rm=T)
             )) %>% bind_rows()

#Merge the two tibble results together

sarcall_cox_agetimescale_altprs <- bind_rows(
    sarcall_cox_agetimescale_altprs_primarycomp %>% mutate(composite='primary'),
    sarcall_cox_agetimescale_altprs_secondarycomp %>% mutate(composite='secondary')
) %>%
  filter(str_detect(term, 'cv_prs'))

#Plot the normal PRS as a forest plot with the alternate PRSs as well.
sarcomere_cox_plot_tb <- sarcall_cox_agetimescale_altprs %>%
  mutate(term = case_when(
    term == 'cv_prs' ~ 'HCM',
    str_detect(term, 'cv_prs_') ~ str_match(term, 'cv_prs_([A-Za-z\\d]+)')[,2],
    T ~NA)) %>%
    mutate(composite = str_to_title(composite)) %>%
    mutate(Group = ifelse(term=='HCM','Continuous PRS (HCM)', 'Continuous PRS (Alt.)')) %>%
    select(term,n_events, composite, hr, hr_low, hr_high, p.value, Group) %>%
    filter(!is.na(term)) %>%
    mutate(adj_pvalue = p.adjust(p.value, method='fdr'))

#Plot the forest plot
forest_plot_cox(
  sarcomere_cox_plot_tb,
  output_dir = 'genetic_subset_analysis/PRS/3_output/alt_prs/4_Severity/1_Cox/1_AgeAsTimescale/sarcall',
  suffix='_altprs_Ig', group_palette='Set2', ylabel='PRS'
)

rm(sarcall_cox_agetimescale_altprs_primarycomp, sarcall_cox_agetimescale_altprs_secondarycomp, sarcall_cox_agetimescale_altprs)
```