---
title: "Joint Modelling of All Predictors vs Endophenotypes in HCMR"
author: "Jonathan Chan"
date: "2026-06-02"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(tictoc)
library(kableExtra)
library(readxl)
library(ggrepel)
library(patchwork)
library(gtsummary)

rm(list=ls())

theme_set(theme_classic(base_size=14))
source('genetic_subset_analysis/common/hcmr_endophenotype_functions.R')
```

# Overview

This script mirrors the endophenotype-association analysis in `1_HCMR_PRS_Endophenotype_Analysis.Rmd` but, instead of focusing on the PRS and sarcomere status alone, it fits a **single joint linear model per endophenotype** that additionally includes the clinical risk factors (CRFs) and plasma proteins as predictors:

- The HCM **PRS** (`cv_prs`, per-SD)
- **Sarcomere status** (positive vs negative)
- **Clinical risk factors**: BMI, DBP, smoking (ever vs never) and the prior-history binary flags
- **Plasma proteins**: NT-proBNP and Troponin T (`TnTStat`)

Age, sex, MRI field strength and the first 3 genetic PCs (plus BSA for wall thickness) are included only as adjustment covariates and are **not** reported. Each endophenotype is standardised to SD units so the effect sizes are comparable.
Outputs are (1) a forest plot of every reported predictor for each endophenotype with significance stars and (2) a volcano-style summary scatter plot of every predictor x endophenotype comparison.

# Import

This imports the HCMR data after preprocessing (identical inputs to script 1).

```{r}
total_df_notransform <- readRDS('output/data/total_notransform_nogenofilter_df.rds')

#Use the 2446 individuals
hcmr_ids <- read_tsv('genetic_subset_analysis/common/output/data/hcmr_agesex_rare_common_filtered_2446.tsv')

#Convert false positive individuals to sarcomere-negative individuals (due to false positive MYBPC3-VUS variants)
falsepositive_mybpc3_vus_individuals <- readRDS('genetic_subset_analysis/common/falsepositive_mybpc3_vus_individuals.rds')
total_df_notransform <- total_df_notransform %>%
  filter(HCR_IDs %in% hcmr_ids$HCR_IDs) %>% #Filter to 2446
  mutate(sarcomere = ifelse(HCR_IDs %in% falsepositive_mybpc3_vus_individuals, FALSE, sarcomere))
rm(hcmr_ids)

#Add in the HCMR PRS scores (this is with the minimac-r2-filter for INFO 0.3-1.0) and computed with the intersectvars with UKB
hcmr_prs <- read_tsv('genetic_subset_analysis/PRS/1_input/hcmr_imputedvars_prs.sscore')

# Add in the HCMR PCs which were derived from flashPCA2 from the HCMR hard-called genotypes
hcmr_pcs <- read_delim('popgen/2_gwas/data/hcmr/HCMR_HARDCALLS/pcs_hcmr_10pcs')

#---- Predictors -------------------------------------------------------------
#Clinical risk factors (column name -> display label)
crf_labels <- c(
  'bmi'      = 'BMI (per SD)',
  'cmrdiabp' = 'DBP (per SD)',
  'smoking'  = 'Smoking (ever)',
  'hxscd'    = 'Prev. SCD',
  'hxafib'   = 'Prev. AF',
  'hxvt'     = 'Prev. VT',
  'hxhf'     = 'Prev. HF',
  'hxstrk'   = 'Prev. Stroke',
  'diatyp2'  = 'Prev. T2D',
  'blcp'     = 'Prev. Chest Pain',
  'bldysp'   = 'Prev. Dyspnea',
  'hxsync'   = 'Prev. Syncope'
)
crfs <- names(crf_labels)

#Plasma proteins (column name -> display label)
protein_labels <- c('NTproBNP' = 'NTproBNP (per SD log-scale)', 'TnTStat' = 'Troponin T (per SD log-scale)')
proteins <- names(protein_labels)

#Continuous predictors standardised per-SD; proteins are heavily skewed so log1p first
cont_predictors <- c('bmi','cmrdiabp')
binary_crfs     <- setdiff(crfs, cont_predictors) #prior-history + smoking 0/1 flags, left on the native scale (hxafib/hxvt/smoking recoded to 0/1 in preprocessing)

#Predictors reported in the figures (everything except age/sex/3PCs/BSA)
extra_predictors <- c('sarcomere', crfs, proteins)        #reported alongside cv_prs
reported_predictors <- c('cv_prs', extra_predictors)

#Endophenotypes of interest (column -> human label), mirroring script 1
continuous_param_list <- list(
  c('wallthkmax','Maximal LV Wall Thickness (mm)'),
  c('lvmassi','LV Mass Index (g/m2)'),
  c('lge_total','LV LGE Mass/ LV Mass (%)'),
  c('ecvfwhole','LV Extracellular Volume Fraction (%)'),
  c('lvef','LV Ejection Fraction (%)'),
  c('lvs_ellmean','-Longitudinal LV Strain'),
  c('lvs_eccmean','-Circumferential LV Strain'),
  c('lvs_errmean','Radial LV Strain')
)
endophenos <- map_chr(continuous_param_list, 1)

#Lookup tables for relabelling / colouring
pheno_labels <- tibble(
  pheno = endophenos,
  Label = map_chr(continuous_param_list, 2),
  Label_short = c('maxLVWT','LVMi','LGE','ECVF','LVEF','-LongStrain','-CircStrain','RadStrain')
)

predictor_label_map <- c('cv_prs' = 'PRS (per SD)', 'sarcomere' = 'Sarcomere-positive',
                         crf_labels, protein_labels)

#Display order (y-axis) for the forest plot: PRS, Sarcomere, proteins, continuous CRFs, then binary CRFs
predictor_order <- c('PRS (per SD)','Sarcomere-positive',
                     protein_labels[['NTproBNP']], protein_labels[['TnTStat']],
                     crf_labels[['bmi']], crf_labels[['cmrdiabp']],
                     crf_labels[binary_crfs])

#Assign each predictor to a category for colouring
group_of <- function(t){
  case_when(
    t == 'cv_prs'    ~ 'PRS',
    t == 'sarcomere' ~ 'Sarcomere',
    t %in% proteins  ~ 'Plasma Protein',
    TRUE             ~ 'Clinical Risk Factor'
  )
}
```

## Investigating the discordant sarcomere associations (maxLVWT vs LVMi)

The joint model gives sarcomere-positive status a **positive** effect on maximal wall thickness
(`wallthkmax`) but a **negative** effect on LV mass index (`lvmassi`), even though `wallthkmax` and
`lvmassi` are positively correlated overall.

A BSA-indexing artefact was the first hypothesis but is **ruled out**: the discordance is already present
in the *raw* values — sarcomere-positive patients have **higher** maximal wall thickness yet **lower**
LV mass (both indexed *and* unindexed) — and adjusting for age/sex/BSA does **not** flip either sign.
So the effect is real, not an artefact of indexing or the covariate set.

The remaining explanation is **morphological**: sarcomeric HCM is characteristically **focal/asymmetric**
hypertrophy — a single segment (typically septal) is very thick, driving up `wallthkmax`, while the
*total* myocardial mass stays lower (smaller cavity, less of the global/concentric remodelling seen in
the predominantly older, hypertensive sarcomere-negative group). `wallthkmax` captures the single
thickest segment (a *focal* measure) whereas `lvmassi` captures *global* mass, so sarcomere status pushes
the two in opposite directions even though they are positively correlated population-wide.

This block decomposes the paradox with four diagnostic panels:
(1) the joint scatter with per-group regression lines and group-stratified Spearman correlations
(sarcomere-positive should sit **down-and-right** — thicker peak wall at lower mass);
(2) the raw marginal distribution of each phenotype by sarcomere status (the discordance pre-adjustment);
(3) the sarcomere effect on each (SD-standardised) phenotype **unadjusted vs adjusted** for age/sex/BSA,
confirming the signs are stable (no indexing/covariate flip); and
(4) a **focality index** — peak wall thickness per unit LV mass — which should be **higher** in
sarcomere-positive if hypertrophy is concentrated focally rather than spread globally.

```{r sarcomere_discordance, fig.width=12, fig.height=11}
invest_dir <- 'genetic_subset_analysis/PRS/3_output/1_Endophenotype/2_LM/sarcall/allpredictor/sarcomere_discordance/'
dir.create(invest_dir, recursive = TRUE, showWarnings = FALSE)

#Covariates available at this point (PCs/proteins not yet joined); BSA is the key indexing covariate
adj_covars_invest <- intersect(c('age','gender','bsa'), names(total_df_notransform))

invest_df <- total_df_notransform %>%
  mutate(sarc = factor(ifelse(sarcomere, 'Positive', 'Negative'), levels = c('Negative','Positive'))) %>%
  filter(!is.na(wallthkmax), !is.na(lvmassi))

sarc_pal <- c('Negative' = 'grey55', 'Positive' = '#D55E00')

#--- (1) Joint scatter + per-group regression lines, with stratified Spearman -------------------
overall_cor <- cor(invest_df$wallthkmax, invest_df$lvmassi, method = 'spearman', use = 'pairwise.complete.obs')
print(overall_cor)
strat_cor <- invest_df %>%
  group_by(sarc) %>%
  summarise(rho = cor(wallthkmax, lvmassi, method = 'spearman', use = 'pairwise.complete.obs'),
            n = sum(is.finite(wallthkmax) & is.finite(lvmassi)), .groups = 'drop')
print(strat_cor)  #Spearman correlation within each sarcomere group

cor_subtitle <- str_glue('Overall Spearman ρ = {round(overall_cor,2)}  |  ',
                         'Neg ρ = {round(strat_cor$rho[strat_cor$sarc=="Negative"],2)}, ',
                         'Pos ρ = {round(strat_cor$rho[strat_cor$sarc=="Positive"],2)}')

p_scatter <- ggplot(invest_df, aes(wallthkmax, lvmassi, colour = sarc)) +
  geom_point(alpha = 0.22, size = 1) +
  geom_smooth(method = 'lm', se = TRUE, linewidth = 1) +
  scale_colour_manual(values = sarc_pal) +
  labs(title = 'maxLVWT vs LVMi by sarcomere status', subtitle = cor_subtitle,
       x = 'Maximal LV Wall Thickness (mm)', y = 'LV Mass Index (g/m2)', colour = 'Sarcomere')

#--- (2) Raw marginal distribution of each phenotype by sarcomere -------------------------------
#Wilcoxon rank-sum (Pos vs Neg) for each phenotype, surfaced in the subtitle
marg_p <- c(
  wallthkmax = wilcox.test(wallthkmax ~ sarc, data = invest_df)$p.value,
  lvmassi    = wilcox.test(lvmassi  ~ sarc, data = invest_df)$p.value
)
marg_subtitle <- str_glue('Wilcoxon Pos vs Neg — maxLVWT p = {signif(marg_p[["wallthkmax"]], 2)}; ',
                          'LVMi p = {signif(marg_p[["lvmassi"]], 2)}')

p_marg <- invest_df %>%
  select(sarc, wallthkmax, lvmassi) %>%
  pivot_longer(c(wallthkmax, lvmassi), names_to = 'pheno', values_to = 'value') %>%
  mutate(pheno = recode(pheno,
                        wallthkmax = 'Maximal LV Wall Thickness (mm)',
                        lvmassi    = 'LV Mass Index (g/m2)')) %>%
  ggplot(aes(sarc, value, fill = sarc)) +
  geom_violin(alpha = 0.4, trim = TRUE) +
  geom_boxplot(width = 0.15, outlier.size = 0.4, alpha = 0.7) +
  facet_wrap(~ pheno, scales = 'free_y') +
  scale_fill_manual(values = sarc_pal) +
  labs(title = 'Raw marginal distributions by sarcomere status', subtitle = marg_subtitle,
       x = NULL, y = 'Raw value', fill = 'Sarcomere') +
  theme(legend.position = 'none')

#--- (3) Sarcomere effect: unadjusted vs adjusted vs +Morphology (SD-standardised outcomes) -----
#Direct test of the focal mechanism: if LV morphology mediates the discordance, adding Morphology as
#a covariate should attenuate the sarcomere-positive effect on BOTH maxLVWT and LVMi toward zero.
std <- function(x) as.numeric(scale(x))
invest_std <- invest_df %>%
  mutate(across(c(wallthkmax, lvmassi), std),
         Morphology = factor(as.character(Morphology)))

adj_lab  <- str_glue('Adjusted ({paste(adj_covars_invest, collapse="+")})')
morph_lab <- str_glue('{adj_lab} + Morphology')

sarc_contrast <- function(y){
  raw   <- lm(reformulate('sarc', y), data = invest_std)
  adj   <- lm(reformulate(c('sarc', adj_covars_invest), y), data = invest_std)
  morph <- lm(reformulate(c('sarc', adj_covars_invest, 'Morphology'), y), data = invest_std)
  pull_beta <- function(m, model){
    co <- summary(m)$coefficients['sarcPositive', ]
    tibble(pheno = y, model = model, beta = co['Estimate'], se = co['Std. Error'], n = nobs(m))
  }
  bind_rows(pull_beta(raw, 'Unadjusted'),
            pull_beta(adj, adj_lab),
            pull_beta(morph, morph_lab))
}

contrasts <- map_dfr(c('wallthkmax','lvmassi'), sarc_contrast) %>%
  mutate(lo = beta - 1.96*se, hi = beta + 1.96*se,
         pheno = recode(pheno, wallthkmax = 'maxLVWT', lvmassi = 'LVMi'),
         model = factor(model, levels = c('Unadjusted', adj_lab, morph_lab)))
print(contrasts)  #sarcomere-positive beta (SD) per model; expect attenuation toward 0 with Morphology

p_contrast <- ggplot(contrasts, aes(beta, pheno, colour = model)) +
  geom_vline(xintercept = 0, linetype = 'dashed', colour = 'grey50') +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.2, position = position_dodge(width = 0.5)) +
  geom_point(size = 2.5, position = position_dodge(width = 0.5)) +
  scale_colour_manual(values = c('#0072B2', '#D55E00', '#009E73')) +
  labs(title = 'Sarcomere-positive effect: unadjusted vs adjusted vs +Morphology',
       subtitle = 'Outcomes standardised to SD; adding Morphology should pull both effects toward zero',
       x = 'Sarcomere-positive effect (SD)', y = NULL, colour = NULL) +
  theme(legend.position = 'bottom')

#--- (4) Focality index: peak wall thickness per unit LV mass -----------------------------------
#High peak thickness with low total mass = focal/asymmetric hypertrophy. Use unindexed lvmass if
#present, else fall back to the indexed LVMi (the ratio's direction across groups is unaffected).
mass_col   <- if ('lvmass' %in% names(invest_df)) 'lvmass' else 'lvmassi'
mass_label <- if (mass_col == 'lvmass') 'LV mass (g)' else 'LVMi (g/m2)'

focality_df <- invest_df %>%
  filter(is.finite(wallthkmax), is.finite(.data[[mass_col]]), .data[[mass_col]] > 0) %>%
  mutate(focality = wallthkmax / .data[[mass_col]])

#Wilcoxon test of the focality index between groups
foc_p <- wilcox.test(focality ~ sarc, data = focality_df)$p.value
foc_med <- focality_df %>% group_by(sarc) %>% summarise(med = median(focality), .groups = 'drop')
print(foc_med)  #Median focality (peak WT per unit mass) by group

p_focality <- ggplot(focality_df, aes(sarc, focality, fill = sarc)) +
  geom_violin(alpha = 0.4, trim = TRUE) +
  geom_boxplot(width = 0.15, outlier.size = 0.4, alpha = 0.7) +
  scale_fill_manual(values = sarc_pal) +
  labs(title = 'Focality index: peak wall thickness per unit mass',
       subtitle = str_glue('maxLVWT / {mass_label}  (Wilcoxon p = {signif(foc_p, 2)})'),
       x = NULL, y = str_glue('maxLVWT / {mass_label}'), fill = 'Sarcomere') +
  theme(legend.position = 'none')

discordance_plot <- (p_scatter + p_marg) / (p_contrast + p_focality) +
  plot_annotation(title = 'Why is sarcomere +ve for maxLVWT but -ve for LVMi?',
                  theme = theme(plot.title = element_text(face = 'bold', size = 15)))

ggsave(file.path(invest_dir, 'sarcomere_maxLVWT_LVMi_discordance.png'),
       discordance_plot, width = 16, height = 11, dpi = 600)

discordance_plot
```

### Morphology confirms the focal mechanism

The higher focality index in sarcomere-positive patients points to asymmetric/focal hypertrophy. The
`Morphology` factor lets us test this directly: its levels split into **asymmetric/focal** patterns
(*Localised basal septal*, *Reverse curve septal*, *Apical*) versus **global** hypertrophy (*Concentric*),
with *Mid-cavity obstruction with apical aneurysm* and *Other* kept separate. If the mechanism holds,
sarcomere-positive patients should be **enriched for focal patterns and depleted of Concentric**, and the
focal morphologies should themselves carry **high `wallthkmax` but lower `lvmassi`** — reproducing the
discordance at the morphology level.

```{r sarcomere_morphology, fig.width=12, fig.height=15}
#Morphology level order as stored in the data; group into focal vs global remodelling patterns
morph_levels <- c('Localised basal septal','Reverse curve septal','Apical',
                  'Concentric','Mid-cavity obstruction with apical aneurysm','Other')

morph_df <- total_df_notransform %>%
  mutate(sarc = factor(ifelse(sarcomere, 'Positive', 'Negative'), levels = c('Negative','Positive'))) %>%
  filter(!is.na(Morphology)) %>%
  mutate(Morphology = factor(as.character(Morphology), levels = morph_levels),
         pattern = case_when(
           Morphology == 'Concentric' ~ 'Concentric (global)',
           Morphology == 'Other'      ~ 'Other',
           TRUE                       ~ 'Asymmetric / focal'))

pattern_pal <- c('Asymmetric / focal' = '#D55E00', 'Concentric (global)' = '#0072B2', 'Other' = 'grey60')

#--- (A) Morphology composition within each sarcomere group (chi-square on the full table) -------
morph_chisq_p <- suppressWarnings(chisq.test(table(morph_df$sarc, morph_df$Morphology))$p.value)
print(prop.table(table(morph_df$sarc, morph_df$Morphology), margin = 1))  #row-wise composition

p_morph_comp <- morph_df %>%
  count(sarc, Morphology) %>%
  group_by(sarc) %>% mutate(prop = n / sum(n)) %>% ungroup() %>%
  ggplot(aes(sarc, prop, fill = Morphology)) +
  geom_col(colour = 'white', linewidth = 0.2) +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = 'LV morphology composition by sarcomere status',
       subtitle = str_glue('Chi-square p = {signif(morph_chisq_p, 2)}'),
       x = NULL, y = 'Proportion', fill = 'Morphology')

#--- (B) maxLVWT and LVMi by morphology (ordered by median peak wall thickness) ------------------
morph_order <- morph_df %>%
  group_by(Morphology) %>% summarise(m = median(wallthkmax, na.rm = TRUE), .groups = 'drop') %>%
  arrange(m) %>% pull(Morphology)

p_morph_pheno <- morph_df %>%
  select(Morphology, pattern, wallthkmax, lvmassi) %>%
  pivot_longer(c(wallthkmax, lvmassi), names_to = 'pheno', values_to = 'value') %>%
  filter(is.finite(value)) %>%
  mutate(pheno = recode(pheno, wallthkmax = 'maxLVWT (mm)', lvmassi = 'LVMi (g/m2)'),
         Morphology = factor(Morphology, levels = morph_order)) %>%
  ggplot(aes(Morphology, value, fill = pattern)) +
  geom_boxplot(outlier.size = 0.4) +
  facet_wrap(~ pheno, scales = 'free_x') +
  coord_flip() +
  scale_fill_manual(values = pattern_pal) +
  labs(title = 'maxLVWT and LVMi by LV morphology',
       x = NULL, y = 'Raw value', fill = 'Pattern') +
  theme(legend.position = 'bottom')

#Vertical stack: raw maxLVWT/LVMi by sarcomere (Wilcoxon), then morphology composition, then
#morphology-resolved maxLVWT/LVMi — one combined figure of the focal mechanism
morph_plot <- p_marg / p_morph_comp / p_morph_pheno +
  plot_layout(heights = c(1, 1, 1.3)) +
  plot_annotation(title = 'Sarcomeric HCM is focal: morphology enrichment explains the maxLVWT/LVMi discordance',
                  theme = theme(plot.title = element_text(face = 'bold', size = 14)))

ggsave(file.path(invest_dir, 'sarcomere_morphology_mechanism.png'),
       morph_plot, width = 12, height = 15, dpi = 600)

morph_plot
```


# Baseline Summary Table

This produces a baseline-characteristics (Table 1) summary of the analysed HCMR cohort using `gtsummary`, mirroring the phenotypes reported in the rare-variant baseline summary table. Continuous variables are shown as mean (SD) and binary flags as n (%). No subgroup comparison is made — a single overall column across the whole cohort is reported, and the variable recodings mirror those in the rare-variant baseline table. Related phenotypes are grouped under bold section headers (Ethnicity, Family History of HCM, Comorbidities, NYHA Functional Class, etc.) by stacking one `tbl_summary` per section.

```{r baseline_summary_table}
#Recreate the baseline phenotype definitions (recodings mirror the rare-variant baseline table)
baseline_tb <- total_df_notransform %>%
  mutate(
    genderMale     = gender == 'Male',
    race_white     = race == 'White',
    race_black     = race == 'Black/Afr.Am',
    race_SA        = race == 'SA',
    race_EA        = race == 'EA',
    race_other     = race %in% c('Pacific Islander', ' American Indian/Alaskan'),
    hxhcm1st       = hxhcm1st >= 1,
    hxhcm2nd       = hxhcm2nd >= 1,
    hxhcm1st2nd    = hxhcm1st & hxhcm2nd,
    nyha1          = blnyha == 1,
    nyha2          = blnyha == 2,
    nyha3_4        = blnyha %in% c(3, 4),
    hxafib_persist = hxafib == 'Persistent',
    hxafib_paroxy  = hxafib == 'Paroxysmal'
  )

#0/1 raw binary flags -> logical so gtsummary treats them as dichotomous (positive count only)
raw_binary_flags <- c('hxhtn','diatyp2','cursmoke','hxsync','hxhf','hxstrk','holtnsvt',
                      'blcp','bldysp','blbeta','blcalc','blacearb','bldisop','blstat',
                      'bldiur','blantico','blantipl')
baseline_tb <- baseline_tb %>%
  mutate(across(any_of(raw_binary_flags), ~ . == 1))

#Display labels (column -> row label, as it appears under its section header)
baseline_labels <- c(
  age            = 'Age (years)',
  genderMale     = 'Male',
  race_white     = 'White',
  race_black     = 'Black',
  race_SA        = 'South Asian',
  race_EA        = 'East Asian',
  race_other     = 'Other',
  bmi            = 'BMI (kg/m2)',
  hxhcm1st       = '1st degree',
  hxhcm2nd       = '2nd degree',
  hxhcm1st2nd    = 'Both 1st and 2nd',
  hxhtn          = 'Hypertension',
  diatyp2        = 'Type 2 Diabetes',
  cursmoke       = 'Current Smoker',
  nyha1          = 'I',
  nyha2          = 'II',
  nyha3_4        = 'III or IV',
  hxsync         = 'Syncope',
  hxhf           = 'Heart Failure',
  hxstrk         = 'Stroke',
  holtnsvt       = 'Non-sustained Ventricular Tachycardia',
  holvtrun       = 'Number of VT Runs',
  hxafib_persist = 'Atrial Fibrillation - Persistent',
  hxafib_paroxy  = 'Atrial Fibrillation - Paroxysmal',
  blcp           = 'Chest Pain',
  bldysp         = 'Dyspnea',
  blbeta         = 'Beta-blockers',
  blcalc         = 'Calcium-channel blocker',
  blacearb       = 'ACE/ARB',
  bldisop        = 'Disopyramide',
  blstat         = 'Statins',
  bldiur         = 'Diuretics',
  blantico       = 'Oral anticoagulant',
  blantipl       = 'Oral antiplatelet agent'
)

#Section structure: each block is a contiguous run of variables, optionally under a bold
#group header (NA = standalone rows, no header). Stacked together to get the factored layout.
baseline_sections <- list(
  list(header = NA_character_, vars = c('age','genderMale')),
  list(header = 'Ethnicity',              vars = c('race_white','race_black','race_SA','race_EA','race_other')),
  list(header = NA_character_, vars = 'bmi'),
  list(header = 'Family History of HCM',  vars = c('hxhcm1st','hxhcm2nd','hxhcm1st2nd')),
  list(header = 'Comorbidities',          vars = c('hxhtn','diatyp2','cursmoke')),
  list(header = 'NYHA Functional Class',  vars = c('nyha1','nyha2','nyha3_4')),
  list(header = 'Other clinical history', vars = c('hxsync','hxhf','hxstrk','holtnsvt','holvtrun','hxafib_persist','hxafib_paroxy')),
  list(header = 'Symptoms at Enrollment', vars = c('blcp','bldysp')),
  list(header = 'Medications at Enrollment', vars = c('blbeta','blcalc','blacearb','bldisop','blstat','bldiur','blantico','blantipl'))
)

#One tbl_summary per section, dropping any variables absent from the data
make_section_tbl <- function(vars){
  vars <- intersect(vars, names(baseline_tb))
  if(length(vars) == 0) return(NULL)
  label_list <- map2(vars, unname(baseline_labels[vars]),
                     ~ rlang::new_formula(rlang::sym(.x), .y))
  baseline_tb %>%
    select(all_of(vars)) %>%
    tbl_summary(
      label     = label_list,
      statistic = list(all_continuous() ~ '{mean} ({sd})',
                       all_categorical() ~ '{n} ({p}%)'),
      digits    = all_continuous() ~ 1,
      missing   = 'no'
    )
}

section_tbls <- map(baseline_sections, ~ make_section_tbl(.x$vars))
keep         <- !map_lgl(section_tbls, is.null)

#Stack the sections, using the headers as bold row-group labels (NA -> ungrouped rows)
baseline_summary_tbl <- tbl_stack(
  section_tbls[keep],
  group_header = map_chr(baseline_sections, 'header')[keep],
  quiet = TRUE
) %>%
  modify_header(label ~ '**Characteristic**') %>%
  bold_labels()

#Save as html and render inline
baseline_out_dir <- 'genetic_subset_analysis/PRS/3_output/1_Endophenotype/2_LM/sarcall/allpredictor/'
dir.create(baseline_out_dir, recursive = TRUE, showWarnings = FALSE)
gtsummary::as_gt(baseline_summary_tbl) %>%
  gt::gtsave(file.path(baseline_out_dir, 'baseline_summary_table.html'))

baseline_summary_tbl
```

# Data Preprocessing

This filters to the analysed cohort, joins the PRS (scaled per-SD) and PCs, and retains the CRF, plasma-protein and endophenotype columns. 
`cmrdiabp` is kept on its original name and relabelled to DBP only in the figures (avoiding a rename-before-select ordering issue).

```{r}
data_preprocessor <- function(total_df, hcmr_prs, vars_keep, sarcomere_status=c(T,F), prs_scale=TRUE, hcmr_pcs=NULL){
  out <- total_df %>%
    filter(sarcomere %in% sarcomere_status) %>%
    left_join(select(hcmr_prs, HCR_IDs=IID, cv_prs=SCORE1_AVG), by='HCR_IDs')

  if(isTRUE(prs_scale)){
    out <- mutate(out, cv_prs = scale(cv_prs)[,1])
  }

  #Derive smoking as a 0/1 ever-vs-never flag from the current/past smoking fields (mirrors script 1)
  if(all(c('cursmoke','pstsmoke') %in% names(out))){
    out <- mutate(out, smoking = case_when(
      cursmoke == 1 | pstsmoke == 1 ~ 1,   #current OR past smoker -> ever
      cursmoke == 0 & pstsmoke == 0 ~ 0,   #neither -> never
      TRUE ~ NA_real_
    ))
  }

  #Recode categorical prior-history fields to 0/1 (any positive category -> 1, 'None' -> 0)
  if('hxafib' %in% names(out)){
    out <- mutate(out, hxafib = case_when(
      hxafib %in% c('Persistent','Paroxysmal') ~ 1,
      hxafib %in% 'None' ~ 0,
      TRUE ~ NA_real_
    ))
  }
  if('hxvt' %in% names(out)){
    out <- mutate(out, hxvt = case_when(
      hxvt %in% c('Sustained','Nonsustained') ~ 1,
      hxvt %in% 'None' ~ 0,
      TRUE ~ NA_real_
    ))
  }

  out <- out %>%
    select(HCR_IDs, age, sex=gender, bsa, FieldStrength, sarcomere, cv_prs, any_of(vars_keep))

  if(!is.null(hcmr_pcs)){
    out <- left_join(out, select(hcmr_pcs, HCR_IDs=IID, PC1:PC10), by='HCR_IDs')
  }
  out
}

sarcall <- data_preprocessor(total_df_notransform, hcmr_prs,
                             vars_keep = c(crfs, proteins, endophenos),
                             sarcomere_status=c(T,F), hcmr_pcs=hcmr_pcs)

rm(total_df_notransform, hcmr_prs)

#Sanity check: every reported predictor + endophenotype is present
setdiff(c(reported_predictors, endophenos), names(sarcall))
```

This standardises the continuous outcomes and predictors to SD units (after log1p for the skewed LGE outcome and the plasma proteins) so the betas are directly comparable; the binary CRFs and sarcomere status are left on the native 0/1 scale.

```{r}
#Toggle: TRUE = log1p the skewed vars before standardising; FALSE = standardise only.
#Standardisation is linear so does not affect inference on its own; this flag only controls
#whether the skewed LGE outcome + plasma proteins are log-transformed (the sole nonlinear step).
apply_log <- TRUE

#Heavily right-skewed continuous variables considered for log1p
log_cols <- c('lge_total', proteins)

#Snapshot the raw (untransformed) values for the before/after distribution plots
dist_raw <- sarcall %>%
  select(any_of(log_cols)) %>%
  mutate(across(everything(), as.numeric))

if (isTRUE(apply_log)) {
  sarcall <- sarcall %>%
    mutate(across(any_of(log_cols), ~ log1p(as.numeric(.))))
}

#Standardise continuous outcomes + continuous predictors to SD (cv_prs already scaled)
sd_cols <- c(endophenos, cont_predictors, proteins)
sarcall_scaled <- sarcall %>%
  mutate(across(any_of(sd_cols), ~ as.numeric(scale(.)))) %>%
  mutate(across(any_of(binary_crfs), as.numeric))  #ensure native-scale numeric 0/1

#Double-check every prior-history/smoking CRF is a strict 0/1 flag (NA allowed); stop if any other value remains
binary_bad <- sarcall_scaled %>%
  select(any_of(binary_crfs)) %>%
  map_lgl(~ any(!(.x %in% c(0, 1, NA))))
if (any(binary_bad)) {
  stop('Non-0/1 values remain in binary CRF(s): ', paste(names(binary_bad)[binary_bad], collapse = ', '))
}
```

This checks the effect of the log1p + SD standardisation on the skewed variables (LGE and the plasma proteins), plotting the raw distribution (left column) against the transformed distribution (right column) for each.

```{r, fig.width=10, fig.height=10}
#Display names + raw-scale units for the skewed variables
transformed_var_disp <- c(
  'lge_total' = 'Total LGE (%)',
  'NTproBNP'  = 'NT-proBNP (pg/mL)',
  'TnTStat'   = 'Troponin T (ng/L)'
)
stage_fill <- c('Before' = '#E69F00', 'After' = '#56B4E9')

density_panel <- function(values, title, xlab, fill){
  ggplot(tibble(value = values[is.finite(values)]), aes(x = value)) +
    geom_density(fill = fill, alpha = 0.6) +
    labs(title = title, x = xlab, y = 'Density') +
    theme_classic(base_size = 12)
}

#Build before (raw) and after (log1p + SD) panels for each variable, ordered row-wise
beforeafter_panels <- map(names(transformed_var_disp), function(v){
  list(
    density_panel(dist_raw[[v]],        str_glue('{transformed_var_disp[[v]]}\nBefore (raw)'),
                  transformed_var_disp[[v]], stage_fill[['Before']]),
    density_panel(sarcall_scaled[[v]],  str_glue('{transformed_var_disp[[v]]}\nAfter (log1p + standardisation)'),
                  'Standardised value (SD of log)', stage_fill[['After']])
  )
}) %>% flatten()

beforeafter_plot <- wrap_plots(beforeafter_panels, ncol = 2, byrow = TRUE) +
  plot_annotation(title = 'Effect of log1p + SD Standardisation on Skewed Variables')

dir.create('genetic_subset_analysis/PRS/3_output/1_Endophenotype/2_LM/sarcall/allpredictor/',
           recursive = TRUE, showWarnings = FALSE)
ggsave('genetic_subset_analysis/PRS/3_output/1_Endophenotype/2_LM/sarcall/allpredictor/transformation_before_after_distributions.png',
       beforeafter_plot, width = 10, height = 10, dpi = 600)

beforeafter_plot
```

This visualises the distribution of each (standardised) endophenotype.

```{r}
sarcall_scaled %>%
  select(all_of(endophenos)) %>%
  pivot_longer(everything(), names_to='pheno', values_to='value') %>%
  ggplot(aes(x=value)) +
  geom_density(fill='lightblue', alpha=0.5) +
  facet_wrap(~pheno, scales='free') +
  labs(title='Density Plots of Standardised Endophenotypes')
```

# Joint Linear Regression

For each endophenotype this fits the joint model
`pheno ~ age + sex + 3PCs (+ BSA for wall thickness) + cv_prs + sarcomere + CRFs + plasma proteins`.
`linear_logistic_regression_modeller` reports the PRS plus every additional predictor from this single mutually-adjusted fit, so all effects are conditional on one another.

```{r}
adjust_covars  <- c('age','sex','FieldStrength','PC1','PC2','PC3')        #not reported; BSA auto-added for wallthkmax
model_covars   <- c(adjust_covars, extra_predictors)       #full RHS (minus the main predictor cv_prs)

joint_lm <- map(continuous_param_list, ~ linear_logistic_regression_modeller(
  sarcall_scaled,
  .x,
  predictor = 'cv_prs',
  predictor_type = 'continuous',
  covars = model_covars,
  additional_predictors_to_report = extra_predictors
)) %>% bind_rows()

#Attach phenotype labels + colour, then FDR-correct across all predictor x endophenotype tests
joint_lm <- joint_lm %>%
  left_join(pheno_labels, by='pheno') %>%
  colour_hardcoder() %>%
  mutate(adj_p = p.adjust(pval, method='fdr'),
         predictor_label = recode(predictor_reported, !!!predictor_label_map),
         Group = group_of(predictor_reported),
         #Human-readable effect string: beta [lower, upper] 95% CI (SD units)
         beta_ci = sprintf('%.2f [%.2f, %.2f]', estimate, estimate - 1.96*SE, estimate + 1.96*SE))

out_dir <- 'genetic_subset_analysis/PRS/3_output/1_Endophenotype/2_LM/sarcall/allpredictor/'
dir.create(out_dir, recursive=TRUE, showWarnings=FALSE)
write_tsv(joint_lm %>% select(pheno, Label, predictor_reported, predictor_label, Group, estimate, SE, beta_ci, pval, adj_p),
          file.path(out_dir, 'summary_allpredictor_joint_lm_results.tsv'))
```

### Raw-Unit Outcome Models

This refits the identical joint models but with each endophenotype left in its **raw units** —
no SD standardisation and no log1p for LGE. The predictors keep their original treatment
(`cv_prs` per-SD, proteins log1p + SD, continuous CRFs per-SD, binary CRFs/sarcomere native 0/1),
so the betas are now interpretable in each endophenotype's native units (mm, g/m2, %, etc.).
The raw outcomes are reconstructed from `sarcall` (non-logged endophenotypes are still raw there)
and `dist_raw` (the snapshot of raw, un-logged LGE) since `total_df_notransform` has been removed.

```{r}
#Raw endophenotypes: non-log outcomes are untouched in `sarcall`; restore raw LGE from dist_raw
raw_endos_df <- sarcall %>%
  select(all_of(endophenos)) %>%
  mutate(lge_total = dist_raw$lge_total) %>%
  mutate(ecvfwhole = ecvfwhole * 100) #To convert to %

#Swap the scaled outcomes for raw ones; predictors stay as in sarcall_scaled (same row order)
sarcall_rawpheno <- sarcall_scaled %>%
  mutate(across(all_of(endophenos), ~ raw_endos_df[[cur_column()]]))

joint_lm_raw <- map(continuous_param_list, ~ linear_logistic_regression_modeller(
  sarcall_rawpheno,
  .x,
  predictor = 'cv_prs',
  predictor_type = 'continuous',
  covars = model_covars,
  additional_predictors_to_report = extra_predictors
)) %>% bind_rows()

joint_lm_raw <- joint_lm_raw %>%
  left_join(pheno_labels, by='pheno') %>%
  colour_hardcoder() %>%
  mutate(adj_p = p.adjust(pval, method='fdr'),
         predictor_label = recode(predictor_reported, !!!predictor_label_map),
         Group = group_of(predictor_reported),
         #Effect string in raw outcome units: beta [lower, upper] 95% CI
         beta_ci = sprintf('%.2g [%.2g, %.2g]', estimate, estimate - 1.96*SE, estimate + 1.96*SE))

write_tsv(joint_lm_raw %>% select(pheno, Label, predictor_reported, predictor_label, Group, estimate, SE, beta_ci, pval, adj_p),
          file.path(out_dir, 'summary_allpredictor_joint_lm_results_rawpheno.tsv'))
```

## Summary Table

```{r}
joint_lm %>%
  transmute(Endophenotype = Label, Predictor = predictor_label, Group,
            Beta_SD = estimate, SE, P = pval, FDR_P = adj_p) %>%
  arrange(Endophenotype, Group, P) %>%
  mutate(across(where(is.numeric), ~signif(.,3))) %>%
  kable(format='html') %>%
  kable_styling(full_width = F) %>%
  scroll_box(height = '500px')
```

## Forest Plots of All Predictors (by Endophenotype Category)

This plots the mutually-adjusted standardised betas for every reported predictor, coloured by predictor
category, split into three single-row forest plots by endophenotype domain: **hypertrophy** (LV mass index,
maximal wall thickness), **fibrosis** (total LGE, ECV fraction) and **contractility** (LVEF, LV strain).
Significance stars use the FDR-adjusted p-values (`*` < 0.05, `**` < 0.01, `***` < 0.001).

```{r}
#Group each endophenotype into a functional domain
pheno_domains <- list(
  hypertrophy   = c('lvmassi','wallthkmax'),
  fibrosis      = c('lge_total','ecvfwhole'),
  contractility = c('lvef','lvs_ellmean','lvs_eccmean','lvs_errmean')
)

forest_plot_allpredictors <- function(results, output_dir, predictor_order, phenos = NULL,
                                       file_name = 'forest_plot_allpredictor.png', group_palette = 'Set1',
                                       nrow = 1, width = 12, height = 6, dpi = 600){
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

  dat <- results %>%
    { if (is.null(phenos)) . else filter(., pheno %in% phenos) } %>%
    filter(is.finite(estimate), is.finite(SE)) %>%
    mutate(conf_low  = estimate - 1.96*SE,
           conf_high = estimate + 1.96*SE,
           sig_label = case_when(
             adj_p <= 0.001 ~ '***',
             adj_p <= 0.01  ~ '**',
             adj_p <= 0.05  ~ '*',
             TRUE ~ ''),
           predictor_label = factor(predictor_label, levels = rev(predictor_order)),
           Group = factor(Group, levels = c('PRS','Sarcomere','Clinical Risk Factor','Plasma Protein')))

  #Drop the trailing unit in parentheses from facet labels (all phenotypes standardised to SD)
  strip_units <- function(x) str_trim(str_remove(x, '\\s*\\([^()]*\\)\\s*$'))
  dat <- mutate(dat, Label = strip_units(Label))

  #Facet (endophenotype) order follows the supplied `phenos` order, not alphabetical
  ord_phenos   <- if (is.null(phenos)) pheno_labels$pheno else phenos
  label_levels <- strip_units(pheno_labels$Label[match(ord_phenos, pheno_labels$pheno)])
  label_levels <- label_levels[label_levels %in% dat$Label]   #keep only facets actually present
  dat <- mutate(dat, Label = factor(Label, levels = label_levels))

  p <- ggplot(dat, aes(x = estimate, y = predictor_label, fill = Group)) +
    geom_vline(xintercept = 0, linetype = 'dashed', colour = 'grey50') +
    geom_errorbarh(aes(xmin = conf_low, xmax = conf_high), height = 0.2) +
    geom_point(shape = 23, size = 2, colour = 'black', stroke = 0.4) +
    geom_text(aes(label = sig_label), vjust = -0.05, size = 4, show.legend = FALSE) +
    facet_wrap(~ Label, nrow = nrow, scales = 'free_x') +
    scale_fill_brewer(palette = group_palette) +
    labs(x = 'Effect on Endophenotype (SD)', y = NULL, fill = 'Predictor type') +
    theme_classic(base_size = 13) +
    theme(panel.grid.minor = element_blank(),
          strip.background = element_rect(fill = 'grey95'),
          strip.text = element_text(face = 'bold'),
          legend.position = 'bottom')

  ggsave(file.path(output_dir, file_name), p, width = width, height = height, dpi = dpi)
  p
}

forest_out_dir <- 'genetic_subset_analysis/PRS/3_output/1_Endophenotype/2_LM/sarcall/allpredictor/'

#One single-row forest plot per endophenotype domain (facet order = the order in pheno_domains)
forest_plots <- imap(pheno_domains, ~ forest_plot_allpredictors(
  joint_lm,
  output_dir = forest_out_dir,
  predictor_order = predictor_order,
  phenos = .x,
  file_name = str_glue('forest_plot_allpredictor_{.y}.png'),
  height = 5,
  width = 5 + 2 * length(.x)
))

#Overall forest plot: combine the 3 domain plots with patchwork (hypertrophy -> fibrosis -> contractility).
#Drop the redundant y-axis (predictor) labels on the 2nd/3rd panels and collect a single shared legend;
#relative widths track the number of facets per domain.
strip_x <- theme(axis.title.x = element_blank())

forest_plots$overall <- (
  (forest_plots$hypertrophy + strip_x) /
  (forest_plots$fibrosis + strip_x     ) /
  (forest_plots$contractility)
) +
  plot_layout(nrow=3, guides = 'collect',
              widths = map_dbl(pheno_domains, length)) &
  theme(legend.position = 'bottom')

ggsave(file.path(forest_out_dir, 'forest_plot_allpredictor_overall.png'),
       forest_plots$overall, height = 5 + 1 * length(endophenos), width = 10, dpi = 600)
```

```{r, fig.width=9, fig.height=7}
forest_plots$hypertrophy
```

```{r, fig.width=9, fig.height=7}
forest_plots$fibrosis
```

```{r, fig.width=14, fig.height=7}
forest_plots$contractility
```

Overall forest plot across all endophenotypes, ordered hypertrophy -> fibrosis -> contractility.

```{r, fig.width=20, fig.height=7}
forest_plots$overall
```

**Statistical caveats.** 
(1) Each model is complete-case; and the joint N differs across endophenotypes. 
(2) Rare binary CRFs (e.g. prior VT/SCD) may have few positives, giving wide CIs; non-finite estimates are filtered before plotting. 
(3) FDR is applied jointly across all predictor x endophenotype tests. 
(4) Mutually-adjusted betas assume no severe collinearity among predictors. 
(5) Effect sizes are in SD units of each standardised endophenotype; binary CRFs/sarcomere are presence-vs-absence contrasts, continuous predictors/PRS are per-SD.

# Linear Regression Assumptions

This section evaluates the OLS assumptions for each joint endophenotype model. The standardise/log choices do not materially change these checks.

**Multicollinearity** (relevant with this many jointly-modelled predictors) is quantified by VIF below. The
remaining assumptions are assessed graphically in the diagnostic-plot section, in order of importance at this
sample size:
**linearity** of the conditional mean,
**homoscedasticity**,
**influential points/leverage**,

and—least critical, since the CLT covers large-N Wald inference—approximate **normality of residuals**.

## Refit Models (returning the fitted lm objects)

The same joint models are refit with `model_return = TRUE` so the underlying `lm` objects are available for
diagnostics.

```{r}
joint_lm_models <- map(continuous_param_list, ~ linear_logistic_regression_modeller(
  sarcall_scaled, .x,
  predictor = 'cv_prs', predictor_type = 'continuous',
  covars = model_covars, additional_predictors_to_report = extra_predictors,
  model_return = TRUE
))
names(joint_lm_models) <- endophenos
```

## Multicollinearity (VIF)

This reports, per model, the maximum variance inflation factor (VIF) across design-matrix columns and the
predictor carrying it. All RHS terms here are single-df (continuous or 0/1 binary; sex is a 2-level factor),
so ordinary VIF equals the generalised VIF and is the appropriate metric. The remaining OLS assumptions
(linearity, homoscedasticity, influence, residual normality) are assessed graphically in the next section.
```{r}
#Dependency-free VIF on the design-matrix columns (one per dummy for factor terms)
design_vif <- function(model){
  X <- model.matrix(model)
  X <- X[, colnames(X) != '(Intercept)', drop = FALSE]
  if (ncol(X) < 2) return(setNames(rep(NA_real_, ncol(X)), colnames(X)))
  v <- vapply(seq_len(ncol(X)), function(j){
    r2 <- summary(lm(X[, j] ~ X[, -j, drop = FALSE]))$r.squared
    1 / (1 - r2)
  }, numeric(1))
  setNames(v, colnames(X))
}

#Per model: N, max VIF and the design-matrix column carrying it
vif_summary <- function(model){
  vif <- design_vif(model)
  tibble(
    n           = length(residuals(model)),
    max_vif     = max(vif, na.rm = TRUE),
    max_vif_var = names(which.max(vif))
  )
}

lm_vif_tbl <- imap_dfr(joint_lm_models, ~ vif_summary(.x) %>% mutate(pheno = .y)) %>%
  left_join(pheno_labels, by = 'pheno') %>%
  transmute(Endophenotype = Label, N = n, Max_VIF = max_vif, Max_VIF_Predictor = max_vif_var)

write_tsv(lm_vif_tbl, file.path(out_dir, 'lm_vif_summary.tsv'))

lm_vif_tbl %>%
  mutate(across(where(is.numeric), ~ signif(., 3))) %>%
  kable(format = 'html', caption = 'Maximum VIF per joint endophenotype model') %>%
  kable_styling(full_width = F)
```

Interpretation threshold: **Max_VIF > 5** moderate, **> 10** severe collinearity.

## Diagnostic Plots (per model)

The standard four `plot.lm` panels per model: 
- residuals-vs-fitted (linearity), 
- Normal Q-Q (residual normality),
- scale-location (homoscedasticity) and 
- residuals-vs-leverage (influence). 

Saved to disk and rendered inline.
```{r, fig.width=9, fig.height=8}
diag_dir <- file.path(out_dir, 'assumption_plots')
dir.create(diag_dir, recursive = TRUE, showWarnings = FALSE)

iwalk(joint_lm_models, function(model, ph){
  label <- pheno_labels$Label[pheno_labels$pheno == ph]

  #Save the 2x2 diagnostic panel
  grDevices::png(file.path(diag_dir, str_glue('lm_diag_{ph}.png')), width = 9, height = 8, units = 'in', res = 300)
  op <- par(mfrow = c(2, 2)); plot(model, sub.caption = label); par(op)
  grDevices::dev.off()

  #Render inline
  op <- par(mfrow = c(2, 2)); plot(model, sub.caption = label); par(op)
})
```
