---
title: "Preprocessing of HCMR Outcomes "
author: "Jonathan Chan"
date: "2026-01-20"
output: 
  html_document:
    code_folding: hide   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(tictoc)
library(kableExtra)
library(readxl)
library(ggrepel)

rm(list=ls())

theme_set(theme_classic(base_size=14))
```

## Import
```{r import}
#This flag is crucial to determine whether you want to use 2558 OR 2446 individuals (former = not common filtered) - See the Flowchart and hcmr_overall_preprocessor.Rmd

hcmr_n <- 2558 #For rare variant subset analysis (sarcomere-positive focus analysis) Note this later filters to only sarcomere == T (915)
# hcmr_n <- 2446 #For PRS and rare genetic analysis (sarcomere-all)

if(hcmr_n == 2558){
  hcmr_ids <- read_tsv(
  'genetic_subset_analysis/common/output/data/hcmr_agesex_rare_filtered_2558.tsv')
} else if (hcmr_n ==2446){
  hcmr_ids <- read_tsv('genetic_subset_analysis/common/output/data/hcmr_agesex_rare_common_filtered_2446.tsv')
} 
```

```{r}
#This imports in the preprocessed HCMR baseline and image data
total_df_notransform <- readRDS('output/data/total_notransform_nogenofilter_df.rds') %>% 
  select(-starts_with('eos')) %>% #Remove any previous eos variables if they exist
  filter(sitePatID %in% hcmr_ids$sitePatID) #Filter to only the individuals in hcmr_ids

#Convert false positive individuals to sarcomere-negative individuals (due to false positive MYBPC3-VUS variants)
falsepositive_mybpc3_vus_individuals <- readRDS('genetic_subset_analysis/common/falsepositive_mybpc3_vus_individuals.rds')
total_df_notransform <- total_df_notransform %>%
  mutate(sarcomere = ifelse(HCR_IDs %in% falsepositive_mybpc3_vus_individuals, FALSE, sarcomere))

if(hcmr_n == 2558){
  total_df_notransform <- filter(total_df_notransform, sarcomere==T)
}
```

This imports in the final follow-up/censoring/end-of-study data.
```{r}
hcmr_eos <- read_excel('data/HCMR_2026_01/HCMRBaselineDataset_2025-10-14.xlsx', sheet='data') %>%
  select(sitePatID, starts_with('eos')) 

total_df_notransform <- left_join(total_df_notransform, hcmr_eos, by='sitePatID')

print(unique(total_df_notransform$eosyn))

print(str_glue('Number of individuals which did not complete study: {nrow(filter(total_df_notransform, eosyn==0))}'))
print(str_glue('Number of individuals with non-NA date of study completion/discontinuation: {sum(!is.na(total_df_notransform$eosdt))}'))

#Double check that there aren't any individuals who have non-NA eosdt but eosyn==1
print(nrow(filter(total_df_notransform, !is.na(eosdt) & eosyn==1)))

#Checking the 1 individuals with missing eosdt but eosyn==0
missing_eosdt <- filter(total_df_notransform, is.na(eosdt) & eosyn==0) %>%
  select(sitePatID, eosyn, eosn, eosdt)

#Remove the 1 individuals with missing eosdt but eosyn==0 (i.e. didn't complete study but no dropout date)
total_df_notransform <- filter(total_df_notransform, !sitePatID %in% missing_eosdt$sitePatID) 

```


This imports in the outcome data.
```{r}
#This imports in the HCMR outcomes files
hcmr_adj_dict <- read_excel('data/HCMR_2026_01/HCMRAdjDataset_2025-10-14.xlsx',sheet='dictionary') %>% mutate(Type = case_when(str_detect(Type, 'char') ~ 'text',
                    str_detect(Type, 'int') ~ 'numeric',
                    str_detect(Type, 'date') ~ 'text',
                    TRUE ~ 'text')) %>%
  filter(!Name %in% c('OxfordID','GENE','sitePatID2')) #Remove unneeded columns

hcmr_adj <- read_excel('data/HCMR_2026_01/HCMRAdjDataset_2025-10-14.xlsx',sheet='data', col_types = hcmr_adj_dict$Type) %>% #This lists all the events that occurred
  mutate(eventDate = as_date(eventDate))

hcmr_fup_dataset <- read_excel('data/HCMR_2026_01/HCMRFUPDataset_2025-12-05.xlsx',sheet='data Long', range= cell_cols("A:R"))
hcmr_fup_long_dict <- read_excel('data/HCMR_2026_01/HCMRFUPDataset_2025-12-05.xlsx',sheet='dictionary',range= cell_cols("A:F")) %>%
  dplyr::rename(Type=`Field Type`) %>% 
  filter(Variable %in% colnames(hcmr_fup_dataset)) %>%
  mutate(Type = case_when(
                    str_detect(Type, 'date') ~ 'date',
                    str_detect(Type, 'char') ~ 'text',
                    str_detect(Type, 'int') ~ 'numeric',
                    TRUE ~ 'text')) 

lvad_placement_events_part1 <- read_excel('data/HCMR_2026_01/HCMRFUPDataset_2025-12-05.xlsx',sheet='data Long', range= cell_cols("A:E"),col_types = hcmr_fup_long_dict$Type[1:5]) #This is the long-version of the follow-up dataset which lists events by follow-up visit
lvad_placement_events_part2 <- read_excel('data/HCMR_2026_01/HCMRFUPDataset_2025-12-05.xlsx',sheet='data Long', range= cell_cols("G:R"),col_types = hcmr_fup_long_dict$Type[6:17]) #This is the long-version of the follow-up dataset which lists events by follow-up visit
lvad_placement_events <- bind_cols(lvad_placement_events_part1, lvad_placement_events_part2)
rm(hcmr_adj_dict, hcmr_fup_long_dict, lvad_placement_events_part1, lvad_placement_events_part2)

lvad_placement_events <- lvad_placement_events %>%
  select(sitePatID, fulvad, fulvaddt) %>%
  mutate(fulvad = as.numeric(fulvad),
         fulvaddt = as_date(fulvaddt)) %>%
  filter(fulvad==1) #Only filter for those with LVAD placement
```

This extracts the last follow-up date for each patient and appends it to the hcmr_adj tibble.

```{r}
last_fup <- hcmr_fup_dataset %>%
  filter(fudt <= as_date('2025-12-05')) %>%  #Remove all fudt later than 2025-12-05 i.e. the date of the file)
  group_by(sitePatID) %>%
    slice_max(order_by = fudt, n = 1, with_ties = FALSE) %>% #Extracts the last date of follow-up for each patient
    ungroup() %>%
  rename('last_fudt'='fudt') %>%
  mutate(last_fudt = as_date(last_fudt)) %>%
  select(sitePatID,last_fudt)

print(summary(last_fup$last_fudt))

print(str_glue('Number of NA values for last_fudt: {sum(is.na(last_fup$last_fudt))} of {length(unique(last_fup$sitePatID))} patients'))

total_df_notransform <- total_df_notransform %>%
  left_join(last_fup, by='sitePatID')

print(str_glue('Number of NA values in patients for last_fudt: {sum(is.na(total_df_notransform$last_fudt))}'))
print(str_glue('Number of 915 patients with NA last_fudt and eosdt: {nrow(filter(total_df_notransform, is.na(last_fudt) & is.na(eosdt)))}')) #Yeah so only that one individual with a NA last_fudt in the original last_fup. Double checked the Excel and all empty fields for this individual in the follow-up dataset
  
rm(last_fup)
```


### Primary and Secondary Composite Definition

This outlines the outcomes within the primary composite and the secondary composite and for each composite, it creates a tibble ready for survival analysis.
```{r}
#Primary Outcome
primary_comp_events <- hcmr_adj %>%
  left_join(lvad_placement_events, by='sitePatID') %>%
  filter(criteriamet ==1) %>% #Only filter for those which meet the adjudicated criteria
  select(-criteriamet) %>%
  filter(death_cause == 1 | # Cardiac death
         vtach == 1 | #Ventricular tachyarrhythmia event
         transplant ==1 | #Heart transplant
        fulvad == 1 #LVAD placement
         ) 

#Create a tibble with the following columns: 1) sitePatID 2) event and 3) event_date
primary_comp_events_surv <- primary_comp_events |>
  transmute(
    sitePatID,
    event = case_when(
      death_cause == 1 ~ "cardiac_death",
      vtach == 1 ~ "vtach",
      transplant == 1 ~ "transplant",
      fulvad == 1 ~ "lvad",
      TRUE ~ NA_character_
    ),
    event_date = case_when(
      death_cause == 1 ~ eventDate,
      vtach == 1 ~ eventDate,
      transplant == 1 ~ eventDate,
      fulvad == 1 ~ fulvaddt,
      TRUE ~ as_date(NA_character_)
    )
  ) |>
  filter(!is.na(event), !is.na(event_date))
```


```{r}
#Secondary outcomes
secondary_comp_events <- hcmr_adj %>%
  left_join(lvad_placement_events, by='sitePatID') %>%
  filter(criteriamet ==1) %>% #Only filter for those which meet the adjudicated criteria
  select(-criteriamet) %>%
  filter( #Primary Outcomes
        death_cause == 1 | # Cardiac death
         vtach == 1 | #Ventricular tachyarrhythmia event
         transplant ==1 | #Heart transplant
        fulvad == 1 | #LVAD placement
        #Secondary outcomes
        death_cause ==2 | #Non-cardiac death
        hf==1 | #Heart failure hospitlisation
        afib ==1 |  #Atrial fibrillation
        (death==0 & stroke==1) #Non-fatal stroke
         )
#Create a tibble with the following columns: 1) sitePatID 2) event and 3) event_date
secondary_comp_events_surv <- secondary_comp_events %>%
  transmute(
    sitePatID,
    event = case_when(
      death_cause == 1 ~ "cardiac_death",
      vtach == 1 ~ "vtach",
      transplant == 1 ~ "transplant",
      fulvad == 1 ~ "lvad",
      death_cause == 2 ~ "non_cardiac_death",
      hf == 1 ~ "heart_failure_hospitalisation",
      afib == 1 ~ "atrial_fibrillation",
      (death == 0 & stroke == 1) ~ "non_fatal_stroke",
      TRUE ~ NA_character_
    ),
    event_date = case_when(
      death_cause == 1 ~ eventDate,
      vtach == 1 ~ eventDate,
      transplant == 1 ~ eventDate,
      fulvad == 1 ~ fulvaddt,
      death_cause == 2 ~ eventDate,
      hf == 1 ~ eventDate,
      afib == 1 ~ eventDate,
      (death == 0 & stroke == 1) ~ eventDate,
      TRUE ~ as_date(NA_character_)
    )
  ) %>%
  filter(!is.na(event), !is.na(event_date))

#Check which rows are in the hcmr_adj but not in the secondary_comp_events_surv
missing_secondary_events <- anti_join(hcmr_adj, secondary_comp_events_surv, by = "sitePatID") #Either death_cause unknown or no)
rm(secondary_comp_events, primary_comp_events)

print(sum(is.na(primary_comp_events_surv$event_date)))
print(sum(is.na(secondary_comp_events_surv$event_date)))
```

### Custom Composite Definition 

This also creates tibbles for custom composites (beyond the primary/secondary defined by HCMR).

```{r}
#HF-related composite
hf_comp_events <- hcmr_adj %>%
  left_join(lvad_placement_events, by='sitePatID') %>%
  filter(criteriamet ==1) %>% #Only filter for those which meet the adjudicated criteria
  select(-criteriamet) %>%
  filter(
        (death_cause == 1 & deathCardio == 3)| # Cardiac death (HF-related specifically)
        transplant ==1 | #Heart transplant
        fulvad == 1 | #LVAD placement
        hf==1 #Heart failure hospitlisation
         )

hf_comp_events_surv <- hf_comp_events %>%
  transmute(
    sitePatID,
    event = case_when(
      death_cause == 1 & deathCardio == 3 ~ "hf_death",
      transplant == 1 ~ "transplant",
      fulvad == 1 ~ "lvad",
      hf == 1 ~ "hf_hospitalisation",
      TRUE ~ NA_character_
    ),
    event_date = case_when(
      death_cause == 1 & deathCardio == 3 ~ eventDate,
      transplant == 1 ~ eventDate,
      fulvad == 1 ~ fulvaddt,
      hf == 1 ~ eventDate,
      TRUE ~ as_date(NA_character_)
    )
  ) %>%
  filter(!is.na(event), !is.na(event_date))

rm(hf_comp_events)
```

```{r}
#VA-related composite
va_comp_events <- hcmr_adj %>%
  filter(criteriamet ==1) %>% #Only filter for those which meet the adjudicated criteria
  select(-criteriamet) %>%
  filter(
        (death_cause == 1 & deathCardio == 1)| # Cardiac death (sudden arrhythmic)
        (death_cause == 1 & deathCardio == 2)| # Cardiac death (non-sudden arrhythmic)
        vtach ==1 #Ventricular tachyarrhythmia event e.g. aborted SCD or VT
         )

va_comp_events_surv <- va_comp_events %>%
  transmute(
    sitePatID,
    event = case_when(
      death_cause == 1 & deathCardio == 1 ~ "va_death",
      death_cause == 1 & deathCardio == 2 ~ "va_death",
      vtach == 1 ~ "vtach",
      TRUE ~ NA_character_
    ),
    event_date = case_when(
      death_cause == 1 & deathCardio == 1 ~ eventDate,
      death_cause == 1 & deathCardio == 2 ~ eventDate,
      vtach == 1 ~ eventDate,
      TRUE ~ as_date(NA_character_)
    )
  ) %>%
  filter(!is.na(event), !is.na(event_date))

rm(va_comp_events)
```

```{r}
#AF events
af_events <- hcmr_adj %>%
  filter(criteriamet ==1) %>% #Only filter for those which meet the adjudicated criteria
  select(-criteriamet) %>%
  filter(
        afib ==1 #Atrial fibrillation
         )

af_events_surv <- af_events %>%
  transmute(
    sitePatID,
    event = case_when(
      afib == 1 ~ "atrial_fibrillation",
      TRUE ~ NA_character_
    ),
    event_date = case_when(
      afib == 1 ~ eventDate,
      TRUE ~ as_date(NA_character_)
    )
  ) %>%
  filter(!is.na(event), !is.na(event_date))

rm(af_events)
```

```{r}
#Stroke-related composite
stroke_comp_events <- hcmr_adj %>%
  filter(criteriamet ==1) %>% #Only filter for those which meet the adjudicated criteria
  select(-criteriamet) %>%
  filter(
        (death_cause == 1 & deathCardio == 8)| # Cardiac death (stroke-related)
        (death == 0 & stroke == 1) #Non-fatal stroke
         )

stroke_comp_events_surv <- stroke_comp_events %>%
  transmute(
    sitePatID,
    event = case_when(
      death_cause == 1 & deathCardio == 8 ~ "fatal_stroke",
      (death == 0 & stroke == 1) ~ "nonfatal_stroke",
      TRUE ~ NA_character_
    ),
    event_date = case_when(
      death_cause == 1 & deathCardio == 8 ~ eventDate,
      (death == 0 & stroke == 1) ~ eventDate,
      TRUE ~ as_date(NA_character_)
    )
  ) %>%
  filter(!is.na(event), !is.na(event_date))

rm(stroke_comp_events)
```

### Censoring of Individuals with Unknown Death Cause

This includes the individuals who died of unknown causes as censored individuals.

```{r}
unknown_death <- filter(hcmr_adj, death_cause==3) %>%
  select(sitePatID, eventDate) %>%
  mutate(event = 'unknown_death',
         event_date = eventDate) %>%
  select(-eventDate)
```

## Final Dataset Construction

This then merges the total_df_notransform with the primary and secondary outcome tibbles to create two final datasets for survival analysis and then for each computes 

1) indicator for the event or censoring 
2) event/censoring date 
3) time to event/censoring in days and 
4) age at event/censoring in years

```{r}
make_outcome_df <- function(total_df_notransform, comp_events_surv, unknown_death, output_basename) {

  output_df <- total_df_notransform %>%
    left_join(comp_events_surv, by = "sitePatID") %>%
    left_join(unknown_death, by = "sitePatID", suffix = c("", "_unknowndeath")) %>%
    mutate(
      event_indicator = ifelse(!is.na(event), 1, 0), # event refers to the composite events in comp_events_surv
      event_or_censoring_date = case_when( #This shows the ranking for how this date is computed with priority of event_date > unknown_death_date > eosdt > last_fudt > icfDate
        event_indicator == 1 ~ event_date,
        event_unknowndeath == "unknown_death" ~ event_date_unknowndeath, # unknown death treated as censored
        eosyn == 0 & !is.na(eosdt) ~ eosdt, # discontinued from study
        !is.na(last_fudt) ~ last_fudt, # last follow-up date for the patient as per the FUP dataset long version
        T ~ icfDate      #If no event and no eosdt and no fudt, then assume the patient was censored at icfdate
        ),
      time_to_event_or_censoring_years =
        as.numeric(difftime(event_or_censoring_date, icfDate, units = "days")) / 365.25,
      age_at_event_or_censoring = age + time_to_event_or_censoring_years
    ) %>%
    # Ensure one row per person by taking the earliest event/censoring date
    group_by(sitePatID) %>%
    slice_min(order_by = event_or_censoring_date, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    select(-event_unknowndeath, -event_date_unknowndeath)
  
  #Output error if there is any individual with NA event_or_censoring_date
  if(nrow(filter(output_df, is.na(event_or_censoring_date))) >0){
    print('NA value in the event_or_censoring date for some individual in this dataset')
  }

  composite <- str_match(output_basename, '_([A-Za-z]+_outcome_df)')[,2]
  print(str_glue("Number of individuals with {composite} composite events: {sum(output_df$event_indicator)}"))

    #This saves the primary and secondary outcome datasets for survival analysis as .rds files
  saveRDS(output_df, str_glue('{output_basename}_{nrow(output_df)}.rds'))

  return(output_df)
}
```

```{r}
#Make the outcome_dfs
composites <- c('primary','secondary','hf','va','af','stroke')

outcome_dfs <- map2(composites, list(primary_comp_events_surv, secondary_comp_events_surv, hf_comp_events_surv, va_comp_events_surv, af_events_surv, stroke_comp_events_surv), ~make_outcome_df(
  total_df_notransform, 
  comp_events_surv = .y,
  unknown_death = unknown_death,
  output_basename = str_glue('genetic_subset_analysis/common/output/data/hcmr_{.x}_outcome_df')
))

names(outcome_dfs) <- composites

rm(list=ls(pattern='surv'))
```

```{r}
#Follow-up times are computed as the time from baseline to the event date or the censoring date (whichever comes first) and then converted to years. The median, Q1, Q3, mean and SD of follow-up times are then printed.
#I repeat this for both primary and secondary composites.

fup_times <- map(outcome_dfs[1:2], ~mutate(.x, fup_time_years = as.numeric(difftime(event_or_censoring_date, icfDate, units = "days")) / 365.25))  #Primary and secondary composites only
fup_quartiles <- map(fup_times, ~quantile(.x$fup_time_years, probs = c(0.25, 0.5, 0.75), na.rm = TRUE))

print(fup_quartiles)

mean_fup_times <- map(fup_times, ~mean(.x$fup_time_years))
sd_fup_times <- map(fup_times, ~sd(.x$fup_time_years))

walk2(mean_fup_times,sd_fup_times, ~print(str_glue('Follow-up time (years) mean and SD: {.x} ± {.y}')))

rm(fup_times, fup_quartiles)
```

```{r}
kable_prep <- function(input_df){

  output_df <- input_df %>%
    group_by(event) %>%
    summarise(count = n()) %>%
    arrange(desc(count))

  return(output_df)
}
```

```{r}
#Output a HTML table of the breakdown of composite events 

kable(kable_prep(outcome_dfs[[1]]), caption = "Breakdown of Primary Composite Events") %>%
  kable_styling(full_width = F)
```

```{r}
kable(kable_prep(outcome_dfs[[2]]), caption = "Breakdown of Secondary Composite Events") %>%
  kable_styling(full_width = F)
```

```{r}
kable(kable_prep(outcome_dfs[[3]]), caption = "Breakdown of HF Composite Events") %>%
  kable_styling(full_width = F)
```

```{r}
kable(kable_prep(outcome_dfs[[4]]), caption = "Breakdown of VA Composite Events") %>%
  kable_styling(full_width = F)
```

```{r}
kable(kable_prep(outcome_dfs[[5]]), caption = "Breakdown of AF Events") %>%
  kable_styling(full_width = F)
```

```{r}
kable(kable_prep(outcome_dfs[[6]]), caption = "Breakdown of Stroke Composite Events") %>%
  kable_styling(full_width = F)
```