---
title: "Comparison of HCMR Case Rare-Variant Carriers on Endophenotype - Powell Classification"
author: "Jonathan Chan"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(factoextra)
library(tictoc)
library(readxl)

rm(list=ls())

theme_set(theme_classic(base_size=14))

source('genetic_subset_analysis/common/hcmr_endophenotype_functions.R')
```

This reruns HCMR_RareVar_Subset_Comparison_Endophenotype_Analysis.Rmd as a sensitivity analysis but using George Powell's variant classifications (output of RareVar_Classification/HCMR_RareVar_Classification_Powell.Rmd) instead of the KT classifications.

Differences vs. the main analysis:

- Sarcomere-positive = carries >=1 Powell SARC+ variant (rather than KT sarcomere status with the MYBPC3-VUS false-positive correction)
- Powell's classification has no pathogenicity tier, so the P/LP vs. VUS comparison is not performed (3 comparisons only)
- All outputs are written with a _powell suffix or to a Powell/ subdirectory so the main analysis outputs are not overwritten

## Import

```{r import, echo=F}
#Import in the total_df_notransform from EDA_HCMR.Rmd
hcmr_ids_of_interest <- read_tsv('genetic_subset_analysis/common/output/data/hcmr_agesex_rare_filtered_2558.tsv')
hcmr_tb <- readRDS('output/data/total_notransform_nogenofilter_df.rds') %>% #I use the 2750 -> filtered for those who have HCR IDs (2677) (i.e. without filtering for hard-called genotypes)
  distinct()

hcmr_tb <- hcmr_tb %>%
  filter(HCR_IDs %in% hcmr_ids_of_interest$HCR_IDs) #Now filter to those individuals of interest

pergene_rarevar_classification <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_rarevar_class_powell.tsv')
hcmr_tb <- hcmr_tb %>%
  mutate(sarcomere_kt = sarcomere) %>% #Retain the original (KT) sarcomere status for the classification comparison below
  mutate(sarcomere = HCR_IDs %in% pergene_rarevar_classification$HCR_ID) #Sarcomere-positive redefined as carrying >=1 Powell SARC+ variant

rm(hcmr_ids_of_interest)

output_data_dir <- 'genetic_subset_analysis/RareVar/3_output/data/'
output_plot_dir <- 'genetic_subset_analysis/RareVar/3_output/plots/1_Endophenotype/Powell/'
walk(str_c(output_plot_dir, c('0_Classification_Comparison/','1_LM/','2_Wilcoxon/')), ~dir.create(., showWarnings=F, recursive=T))
```

## Variant Classification Comparison: KT vs. Powell

This compares sarcomere-positive status under the original (KT) variant classification (sarcomere==TRUE in hcmr_tb) against the Powell classification (carrying >=1 Powell SARC+ variant), within the analysed HCMR cohort.

N.B. KT status here is the raw sarcomere column, i.e. without the MYBPC3-VUS false-positive correction used in the main analysis.

```{r classification_comparison}
class_comparison_tb <- hcmr_tb %>%
  select(HCR_IDs, sarcomere_kt, sarcomere_powell = sarcomere) %>%
  distinct() %>%
  left_join(select(pergene_rarevar_classification, HCR_ID, powell_class = class), by=c('HCR_IDs'='HCR_ID')) %>%
  mutate(concordance = case_when(
    is.na(sarcomere_kt) ~ 'KT status missing',
    sarcomere_kt & sarcomere_powell ~ 'SARC+ in both',
    !sarcomere_kt & !sarcomere_powell ~ 'SARC- in both',
    sarcomere_kt & !sarcomere_powell ~ 'SARC+ in KT only',
    !sarcomere_kt & sarcomere_powell ~ 'SARC+ in Powell only'
  ))

print(str_glue('{sum(class_comparison_tb$sarcomere_kt, na.rm=T)} sarcomere-positive individuals as per KT classification (of {nrow(class_comparison_tb)}; {sum(is.na(class_comparison_tb$sarcomere_kt))} with missing KT status)'))
print(str_glue('{sum(class_comparison_tb$sarcomere_powell)} sarcomere-positive individuals as per Powell classification (of {nrow(class_comparison_tb)})'))
print(str_glue('{sum(!pergene_rarevar_classification$HCR_ID %in% class_comparison_tb$HCR_IDs)} Powell SARC+ individuals are not in the analysed HCMR cohort'))

#Concordance counts
class_comparison_tb %>%
  count(concordance) %>%
  mutate(pct = round(100 * n / sum(n), 1))

#2x2 contingency table (KT rows x Powell columns), excluding missing KT status
class_comparison_complete_tb <- class_comparison_tb %>%
  filter(!is.na(sarcomere_kt)) %>%
  mutate(across(c(sarcomere_kt, sarcomere_powell), ~factor(if_else(., 'SARC+', 'SARC-'), levels=c('SARC+','SARC-'))))

class_contingency <- table(KT = class_comparison_complete_tb$sarcomere_kt, Powell = class_comparison_complete_tb$sarcomere_powell)
addmargins(class_contingency)

#Agreement statistics: % agreement, Cohen's kappa and McNemar's test for asymmetry in discordant pairs
po <- sum(diag(class_contingency)) / sum(class_contingency)
pe <- sum(rowSums(class_contingency) * colSums(class_contingency)) / sum(class_contingency)^2
kappa <- (po - pe) / (1 - pe)
print(str_glue('Percentage agreement: {round(100 * po, 1)}%; Cohen\'s kappa: {round(kappa, 3)}'))
mcnemar.test(class_contingency)

write_tsv(class_comparison_tb, str_c(output_data_dir, 'hcmr_sarcomere_kt_vs_powell_comparison.tsv'))
```

This plots the 2x2 contingency table as a heatmap.

```{r classification_comparison_plot, fig.width=6, fig.height=5}
class_contingency_plot <- as_tibble(class_contingency) %>%
  mutate(pct = 100 * n / sum(n)) %>%
  ggplot(aes(x = Powell, y = KT, fill = n)) +
  geom_tile(colour = 'white') +
  geom_text(aes(label = str_c(n, '\n(', round(pct, 1), '%)')), size = 5, colour = 'white') +
  scale_fill_viridis(option = 'D', end = 0.85) +
  scale_y_discrete(limits = rev) +
  labs(x = 'Powell Classification', y = 'KT Classification', fill = 'N',
       title = 'Sarcomere Status: KT vs. Powell',
       subtitle = str_glue('Agreement = {round(100 * po, 1)}%; Cohen\'s kappa = {round(kappa, 3)}'))

print(class_contingency_plot)
ggsave(str_c(output_plot_dir, '0_Classification_Comparison/sarcomere_kt_vs_powell_contingency.png'), class_contingency_plot, width=6, height=5, dpi=300)
```

This breaks down the Powell per-gene class for Powell sarcomere-positive individuals by concordance with KT, to show which genes/variant types drive the discordance.

```{r classification_comparison_breakdown}
class_comparison_tb %>%
  filter(sarcomere_powell) %>%
  count(concordance, powell_class) %>%
  pivot_wider(names_from = concordance, values_from = n, values_fill = 0) %>%
  arrange(desc(rowSums(across(where(is.numeric)))))

rm(class_comparison_complete_tb, class_contingency, class_contingency_plot, po, pe, kappa)
```

## Phenotypes for Comparison

The following phenotypes will be analysed each reflecting some component of HCM disease.

- Hypertrophy: wallthkmax; lvmassi
- Fibrosis: lge_total; ecvfwhole
- Hypercontractility: LVEF (from CMR) and LV strain parameters (lvs_ellmean; eccmean; errmean)

Note that lge_total is treated both as a continuous and a discrete trait via binning.

```{r phenotype_tb}
phenos_of_interest <- c('lge_total','wallthkmax','lvmassi','lge_total','ecvfwhole','lvef','lvs_ellmean','lvs_eccmean','lvs_errmean')
contphenos <- phenos_of_interest

continuous_param_list <- list(
  c('wallthkmax','Maximal LV Wall Thickness (mm)'),
  c('lvmassi','LV Mass Index (g/m2)'),
  c('lge_total','LGE Mass over All Segments / LV Mass (%)'),
  c('ecvfwhole','Whole LV Extracellular Volume Fraction (%)'),
  c('lvef','LV Ejection Fraction (%)'),
  c('lvs_ellmean','-Longitudinal LV Strain'),
  c('lvs_eccmean','-Circumferential LV Strain'),
  c('lvs_errmean','Radial LV Strain')
)
names(continuous_param_list) <- map_chr(continuous_param_list, ~.[[1]])

label_tb <- read_tsv('output/data/combo_labels_20260123.tsv')
```

```{r sarcpos_filter}
print(str_glue('Filtering out {nrow(hcmr_tb) - nrow(hcmr_tb %>% filter(sarcomere==T))} Powell sarcomere-negative individuals from HCMR cohort for analysis.'))

hcmr_sarcpos_tb <- hcmr_tb %>%
  filter(sarcomere==T) #Filter to only Powell sarcomere-positive individuals

print(str_glue('There are {length(unique(hcmr_sarcpos_tb$HCR_IDs))} sarcomere-positive individuals as per Powell classification'))
```

### Phenotype Tibble Builder

This builds a single overall phenotype tibble with the primary rare variant classification (per-gene) from Powell's SARC+ variants.

```{r}
#Build the phenotype tibble with the Powell per-gene classification
hcmr_sarcpos_tb <- hcmr_sarcpos_tb %>%
  left_join(select(pergene_rarevar_classification,HCR_ID, class_primary=class), by=c('HCR_IDs'='HCR_ID'))

#Some phenotype-specific preprocessing
hcmr_sarcpos_tb <- hcmr_sarcpos_tb %>%
    mutate(lge_total_bin = case_when(lge_total==0 ~ '0%',
                                     lge_total >0 & lge_total <=5 ~ '0-5%',
                                     lge_total > 5 & lge_total <=10 ~ '5-10%',
                                     lge_total > 10 & lge_total <= 15 ~ '10-15%',
                                     lge_total > 15 ~ '>15%',
                                     T~NA)) %>%
    mutate(lge_total_bin = factor(lge_total_bin, levels=c('0%','0-5%','5-10%','10-15%','>15%'), ordered=T)) %>%
    mutate(ecvfwhole=ecvfwhole * 100)

#Output the hcmr_sarcpos_tb with primary class attached and with no transformation and without filtering for the NA values in the endophenotypes of interest
hcmr_sarcpos_tb_noNAfiltered <- hcmr_sarcpos_tb
write_tsv(hcmr_sarcpos_tb_noNAfiltered, str_c(output_data_dir, 'hcmr_sarcpos_pheno_rarevarclasses_noNAfiltered_notransform_powell.tsv'))
```

```{r}
covs <- c('bmi','sarcomere','FieldStrength')

#Build the initial tb for all Powell sarc_positive individuals
hcmr_sarcpos_tb <- hcmr_sarcpos_tb %>%
  select(HCR_IDs,age,sex=gender,bsa, all_of(covs), class_primary,
         all_of(phenos_of_interest))

hcmr_sarcpos_tb_noNAfiltered <- hcmr_sarcpos_tb_noNAfiltered %>%
  select(HCR_IDs,age,sex=gender,bsa, all_of(covs), class_primary,
         all_of(phenos_of_interest))

#Only apply this filtering step to the hcmr_sarcpos_tb
hcmr_sarcpos_tb <- hcmr_sarcpos_tb %>%
  filter(if_all(c(covs), negate(is.na))) %>%
  filter(if_any(c(contphenos), negate(is.na))) #Filter out all individuals who have NA values for all continuous phenotypes or covariates

print(str_glue('{nrow(hcmr_sarcpos_tb_noNAfiltered)} Powell sarcomere-positive individuals; {nrow(hcmr_sarcpos_tb)} after filtering for non-NA covariates/phenotypes.'))
```

This creates variations of the phenotype tibble for downstream analyses.

```{r}
hcmr_sarcpos_tb_notransform <- hcmr_sarcpos_tb #Non-transformed version of the tibble (but has been filtered for non-NA individuals)

hcmr_sarcpos_tb_notransform_lgelogged <- hcmr_sarcpos_tb_notransform %>%
  mutate(lge_total = log1p(lge_total)) #Log-transform LGE to correct for skewness (add a small constant to avoid log(0))

#Log-transform the skewed LGE and also scale into SD
scaled_cols <- c('lge_total')
hcmr_sarcpos_tb <- hcmr_sarcpos_tb %>%
  mutate(across(scaled_cols, ~log1p(.))) %>% #Log-transform LGE to correct for skewness (add a small constant to avoid log(0))
  mutate(across(contphenos, ~scale(.)[,1])) # Standard scaling of all endophenotypes

write_tsv(hcmr_sarcpos_tb_notransform, str_c(output_data_dir, 'hcmr_sarcpos_pheno_rarevarclasses_notransform_powell.tsv'))
write_tsv(hcmr_sarcpos_tb_notransform_lgelogged, str_c(output_data_dir, 'hcmr_sarcpos_pheno_rarevarclasses_notransform_lgelogged_powell.tsv'))
write_tsv(hcmr_sarcpos_tb, str_c(output_data_dir, 'hcmr_sarcpos_pheno_rarevarclasses_transform_powell.tsv'))
rm(hcmr_tb)
```

### Comparison Group Generator

These generate the comparison_tbs for each comparison - either transformed or non-transformed and all for single-dosage individuals.

N.B. comparison_pheno_tb_gen()'s single-dosage filter assumes the KT gene bins (MYBPC3_pLOF:FHOD3), so here it is called with only_singles=F and the single-dosage filter is applied using the Powell gene bins instead.

```{r}
#Single-dosage individuals = total dosage of 1 across all Powell gene bins
powell_single_dosage_ids <- pergene_rarevar_classification %>%
  mutate(total_dosage = rowSums(select(., -HCR_ID, -class), na.rm=T)) %>%
  filter(total_dosage == 1) %>%
  pull(HCR_ID)

powell_comparison_pheno_tb_gen <- function(phenotype_tb, param_list){
  comparison_pheno_tb_gen(phenotype_tb = phenotype_tb, param_list = param_list, rarevar_class_tb = pergene_rarevar_classification, var_class='primary', only_singles=F) %>%
    filter(HCR_IDs %in% powell_single_dosage_ids)
}
```

```{r}
#This code generates the comparison lists
#For thick vs. thin (excluding ACTC1 from Thin btw)
thick_class <- unique(hcmr_sarcpos_tb$class_primary)[!str_detect(unique(hcmr_sarcpos_tb$class_primary),'ACTC1|TPM1|TNNI3|TNNT2') & str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYBPC3|MYH7|MYL2|MYL3')]
thin_class <- unique(hcmr_sarcpos_tb$class_primary)[str_detect(unique(hcmr_sarcpos_tb$class_primary),'TPM1|TNNI3|TNNT2') & !str_detect(unique(hcmr_sarcpos_tb$class_primary),'ACTC1|MYBPC3|MYH7|MYL2|MYL3')]

#For mybpc3_pLOF vs. mybpc3_nonlof
mybpc3_plof_class <- unique(hcmr_sarcpos_tb$class_primary)[str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYBPC3_pLOF') & !str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYBPC3_nonLOF')]
mybpc3_nonlof_class <- unique(hcmr_sarcpos_tb$class_primary)[str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYBPC3_nonLOF') & !str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYBPC3_pLOF')]

#For mybpc3 vs. MYH7
mybpc3_class <- unique(hcmr_sarcpos_tb$class_primary)[str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYBPC3') & !str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYH7')]
myh7_class <- unique(hcmr_sarcpos_tb$class_primary)[str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYH7') & !str_detect(unique(hcmr_sarcpos_tb$class_primary),'MYBPC3')]

#---------------------------------------------------------------------------------------
#Parameter lists for each comparison
param_lists <- list(
  thick_thin_tb = list(thick_class, 'Thick',thin_class, 'Thin'),
  mybpc3_plof_nonlof_tb = list(mybpc3_plof_class, 'MYBPC3_pLOF', mybpc3_nonlof_class, 'MYBPC3_nonLOF'),
  mybc3_myh7_tb = list(mybpc3_class, 'MYBPC3',myh7_class,'MYH7')
)

rm(list=ls(pattern='class$'))
```

```{r}
#These are with the transformed phenotype_tb
comparison_tbs <- map(param_lists, ~powell_comparison_pheno_tb_gen(hcmr_sarcpos_tb, .))
write_rds(comparison_tbs, str_c(output_data_dir, 'hcmr_comparison_tbs_powell.rds'))

#These are with the nontransformed phenotype_tb (but without logged LGE)
comparison_tbs_notransform <- map(param_lists, ~powell_comparison_pheno_tb_gen(hcmr_sarcpos_tb_notransform, .))
write_rds(comparison_tbs_notransform, str_c(output_data_dir, 'hcmr_comparison_tbs_notransform_powell.rds'))

#These are with the nontransformed phenotype_tb without filtering out the individuals with NA values for the continuous endophenotypes (but without LOG of LGE)
comparison_tbs_noNAfiltered <- map(param_lists, ~powell_comparison_pheno_tb_gen(hcmr_sarcpos_tb_noNAfiltered, .))
write_rds(comparison_tbs_noNAfiltered, str_c(output_data_dir, 'hcmr_comparison_tbs_noNAfiltered_powell.rds'))

#Group counts per comparison
map(comparison_tbs, ~count(., group))

rm(param_lists)
```

## Linear/Logistic Regression Modelling

I also fit linear/logistic regression models for the continuous/discrete traits to correct for confounding effects such as age, sex, BMI and BSA (where necessary e.g. fro maxLVWT).

```{r}
comparison_groups <- list(c('Thick','Thin'), c('MYBPC3_pLOF','MYBPC3_nonLOF'), c('MYBPC3','MYH7'))
comporder <- c('Thick vs. Thin','MYBPC3 vs. MYH7','MYBPC3_pLOF vs. MYBPC3_nonLOF')
lm_covars <- c('age','sex','bmi', 'FieldStrength')

#Runs the linear regression for every comparison x continuous phenotype and builds the summary tibble
lm_summary_builder <- function(comparison_tbs_list){
  map2(comparison_tbs_list, comparison_groups, function(tb, grps){
    map(continuous_param_list, ~linear_logistic_regression_modeller(tb, ., grps[1], grps[2], covars=lm_covars)) %>% bind_rows()
  }) %>%
    bind_rows() %>%
    mutate(comparison = str_c(group1, ' vs. ', group2)) %>%
    left_join(select(label_tb,Variable, Label),by=c('pheno'='Variable')) %>%
    mutate(estimate = - estimate) %>% #Invert the estimate to reflect group1 vs. group2
    select(-term) %>%
    mutate(Label_short = case_when(
      pheno == 'wallthkmax' ~ 'maxLVWT',
      pheno == 'lvmassi' ~ 'LVMi',
      pheno == 'lge_total' ~ 'LGE',
      pheno == 'ecvfwhole' ~ 'ECVF',
      pheno == 'lvef' ~ 'LVEF',
      pheno == 'lvs_ellmean' ~ '-LongStrain',
      pheno == 'lvs_eccmean' ~ '-CircStrain',
      pheno == 'lvs_errmean' ~ 'RadStrain'
    )) %>%
    mutate(Label_short = str_c(Label_short, ' (', comparison,')'))
}
```

```{r linlogreg}
cont_lm_summary_tb <- lm_summary_builder(comparison_tbs)

#Makes the label reflect the normalisation
cont_lm_summary_tb <-  cont_lm_summary_tb %>%
  mutate(Label= case_when(
      pheno %in% scaled_cols ~ str_replace(Label, '\\([^ ]+\\)$', '(SD(Log(%)))'),
      pheno %in% phenos_of_interest ~ str_replace(Label, '\\([^ ]+\\)$', '(SD)')))

cont_lm_summary_tb <- colour_hardcoder(cont_lm_summary_tb)
```

This plots out the summary plot from the continuous phenotypes via linear regression - so adjusted for age/sex and BSA if necessary.

```{r lm_summary, fig.width=9, fig.height=6}
summary_plotter(cont_lm_summary_tb,output_path=str_c(output_plot_dir, '1_LM/'), mtc='fdr',confint=T, label=T, test='LinearRegression',xlabel='Coefficient Estimate', height=7, width=9, shape_values=c(15,16,17), comporder=comporder)

write_tsv(cont_lm_summary_tb, str_c(output_plot_dir, '1_LM/cont_summary_tb.tsv'))

#This is for publication
summary_plotter(cont_lm_summary_tb,output_path=str_c(output_plot_dir, '1_LM/'), mtc='fdr',confint=T, poster=T, test='LinearRegression',xlabel='Coefficient Estimate', height=6, width=11, shape_values=c(15,16,17), comporder=comporder)
```

This outputs the summary plot as facetted forest plots instead with significance stars for multiple testing comparison.

```{r}
cont_lm_summary_tb_forest <- read_tsv(str_c(output_plot_dir, '1_LM/cont_summary_tb.tsv')) %>%
  mutate(Label_short = str_match(Label_short, '(^[^()]+)')[,2]) %>%
  select(-Label) %>%
  dplyr::rename(Label = Label_short, predictor_label = comparison) %>%
  mutate(adj_p = p.adjust(pval, method='BH'))  %>%
  mutate(predictor_reported = predictor_label)

summary_forest_plotter(cont_lm_summary_tb_forest, str_c(output_plot_dir, '1_LM/'), file_name = 'LinearRegression_summary_forestplot_fdr.png', sigphenos_only=F,facet_order=c('Thick vs. Thin','MYBPC3 vs. MYH7','MYBPC3_pLOF vs. MYBPC3_nonLOF'), height=9, width=6)
```

This outputs an alternative to the single combined scatter (volcano) summary plot: a 3-panel plot with one panel per genotypic comparison, sharing the same x-axis (standardised coefficient estimate) and y-axis (-log10 FDR-adjusted p-value) across all panels. Points are coloured by phenotype, filled if FDR-significant, with direction-of-effect annotated per panel.

```{r lm_summary_panelplot, fig.width=10, fig.height=8}
#summary_panel_plotter() is defined in genetic_subset_analysis/common/hcmr_endophenotype_functions.R
summary_panel_plotter(read_tsv(str_c(output_plot_dir, '1_LM/cont_summary_tb.tsv')), output_path = str_c(output_plot_dir, '1_LM/'), comporder=comporder)
```

This also outputs the summary tibble but after running the linear regression without transformation (to enable extraction of coefficient values in native measurement units)

```{r}
cont_lm_summary_tb_notransform <- lm_summary_builder(comparison_tbs_notransform) %>%
  colour_hardcoder()

write_tsv(cont_lm_summary_tb_notransform, str_c(output_plot_dir, '1_LM/cont_summary_tb_notransform.tsv'))
```

I also need to check the assumptions of linear modelling to see which are valid models. These include

- Linear relationship between predictor and outcome
- Normally distributed residuals
- Homoscedasticity (residuals have constant variance)
- Independence of residuals

From the diagnostic plots,  for good assumptions met, check for:

1) Plot 1 (residuals vs. fitted) should be a horizontal line without patterns -> linearity relationship assumption OK
2) Plot 2 (QQ) should have dots following the straight, dashed line -> normality of residuals assumption OK
3) Plot 3 (Scale-Location) should have a horizontal line with equally spread points -> homoscedasticity
4) Plot 4 (Residuals vs. Leverage) IDs influential points which might influence the regression fit
  - Generally you would look for outliers with a magnitude of >3 (+3 or -3) standardised residuals (i.e outliers exceeding 3 standard deviations) and high-leverage points.
  - A good high-leverage threshold is 2(p+1)/n where p is the number of predictors and n is the number of observations

```{r lm_assumption_checker}
comparison_names <- c('Thick vs. Thin','MYBPC3_pLOF vs. MYBPC3_nonLOF', 'MYBPC3 vs. MYH7')
phenotype_names <- map_chr(continuous_param_list, ~.[[1]])

lm_list <- map2(comparison_tbs, comparison_groups, function(tb, grps){
  map(continuous_param_list, ~linear_logistic_regression_modeller(tb, ., grps[1], grps[2], covars=lm_covars, model_return=T))
})

#Naming all the lists and models
names(lm_list) <- comparison_names
for (i in seq(length(lm_list))){
  names(lm_list[[i]]) <- phenotype_names
}

for (i in seq(length(lm_list))){
  print(str_c('Diagnostic plots of linear regression for comparison ',names(lm_list)[i]))
  walk2(lm_list[[i]], phenotype_names,~diagnostic_plotter(.x,.y))
}
```

## Wilcoxon Rank Sum Test

Now we generate comparison plots between the two groups of comparison for all MTC-significant signals from LM and simultaneously run a statistical test for that phenotype.

Unlike the main analysis (where the phenotypes are hardcoded), the phenotypes tested per comparison are taken directly from the FDR-significant LM results.

All NA values are filtered out and resulting counts per group are printed in the plot.

```{r plot_test}
mtc_sig_lm <- cont_lm_summary_tb %>%
  mutate(adj_pval = p.adjust(pval, method='fdr')) %>%
  filter(adj_pval < 0.05) %>%
  select(-adj_pval)

print(mtc_sig_lm$Label_short)

#Run the pairwise Wilcoxon rank sum tests and distribution plotting for each MTC-significant comparison but on the non-transformed phenotypes
wilcoxon_subdirs <- c('thick_thin','mybpc3_plof_nonlof','mybpc3_myh7')

cont_wilcoxon_summary_tb <- pmap(list(comparison_tbs_notransform, comparison_groups, wilcoxon_subdirs), function(tb, grps, subdir){
  sig_phenos <- mtc_sig_lm %>%
    filter(group1 == grps[1], group2 == grps[2]) %>%
    pull(pheno)
  map(
    continuous_param_list[names(continuous_param_list) %in% sig_phenos],
    ~plot_tester(input_tb=tb, ., group1_name=grps[1], group2_name=grps[2], output_path=str_c(output_plot_dir, '2_Wilcoxon/', subdir))
  ) %>%
    bind_rows()
}) %>%
  bind_rows()

cont_wilcoxon_summary_tb
```

## Supplementary: Raw Phenotype Distributions for Significant Findings

This outputs a supplementary panelled figure of the raw (non-transformed, native units) phenotype distributions overlaid as histograms (density-scaled by default) across both subgroups, for every FDR-significant comparison x phenotype finding from the linear regression. One panel row per genotypic comparison; within each, one sub-panel per significant phenotype (free scales as units differ). Dashed lines show group medians; sub-panel titles give the phenotype label (with units).

```{r sig_density_plot, fig.width=12, fig.height=12}
#sig_density_plotter() is defined in genetic_subset_analysis/common/hcmr_endophenotype_functions.R
#Significance taken from the standardised LM results; distributions plotted on raw (non-transformed, LGE not logged) values
sig_density_plotter(
  summary_tb = read_tsv(str_c(output_plot_dir, '1_LM/cont_summary_tb.tsv')),
  comparison_tbs_list = read_rds(str_c(output_data_dir, 'hcmr_comparison_tbs_notransform_powell.rds')),
  param_list = continuous_param_list,
  output_path = str_c(output_plot_dir, '2_Wilcoxon/'),
  comporder = comporder
)
```

## Publication Figure: Panelled Summary Scatter + Raw Phenotype Distributions

This combines the 3-panel summary scatter plot (top) with the raw phenotype histograms for significant findings (underneath) into a single publication figure. Panels are tagged sequentially (A = summary scatter; B onwards = one density row per comparison).

```{r combined_publication_figure, fig.width=12, fig.height=20}
#combined_figure_plotter() is defined in genetic_subset_analysis/common/hcmr_endophenotype_functions.R
combined_figure_plotter(
  summary_tb = read_tsv(str_c(output_plot_dir, '1_LM/cont_summary_tb.tsv')),
  comparison_tbs_list = read_rds(str_c(output_data_dir, 'hcmr_comparison_tbs_notransform_powell.rds')),
  param_list = continuous_param_list,
  output_path = str_c(output_plot_dir, '1_LM/'),
  comporder = comporder
)
```

