---
title: "Comparison of HCMR Case Rare-Variant Carriers on Endophenotype - Summary Tables"
author: "Jonathan Chan"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide   
---
## Import

```{r}
library(tidyverse)
rm(list=ls())

source('genetic_subset_analysis/common/hcmr_endophenotype_functions.R')
```

```{r}
#This imports in the 915 sarcomere-positive individuals
hcmr_sarcpos_tb_notransform <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_sarcpos_pheno_rarevarclasses_noNAfiltered_notransform.tsv')

all_merged_overlapvars_tb <- read_tsv('genetic_subset_analysis/RareVar/3_output/individual_class/overlap_variants/hcmr_KT_overlapvariants.tsv') %>%
  filter(HCR_ID %in% hcmr_sarcpos_tb_notransform$HCR_IDs) #Filter only to the 915 sarcomere-positive individuals

#Identify the number of homozygotes and the number of multiple heterozygotes
homozygotes <- all_merged_overlapvars_tb %>% 
  filter(dosage1==2 | dosage2==2) #There are 2 homozygotes 
multi_het <- all_merged_overlapvars_tb %>%
  filter(!is.na(dosage2))

print(str_c('There are', nrow(homozygotes), 'homozygotes and',length(unique(multi_het$HCR_ID)),'multiple heterozygotes with ', length(intersect(homozygotes$HCR_ID, multi_het$HCR_ID)),'individual being both so technically',nrow(homozygotes)+length(unique(multi_het$HCR_ID))-length(intersect(homozygotes$HCR_ID, multi_het$HCR_ID)), 'individuals with >1 dose of a rare variant', sep=' '))

#These are all the individuals with >1 dosage of rare variant
multi_dosage_ids <- unique(c(homozygotes$HCR_ID,multi_het$HCR_ID))

rm(homozygotes, multi_het, all_merged_overlapvars_tb)
```

```{r}
pergene_rarevar_classification <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_rarevar_class.tsv')
pergene_perpatho_rarevar_classification <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_rarevar_class_plp_vusfpvus.tsv')
```

### Gene Carrier Breakdown

This breaks down all 915 sarcomere-positive individuals (both single- and multi-variant carriers) by the gene(s) in which they carry a rare variant. N.B. Multi-variant carriers across >1 gene are counted once in each gene they carry a variant in, so the per-gene counts sum to more than the number of individuals.

```{r gene_carrier_breakdown}
sarc_genes <- c('MYBPC3_pLOF','MYBPC3_nonLOF','MYH7','TNNI3','TNNT2','TPM1','MYL2','MYL3','ACTC1')

carrier_tb <- pergene_rarevar_classification %>%
  filter(HCR_ID %in% hcmr_sarcpos_tb_notransform$HCR_IDs) %>% #Filter only to the 915 sarcomere-positive individuals
  select(HCR_ID, all_of(sarc_genes)) %>%
  mutate(MYBPC3 = MYBPC3_pLOF + MYBPC3_nonLOF) #Any MYBPC3 carrier (pLOF and/or nonLOF)

#Number of carriers (dosage >=1) per gene
gene_carrier_counts <- carrier_tb %>%
  summarise(across(-HCR_ID, ~sum(. >= 1, na.rm=T))) %>%
  pivot_longer(everything(), names_to='Gene', values_to='Number of Carriers') %>%
  mutate(Gene = factor(Gene, levels=c('MYBPC3', sarc_genes))) %>%
  arrange(Gene) #MYBPC3 total is kept in the table alongside the pLOF/nonLOF breakdown

print(gene_carrier_counts) #This is for all 915 so both single and multiple variant carriers so total sum in column > 915
print(str_c('Total number of sarcomere-positive individuals:', n_distinct(carrier_tb$HCR_ID), sep=' '))

write_csv(gene_carrier_counts, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/gene_carrier_counts.csv')

#Single horizontal stacked bar for the Central Illustration - MYBPC3 kept as a single block (pLOF + nonLOF combined)
gene_carrier_plot <- gene_carrier_counts %>%
  filter(!Gene %in% c('MYBPC3_pLOF','MYBPC3_nonLOF')) %>%
  mutate(Gene = fct_rev(fct_drop(Gene))) %>% #So that MYBPC3 sits at the left of the bar
  ggplot(aes(x=`Number of Carriers`, y='', fill=Gene)) +
  geom_col(colour='white', width=0.6) +
  scale_x_continuous(expand=c(0,0)) +
  guides(fill=guide_legend(reverse=T, nrow=1)) +
  labs(x=NULL, y=NULL, fill=NULL) +
  theme_classic(base_size=36) + #2x the default base_size of 11
  theme(axis.line.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank(),
        legend.position='bottom',
        panel.background=element_rect(fill='transparent', colour=NA),
        plot.background=element_rect(fill='transparent', colour=NA),
        legend.background=element_rect(fill='transparent', colour=NA))

print(gene_carrier_plot)

ggsave('genetic_subset_analysis/RareVar/3_output/summary_tbs/gene_carrier_stacked_bar.png', gene_carrier_plot, width=16, height=3.5, dpi=300, bg='transparent')
ggsave('genetic_subset_analysis/RareVar/3_output/summary_tbs/gene_carrier_stacked_bar.pdf', gene_carrier_plot, width=16, height=3.5, bg='transparent')

rm(sarc_genes, carrier_tb, gene_carrier_plot)
```

This outputs the number of individuals in each class.

```{r}
hcmr_singlevar <- hcmr_sarcpos_tb_notransform %>%
  filter(!HCR_IDs %in% multi_dosage_ids) %>%
  left_join(pergene_rarevar_classification, by=c('HCR_IDs'='HCR_ID')) %>%
  left_join(pergene_perpatho_rarevar_classification, by=c('HCR_IDs'='HCR_ID')) %>%
  mutate(patho_class = ifelse(str_detect(class_secondary, 'P_LP'), 'P_LP', 'VUS'))

print(hcmr_singlevar %>%
  group_by(class_primary) %>%
  summarise(n=n()) %>%
  arrange(desc(n)))

print(table(hcmr_singlevar$class_primary, hcmr_singlevar$patho_class))
```

```{r}
#This code generates the comparison lists
#For primary classification comparisons
#For thick vs. thin (excluding ACTC1 from Thin btw)
thick_class <- unique(hcmr_sarcpos_tb_notransform$class_primary)[!str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'ACTC1|TPM1|TNNI3|TNNT2') & str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYBPC3|MYH7|MYL2|MYL3')]
thin_class <- unique(hcmr_sarcpos_tb_notransform$class_primary)[str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'TPM1|TNNI3|TNNT2') & !str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'ACTC1|MYBPC3|MYH7|MYL2|MYL3')]

#For mybpc3_pLOF vs. mybpc3_nonlof
mybpc3_plof_class <- unique(hcmr_sarcpos_tb_notransform$class_primary)[str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYBPC3_pLOF') & !str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYBPC3_nonLOF')]
mybpc3_nonlof_class <- unique(hcmr_sarcpos_tb_notransform$class_primary)[str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYBPC3_nonLOF') & !str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYBPC3_pLOF')]

#For mybpc3 vs. MYH7
mybpc3_class <- unique(hcmr_sarcpos_tb_notransform$class_primary)[str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYBPC3') & !str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYH7')]
myh7_class <- unique(hcmr_sarcpos_tb_notransform$class_primary)[str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYH7') & !str_detect(unique(hcmr_sarcpos_tb_notransform$class_primary),'MYBPC3')]

#For secondary classification comparison
#For P/LP vs. VUS_VUS_FP
p_lp_class <- unique(hcmr_sarcpos_tb_notransform$class_secondary)[str_detect(unique(hcmr_sarcpos_tb_notransform$class_secondary),'P_LP') & !str_detect(unique(hcmr_sarcpos_tb_notransform$class_secondary), 'VUS')] #I use this because I wish to exclude all the individuals who have a combination across P_LP and VUS_VUS_FP
vus_vusfp_class <- unique(hcmr_sarcpos_tb_notransform$class_secondary)[str_detect(unique(hcmr_sarcpos_tb_notransform$class_secondary),'VUS') & !str_detect(unique(hcmr_sarcpos_tb_notransform$class_secondary), 'P_LP')] 

#---------------------------------------------------------------------------------------
#Parameter lists for each comparison
thick_thin_param_list <- list(thick_class, 'Thick',thin_class, 'Thin')
mybpc3_plof_nonlof_param_list <- list(mybpc3_plof_class, 'MYBPC3_pLOF', mybpc3_nonlof_class, 'MYBPC3_nonLOF')
mybpc3_myh7_param_list <- list(mybpc3_class, 'MYBPC3',myh7_class,'MYH7')
p_lp_vus_vusfp_param_list <- list(p_lp_class, 'P/LP', vus_vusfp_class, 'VUS')

rm(list=ls(pattern='class$'))
```

## Publication Display Items

This provides the tables for the display items.

### Baseline Summary Table

```{r table_display_item}
baseline_summary_printer <- function(list_tbs, group_vector_list, var, varname, cont_or_discrete='continuous', order=T){ #Input is a list of tibbles each corresponding to a column

    if(length(list_tbs)==4){
      k <- 4 #This defaults to using the P/LP vs VUS comparison for total
    } else{
      k <- 1 #Otherwise it uses the first tibble in the list of comparison tibbles
    }
  
  if (cont_or_discrete =='continuous'){
    
    g1_means <- map2(list_tbs,group_vector_list, ~ mean(filter(.x, group==.y[1])[[var]], na.rm=T)) #Iterates over every comparison_tb
    g2_means <- map2(list_tbs,group_vector_list, ~ mean(filter(.x, group==.y[2])[[var]], na.rm=T))
    g1_sdevs <- map2(list_tbs,group_vector_list, ~ sd(filter(.x, group==.y[1])[[var]], na.rm=T))
    g2_sdevs <- map2(list_tbs,group_vector_list, ~ sd(filter(.x, group==.y[2])[[var]], na.rm=T))
    
    total_mean <- mean(list_tbs[[k]][[var]],na.rm=T)
    total_sd <- sd(list_tbs[[k]][[var]],na.rm=T)
    
    output_tb <- bind_cols(g1_means) %>% bind_cols(g1_sdevs) %>%bind_cols(g2_means) %>% bind_cols(g2_sdevs) %>%
      mutate(Total_mean=total_mean, Total_sd=total_sd) %>%
      mutate(across(everything(),~round(.,2))) %>%
      mutate(Phenotype=varname) %>%
      select(Phenotype, Total_mean, Total_sd,everything())
    
    g1_names <- sapply(group_vector_list, "[[", 1)
    g2_names <- sapply(group_vector_list, "[[", 2)
    
    colnames(output_tb) <- c('Phenotype',
                             'Total_mean',
                             'Total_sd',
                         str_c(g1_names,'_mean'),
                         str_c(g1_names,'_sd'),
                         str_c(g2_names,'_mean'),
                         str_c(g2_names,'_sd'))
  
    number_comparisons <- length(group_vector_list)
    number_other_cols <- 3
    
    if(isTRUE(order)){
          #Ordering the columns correctly
    output_tb <- output_tb %>%
      select(Phenotype, Total_mean, Total_sd,
             1+number_other_cols,1+number_comparisons+number_other_cols,1+2*number_comparisons+number_other_cols, 1+3*number_comparisons+number_other_cols,
             2+number_other_cols,2+number_comparisons+number_other_cols,2+2*number_comparisons+number_other_cols, 2+3*number_comparisons+number_other_cols,
             3+number_other_cols,3+number_comparisons+number_other_cols,3+2*number_comparisons+number_other_cols, 3+3* number_comparisons+number_other_cols,
             4+number_other_cols,4+number_comparisons+number_other_cols, 4+2*number_comparisons+number_other_cols,4+3*number_comparisons+number_other_cols
             )
    }

    
  } else if (cont_or_discrete =='discrete'){
    
    g1_nums <- map2(list_tbs,group_vector_list, ~ sum(filter(.x, group==.y[1])[[var]], na.rm=T))
    g2_nums <- map2(list_tbs,group_vector_list, ~ sum(filter(.x, group==.y[2])[[var]], na.rm=T))
    g1_percent <- map2(list_tbs,group_vector_list, ~ sum(filter(.x, group==.y[1])[[var]], na.rm=T) / sum(!is.na(filter(.x, group==.y[1])[[var]])) * 100)
    g2_percent <- map2(list_tbs,group_vector_list, ~ sum(filter(.x, group==.y[2])[[var]], na.rm=T)/ sum(!is.na(filter(.x, group==.y[2])[[var]])) * 100)
    
    total_num <- sum(list_tbs[[k]][[var]], na.rm=T) 
    total_percent <- sum(list_tbs[[k]][[var]], na.rm=T) /  sum(!is.na(list_tbs[[k]][[var]])) * 100
      
    output_tb <- bind_cols(g1_nums) %>% bind_cols(g1_percent) %>%bind_cols(g2_nums) %>% bind_cols(g2_percent) %>%
      mutate(Phenotype=varname) %>%
      mutate(Total_num=total_num, Total_percent=total_percent) %>%
      select(Phenotype, Total_num, Total_percent,everything())
    
    g1_names <- sapply(group_vector_list, "[[", 1)
    g2_names <- sapply(group_vector_list, "[[", 2)
    
    colnames(output_tb) <- c('Phenotype',
                             'Total_num',
                             'Total_percent',
                         str_c(g1_names,'_num'),
                         str_c(g1_names,'_percent'),
                         str_c(g2_names,'_num'),
                         str_c(g2_names,'_percent'))
    
    output_tb <-  output_tb %>%
      mutate(across(ends_with('num'),~round(.,0)),
             across(ends_with('percent'), ~round(.,1)))
  
    number_comparisons <- length(group_vector_list)
    number_other_cols <- 3
    
    if(isTRUE(order)){
    #Ordering the columns correctly but this assumes 4 comparisons
    output_tb <- output_tb %>%
      select(Phenotype,Total_num, Total_percent, 
             1+number_other_cols,1+number_comparisons+number_other_cols,1+2*number_comparisons+number_other_cols, 1+3*number_comparisons+number_other_cols,
             2+number_other_cols,2+number_comparisons+number_other_cols,2+2*number_comparisons+number_other_cols, 2+3*number_comparisons+number_other_cols,
             3+number_other_cols,3+number_comparisons+number_other_cols,3+2*number_comparisons+number_other_cols, 3+3* number_comparisons+number_other_cols,
             4+number_other_cols,4+number_comparisons+number_other_cols, 4+2*number_comparisons+number_other_cols,4+3*number_comparisons+number_other_cols
             )
    }
  }
  
  return(output_tb)
}
```

This provides the baseline summary table output (Table 1) of the 892 sarcomere-positive patients (915 sarcpos --> 892 single-variant individuals).

```{r baseline_table}
baseline_cont_phenos <- c('age','bmi','holvtrun') #As per Neubaeur et al, 2019
baseline_cont_phenonames <- c('Age (years)','BMI (kg/msq)','Number of VT Runs')
baseline_discrete_phenos <- c('N','genderMale','race_white','race_black','race_SA','race_EA','race_other','hxhcm1st','hxhcm2nd','hxhcm1st2nd','hxhtn','diatyp2','cursmoke','nyha1','nyha2','nyha3_4','hxsync','hxhf','hxstrk','holtnsvt','hxafib_persist','hxafib_paroxy','blcp','bldysp','blbeta','blcalc','blacearb','bldisop','blstat','bldiur','blantico','blantipl')
baseline_discrete_phenonames <- c('N','Male','White','Black','South Asian','East Asian','Other','1st degree','2nd degree', 'Both 1st and 2nd','Hypertension','Type 2 Diabetes', 'Current Smoker', 'Class I','Class II','Class III/IV','Syncope','Heart Failure', 'Stroke','Non-sustained Ventricular Tachycardia','Persistent','Paroxysysmal','Chest Pain','Dyspnea','Beta-blockers','Calcium-channel blocker','ACE/ARB', 'Disopyramide','Statins','Diuretics','Oral anticoagulant','Oral antiplatelet agent')

baseline_tb<- hcmr_sarcpos_tb_notransform %>%
  mutate(genderMale=ifelse(gender=='Male',T,F)) %>%
  mutate(race_white=ifelse(race=='White',T,F),
         race_black = ifelse(race=='Black/Afr.Am',T,F),
         race_SA=ifelse(race =='SA',T,F),
         race_EA=ifelse(race=='EA',T,F),
         race_other=ifelse(race %in% c('Pacific Islander',' American Indian/Alaskan'),T,F)) %>%
  mutate(hxhcm1st = ifelse(hxhcm1st >=1,T,F),
         hxhcm2nd=ifelse(hxhcm2nd >=1, T, F)) %>%
  mutate(hxhcm1st2nd = ifelse(hxhcm1st & hxhcm2nd, T, F)) %>%
  mutate(nyha1=ifelse(blnyha==1, T,F),
         nyha2=ifelse(blnyha==2,T,F),
         nyha3_4=ifelse(blnyha %in% c(3,4),T,F)) %>%
  mutate(hxafib_persist=ifelse(hxafib=='Persistent',T,F),
         hxafib_paroxy=ifelse(hxafib=='Paroxysmal',T,F)) %>%
  mutate(N=1) %>%
  select(HCR_IDs, class_primary, class_secondary, all_of(baseline_cont_phenos),all_of(baseline_discrete_phenos))

baseline_tb_singlevar <- baseline_tb %>%
    filter(!HCR_IDs %in% multi_dosage_ids)#Filter for only 892 single-variant individuals
```

This outputs the baseline summary tables for the entire 892 individuals as well as the  pairwise comparisons.

```{r}
#This creates the sub-tables across the binary comparisons made
temp <- map(list(thick_thin_param_list,mybpc3_plof_nonlof_param_list,mybpc3_myh7_param_list),~comparison_pheno_tb_gen(phenotype_tb = baseline_tb_singlevar, param_list=.,rarevar_class_tb = pergene_rarevar_classification, var_class='primary'))
names(temp) <- c('thick_thin_tb','mybpc3_plof_nonlof_tb','mybc3_myh7_tb')

temp2<- map(list(p_lp_vus_vusfp_param_list),~comparison_pheno_tb_gen(phenotype_tb = baseline_tb_singlevar, param_list=.,rarevar_class_tb = pergene_perpatho_rarevar_classification, var_class='secondary'))
names(temp2) <- c('p_lp_vus_vusfp_tb')

comparison_tbs_baseline <- c(temp, temp2)
rm(temp,temp2)

#Collating the list of tibbles 
group_list <- list(c('Thick','Thin'),
                    c('MYBPC3_pLOF','MYBPC3_nonLOF'),
                    c('MYBPC3','MYH7'),
                    c('P/LP','VUS'))

baseline_cont_summary_tb <- map2(baseline_cont_phenos, baseline_cont_phenonames, ~baseline_summary_printer(comparison_tbs_baseline,group_list,.x,.y)) %>% bind_rows()
baseline_discrete_summary_tb <- map2(baseline_discrete_phenos, baseline_discrete_phenonames, ~baseline_summary_printer(comparison_tbs_baseline,group_list,.x,.y, cont_or_discrete = 'discrete')) %>% bind_rows()

write_csv(baseline_cont_summary_tb, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/baseline_summary_continuous.csv')
write_csv(baseline_discrete_summary_tb, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/baseline_summary_discrete.csv')
```

This outputs the baseline summary tables for the multi-dosage individuals vs. single-dosage individuals.

```{r}
#Repeat for the multiple dosage individuals only
baseline_tb_multivar <- baseline_tb %>%
  mutate(group = case_when(HCR_IDs %in% multi_dosage_ids~'Multi',
                           T~'Single'))

#Outputting the summary tibbles
baseline_cont_summary_tb_multiple <- map2(baseline_cont_phenos, baseline_cont_phenonames, ~baseline_summary_printer(list(baseline_tb_multivar),list(c('Single','Multi')),.x,.y, order=F))  %>% bind_rows()

baseline_discrete_summary_tb_multiple <- map2(baseline_discrete_phenos, baseline_discrete_phenonames, ~baseline_summary_printer(list(baseline_tb_multivar),list(c('Single','Multi')),.x,.y, order=F, cont_or_discrete = 'discrete'))  %>% bind_rows()

write_csv(baseline_cont_summary_tb_multiple, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/baseline_summary_multiple_continuous.csv')
write_csv(baseline_discrete_summary_tb_multiple, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/baseline_summary_multiple_discrete.csv')

rm(list=ls(pattern='summary_tb'))
rm(list=ls(pattern='baseline'))
```

### CMR Phenotype Table

This plots the summary tables but focusing on the endophenotypes specifically.

This creates the comparison tibbles.
```{r}
#Define the continuous parameters you want to get a summary for
phenos_of_interest <- c('lge_total','wallthkmax','lvmassi','lge_total','ecvfwhole','lvef','lvs_ellmean','lvs_eccmean','lvs_errmean')

cmr_tb <- hcmr_sarcpos_tb_notransform %>%
  mutate(N=1) %>%
  select(HCR_IDs, class_primary, class_secondary, all_of(phenos_of_interest))

cmr_tb_singlevar <- cmr_tb %>%
    filter(!HCR_IDs %in% multi_dosage_ids)#Filter for only 892 single-variant individuals

#This creates the sub-tables across the binary comparisons made
temp <- map(list(thick_thin_param_list,mybpc3_plof_nonlof_param_list,mybpc3_myh7_param_list),~comparison_pheno_tb_gen(phenotype_tb = cmr_tb_singlevar, param_list=.,rarevar_class_tb = pergene_rarevar_classification, var_class='primary'))
names(temp) <- c('thick_thin_tb','mybpc3_plof_nonlof_tb','mybc3_myh7_tb')

temp2<- map(list(p_lp_vus_vusfp_param_list),~comparison_pheno_tb_gen(phenotype_tb = cmr_tb_singlevar, param_list=.,rarevar_class_tb = pergene_perpatho_rarevar_classification, var_class='secondary'))
names(temp2) <- c('p_lp_vus_vusfp_tb')

comparison_tbs_cmr <- c(temp, temp2)
rm(temp,temp2, phenos_of_interest)
```

```{r cmr_phenotype_table}
continuous_summary_printer <- function(list_tbs, group_vector_list, varlist, mean_or_median='mean', column_order=T){ #Input is a list of tibbles each corresponding to a column
    
  var=varlist[[1]]
  varname=varlist[[2]]
  
  if(mean_or_median=='mean'){
    g1_means <- map2(list_tbs,group_vector_list, ~ mean(filter(.x, group==.y[1])[[var]], na.rm=T)) #Iterates over every comparison_tb
    g2_means <- map2(list_tbs,group_vector_list, ~ mean(filter(.x, group==.y[2])[[var]], na.rm=T))
    g1_sdevs <- map2(list_tbs,group_vector_list, ~ sd(filter(.x, group==.y[1])[[var]], na.rm=T))
    g2_sdevs <- map2(list_tbs,group_vector_list, ~ sd(filter(.x, group==.y[2])[[var]], na.rm=T))
    pvals <- map2(list_tbs, group_vector_list, ~ wilcox.test(filter(.x, group==.y[1])[[var]],filter(.x, group==.y[2])[[var]], na.action='na.omit')$p.value)
    
    total_mean <- mean(list_tbs[[1]][[var]],na.rm=T)
    total_sd <- sd(list_tbs[[1]][[var]],na.rm=T)
    
    output_tb <- bind_cols(g1_means) %>% bind_cols(g1_sdevs) %>%bind_cols(g2_means) %>% bind_cols(g2_sdevs) %>%
      mutate(Total_mean=total_mean, Total_sd=total_sd) %>%
      mutate(across(everything(),~round(.,2))) %>%
      bind_cols(pvals) %>%
      mutate(Phenotype=varname) %>%
      select(Phenotype, Total_mean, Total_sd,everything())
    
    g1_names <- sapply(group_vector_list, "[[", 1)
    g2_names <- sapply(group_vector_list, "[[", 2)
    
    colnames(output_tb) <- c('Phenotype',
                             'Total_mean',
                             'Total_sd',
                         str_c(g1_names,'_mean'),
                         str_c(g1_names,'_sd'),
                         str_c(g2_names,'_mean'),
                         str_c(g2_names,'_sd'),
                         str_c(g1_names,'_vs_',g2_names,'_pval'))
    
    output_tb <- output_tb %>%
      mutate(across(ends_with('pval'), ~signif(.,2)))

    number_comparisons <- length(group_vector_list)
    number_other_cols <- 3
    
    if(isTRUE(column_order)){
          #Ordering the columns correctly - N.B This assumes that there are 4 comparisons
    output_tb <- output_tb %>%
      select(Phenotype, Total_mean, Total_sd,
             1+number_other_cols,1+number_comparisons+number_other_cols,1+2*number_comparisons+number_other_cols, 1+3*number_comparisons+number_other_cols,4+3*number_comparisons+number_other_cols+1,
             2+number_other_cols,2+number_comparisons+number_other_cols,2+2*number_comparisons+number_other_cols, 2+3*number_comparisons+number_other_cols,4+3*number_comparisons+number_other_cols+2,
             3+number_other_cols,3+number_comparisons+number_other_cols,3+2*number_comparisons+number_other_cols, 3+3* number_comparisons+number_other_cols,4+3*number_comparisons+number_other_cols+3,
             4+number_other_cols,4+number_comparisons+number_other_cols, 4+2*number_comparisons+number_other_cols,4+3*number_comparisons+number_other_cols,4+3*number_comparisons+number_other_cols+4,
             )
    }

  
  } else if (mean_or_median=='median'){
    
    g1_medians <- map2(list_tbs,group_vector_list, ~ median(filter(.x, group==.y[1])[[var]], na.rm=T)) #Iterates over every comparison_tb
    g2_medians <- map2(list_tbs,group_vector_list, ~ median(filter(.x, group==.y[2])[[var]], na.rm=T))
    g1_fqs <- map2(list_tbs,group_vector_list, ~ quantile(filter(.x, group==.y[1])[[var]],0.25, na.rm=T))
    g2_fqs <- map2(list_tbs,group_vector_list, ~ quantile(filter(.x, group==.y[2])[[var]],0.25, na.rm=T))
    g1_tqs <- map2(list_tbs,group_vector_list, ~ quantile(filter(.x, group==.y[1])[[var]],0.75, na.rm=T))
    g2_tqs <- map2(list_tbs,group_vector_list, ~ quantile(filter(.x, group==.y[2])[[var]],0.75, na.rm=T))
    pvals <- map2(list_tbs, group_vector_list, ~ wilcox.test(filter(.x, group==.y[1])[[var]],filter(.x, group==.y[2])[[var]], na.action='na.omit')$p.value)
    
    total_median <- median(list_tbs[[1]][[var]],na.rm=T)
    total_fq <- quantile(list_tbs[[1]][[var]],0.25,na.rm=T)
    total_tq <- quantile(list_tbs[[1]][[var]],0.75,na.rm=T)
    
    output_tb <- bind_cols(g1_medians) %>% bind_cols(g1_fqs) %>%bind_cols(g1_tqs) %>% bind_cols(g2_medians) %>%bind_cols(g2_fqs) %>% bind_cols(g2_tqs) %>% 
      mutate(Total_median=total_median, Total_fq=total_fq, Total_tq=total_tq) %>%
      mutate(across(everything(),~round(.,2))) %>%
      bind_cols(pvals) %>%
      mutate(Phenotype=varname) %>%
      mutate(Total = str_c(Total_median,' (',Total_fq,'-',Total_tq,')')) %>%
      select(Phenotype, Total, everything()) %>%
      select(-Total_median, -Total_fq, -Total_tq)
    
    g1_names <- sapply(group_vector_list, "[[", 1)
    g2_names <- sapply(group_vector_list, "[[", 2)
    
    colnames(output_tb) <- c('Phenotype',
                             'Total',
                         str_c(g1_names,'_median'),
                         str_c(g1_names,'_fq'),
                         str_c(g1_names,'_tq'),
                         str_c(g2_names,'_median'),
                         str_c(g2_names,'_fq'),
                         str_c(g2_names,'_tq'),
                         str_c(g1_names,'_vs_',g2_names,'_pval'))
    
     output_tb <- output_tb %>%
      mutate(across(ends_with('pval'), ~signif(.,2)))
    
    for (name in c(g1_names, g2_names)){
      median_colname <- str_c(name, '_median')
      fq_colname <- str_c(name,'_fq')
      tq_colname <- str_c(name,'_tq')
      
      output_tb[[name]] <- str_c(output_tb[[median_colname]],' (', output_tb[[fq_colname]],'-',output_tb[[tq_colname]],')')
    }
     
     output_tb <- output_tb %>%
       select(Phenotype, Total, all_of(c(g1_names, g2_names)),ends_with('pval'))
     
    number_comparisons <- length(group_vector_list)
    number_other_cols <- 2
    
        if(isTRUE(column_order)){
    #Ordering the columns correctly N.B This assumes 4 comparisons
    output_tb <- output_tb %>%
      select(Phenotype, Total,
             1+number_other_cols,1+number_comparisons+number_other_cols,4+number_comparisons+number_other_cols+1,
             2+number_other_cols,2+number_comparisons+number_other_cols,4+number_comparisons+number_other_cols+2,
             3+number_other_cols,3+number_comparisons+number_other_cols,4+number_comparisons+number_other_cols+3,
             4+number_other_cols,4+number_comparisons+number_other_cols,4+number_comparisons+number_other_cols+4,
             )
        }
      
  }
  
  return(output_tb)
}
```

```{r}
continuous_param_list_for_summary<- list(
  c('wallthkmax','Maximal LV Wall Thickness (mm)'),
  c('lvmassi','Left Ventricular Mass Index (g/m2)'),
  c('ecvfwhole','Whole LV Extracellular Volume Fraction (%)'),
  c('lvef','Left ventricular Ejection Fraction (%)'),
  c('lvs_ellmean','-Longitudinal LV Strain'),
  c('lvs_eccmean','-Circumferential LV Strain'),
  c('lvs_errmean','Radial LV Strain')
)
lge_summary_vector <- c('lge_total','Total LGE over All Segments (6 SD) (%)')

#This creates the summary_tb itself
continuous_summary_tb <- map(continuous_param_list_for_summary, ~continuous_summary_printer(comparison_tbs_cmr, group_list,., mean_or_median='mean')) %>% bind_rows()
lge_summary_tb <-  continuous_summary_printer(comparison_tbs_cmr, group_list,lge_summary_vector, mean_or_median='median') %>% bind_rows()

write_csv(continuous_summary_tb, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/cmrpheno_summary_continuous.csv')
write_csv(lge_summary_tb, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/cmrlge_summary_continuous.csv')

rm(continuous_summary_tb, lge_summary_tb, comparison_tbs_cmr, group_list)
```

This compares the single vs. multi-variant individuals on the CMR endophenotypes.

```{r multidosage_phenotype_table}
cmr_tb_multivar <- cmr_tb %>%
  mutate(group = case_when(HCR_IDs %in% multi_dosage_ids~'Multi',
                           T~'Single'))

group_list <- list(c('Single','Multi'))

#Outputting the summary tibbles
continuous_summary_tb <- map(continuous_param_list_for_summary, ~continuous_summary_printer(list(cmr_tb_multivar), group_list,., mean_or_median='mean',column_order=F)) %>% bind_rows()
lge_summary_tb <-  continuous_summary_printer(list(cmr_tb_multivar), group_list,lge_summary_vector, mean_or_median='median',column_order=F) %>% bind_rows()

write_csv(continuous_summary_tb, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/multivssingle_cmrpheno_summary_continuous.csv')
write_csv(lge_summary_tb, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/multivssingle_cmrlge_summary_continuous.csv')

rm(list=ls(pattern='summary_tb'))
rm(list=ls(pattern='cmr'))
```

This outputs the hazard ratios as a Table 3 for each pairwise comparison.

```{r}
cox_results_tb <- read_rds('genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hcmr_rarevar_subset_cox_age_timescale.rds')

cox_results_summary_tb <- cox_results_tb %>%
  mutate(Comparison = str_replace_all(Group, "\\\n", ' ')) %>%
  filter(Comparison %in% c('Thick vs. Thin',
                           'MYBPC3_pLOF vs. MYBPC3_nonLOF',
                           'MYBPC3 vs. MYH7',
                           'Sarcomere-P/LP vs. Sarcomere-VUS')) %>%
  select(Comparison, Composite=composite, HR=hr, HR_lowerCI=hr_low, HR_upperCI=hr_high, pval=p.value) %>%
  mutate(HR=round(HR,2),
         HR_lowerCI=round(HR_lowerCI,2),
         HR_upperCI=round(HR_upperCI,2),
         pval=signif(pval,2)) %>%
  mutate(HR_CI = str_c(HR,' (',HR_lowerCI,'-',HR_upperCI,')')) %>%
  select(Comparison, Composite, HR_CI, pval)

write_tsv(cox_results_summary_tb, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/cox_results_summary.tsv')
```

This outputs the hazard ratios for the single-vs-multiple as a Supp. Table.


```{r}
cox_results_tb <- read_rds('genetic_subset_analysis/RareVar/3_output/plots/2_Severity/1_Cox/1_AgeAsTimescale/hcmr_rarevar_subset_cox_age_timescale_multi.rds')

cox_results_summary_tb <- cox_results_tb %>%
  mutate(Comparison = str_replace_all(Group, "\\\n", ' ')) %>%
  filter(Comparison %in% c('Multi vs. Single')) %>%
  select(Comparison, Composite=composite, HR=hr, HR_lowerCI=hr_low, HR_upperCI=hr_high, pval=p.value) %>%
  mutate(HR=round(HR,2),
         HR_lowerCI=round(HR_lowerCI,2),
         HR_upperCI=round(HR_upperCI,2),
         pval=signif(pval,2)) %>%
  mutate(HR_CI = str_c(HR,' (',HR_lowerCI,'-',HR_upperCI,')')) %>%
  select(Comparison, Composite, HR_CI, pval)

write_tsv(cox_results_summary_tb, 'genetic_subset_analysis/RareVar/3_output/summary_tbs/cox_results_summary_multi.tsv')
```