---
title: "Comparison of HCMR Case Rare-Variant Carriers on Endophenotype - Multi-Variant Individuals Analysis"
author: "Jonathan Chan"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(factoextra)
library(tictoc)
library(readxl)

rm(list=ls())

theme_set(theme_classic(base_size=14))

source('genetic_subset_analysis/common/hcmr_endophenotype_functions.R')
```

## Import

```{r import, echo=F}
hcmr_sarcpos_tb <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_sarcpos_pheno_rarevarclasses_transform.tsv')
hcmr_sarcpos_tb_notransform_lgelogged <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_sarcpos_pheno_rarevarclasses_notransform_lgelogged.tsv')
hcmr_sarcpos_tb_notransform <- read_tsv('genetic_subset_analysis/RareVar/3_output/data/hcmr_sarcpos_pheno_rarevarclasses_notransform.tsv')
```

## Phenotypes for Comparison

The following phenotypes will be analysed each reflecting some component of HCM disease.

- Hypertrophy: wallthkmax; lvmassi
- Fibrosis: lge_total; ecvfwhole
- Hypercontractility: LVEF (from CMR) and LV strain parameters (lvs_ellmean; eccmean; errmean)

Note that lge_total is treated both as a continuous and a discrete trait via binning.

```{r phenotype_tb}
phenos_of_interest <- c('lge_total','wallthkmax','lvmassi','lge_total','ecvfwhole','lvef','lvs_ellmean','lvs_eccmean','lvs_errmean')
contphenos <- phenos_of_interest

continuous_param_list <- list(
  c('wallthkmax','Maximal LV Wall Thickness (mm)'),
  c('lvmassi','LV Mass Index (g/m2)'),
  c('lge_total','LGE Mass over All Segments / LV Mass (%)'),
  c('ecvfwhole','Whole LV Extracellular Volume Fraction (%)'),
  c('lvef','LV Ejection Fraction (%)'),
  c('lvs_ellmean','-Longitudinal LV Strain'),
  c('lvs_eccmean','-Circumferential LV Strain'),
  c('lvs_errmean','Radial LV Strain')
  # c('entropy_LVWall_MAG_adj','Entropy of LV Wall from MAG LGE Images'),
  # c('entropy_LVWall_PSIR_adj','Entropy of LV Wall from PSIR LGE Images')
)

scaled_cols <- c('lge_total')

names(continuous_param_list) <- map_chr(continuous_param_list, ~.[[1]])

label_tb <- read_tsv('output/data/combo_labels_20260123.tsv')
```

# Multi-Variant Individuals Analysis

This section analyses the phenotypes of the individuals who have more than one rare variant dosage in KT_HCMR19.

This subset can be broken down into a few sub-types:

1) Homozygotes for a particular variant
2) Heterozygotes for 2 or more variants

```{r multi-dosage individuals}

all_merged_overlapvars_tb <- read_tsv('genetic_subset_analysis/RareVar/3_output/individual_class/overlap_variants/hcmr_KT_overlapvariants.tsv') %>%
  filter(HCR_ID %in% hcmr_sarcpos_tb$HCR_IDs) #Filter only to the 882 sarcomere-positive individuals

#Identify the number of homozygotes and the number of multiple heterozygotes
homozygotes <- all_merged_overlapvars_tb %>% 
  filter(dosage1==2 | dosage2==2) #There are 2 homozygotes 
multi_het <- all_merged_overlapvars_tb %>%
  filter(!is.na(dosage2))

print(str_c('There are', nrow(homozygotes), 'homozygotes and',length(unique(multi_het$HCR_ID)),'multiple heterozygotes with ', length(intersect(homozygotes$HCR_ID, multi_het$HCR_ID)),'individual being both so technically',nrow(homozygotes)+length(unique(multi_het$HCR_ID))-length(intersect(homozygotes$HCR_ID, multi_het$HCR_ID)), 'individuals with >1 dose of a rare variant', sep=' '))

#These are all the individuals with >1 dosage of rare variant
multi_dosage_ids <- unique(c(homozygotes$HCR_ID,multi_het$HCR_ID))
```

Then generate the phenotype tbs for the single vs. multiple-variant comparison.

```{r}
#Generate the phenotype tb for comparison
multi_vs_single_tb <- hcmr_sarcpos_tb %>%
  mutate(group = ifelse(HCR_IDs %in% multi_dosage_ids, 'Multi','Single')) %>%
  mutate(group=factor(group, levels=c('Single','Multi'))) 

multi_vs_single_tb_notransform <- hcmr_sarcpos_tb_notransform%>%
  mutate(group = ifelse(HCR_IDs %in% multi_dosage_ids, 'Multi','Single')) %>%
  mutate(group=factor(group, levels=c('Single','Multi')))
```

## Single vs. Multiple Variant Individual Comparison

```{r multi-dosage individuals_comparison}
#This tests all individuals with >1 rare variant dosage against the other sarcomere-positive individuals

#Linear regression to adjust for age/sex etc. results
multi_vs_single_lm_results <- map(continuous_param_list, ~linear_logistic_regression_modeller(multi_vs_single_tb, .,'Multi','Single', covars=c('age','sex','bmi','FieldStrength'))) %>% bind_rows()  %>%
  mutate(comparison = str_c(group1, ' vs. ', group2)) %>%
  left_join(select(label_tb,Variable, Label),by=c('pheno'='Variable')) %>%
  mutate(estimate = - estimate) %>% #Invert the estimate to reflect group1 vs. group2 
  select(-term) %>%
  mutate(Label_short = case_when(
    pheno == 'wallthkmax' ~ 'maxLVWT',
    pheno == 'lvmassi' ~ 'LVMi',
    pheno == 'lge_total' ~ 'LGE',
    pheno == 'ecvfwhole' ~ 'ECVF',
    pheno == 'lvef' ~ 'LVEF',
    pheno == 'lvs_ellmean' ~ '-LongStrain',
    pheno == 'lvs_eccmean' ~ '-CircStrain',
    pheno == 'lvs_errmean' ~ 'RadStrain'
  )) %>%
    mutate(Label_short = str_c(Label_short, ' (', comparison,')')) %>%
  mutate(Label= case_when(
      pheno %in% scaled_cols ~ str_replace(Label, '\\([^ ]+\\)$', '(SD(Log(%)))'),
      pheno %in% phenos_of_interest ~ str_replace(Label, '\\([^ ]+\\)$', '(SD)')))

multi_vs_single_lm_results <- colour_hardcoder(multi_vs_single_lm_results)


summary_plotter(multi_vs_single_lm_results,output_path='genetic_subset_analysis/RareVar/3_output/plots/1_Endophenotype/1_LM/MultiVariant/',  mtc='fdr',test='LinearRegression',xlabel='Coefficient Estimate',confint=T, label=T, comporder = c('Multi vs. Single'))

write_tsv(multi_vs_single_lm_results,'genetic_subset_analysis/RareVar/3_output/plots/1_Endophenotype/1_LM/MultiVariant/allmulti_vs_single_lm_results.tsv')

summary_plotter(multi_vs_single_lm_results,output_path='genetic_subset_analysis/RareVar/3_output/plots/1_Endophenotype/1_LM/MultiVariant/',  mtc='fdr',test='LinearRegression',xlabel='Coefficient Estimate',confint=T, poster=T, comporder = c('Multi vs. Single'), height=6, width=10)
```

This plots the summary plot as a facetted forest plot instead.

```{r}
multi_vs_single_lm_results <- read_tsv('genetic_subset_analysis/RareVar/3_output/plots/1_Endophenotype/1_LM/MultiVariant/allmulti_vs_single_lm_results.tsv') %>%
  mutate(Label_short = str_match(Label_short, '(^[^()]+)')[,2]) %>%
  select(-Label) %>%
  dplyr::rename(Label = Label_short, predictor_label = comparison) %>%
  mutate(adj_p = p.adjust(pval, method='BH'))  %>%
  mutate(predictor_reported = predictor_label)

summary_forest_plotter(multi_vs_single_lm_results, 'genetic_subset_analysis/RareVar/3_output/plots/1_Endophenotype/1_LM/MultiVariant/', file_name = 'LinearRegression_multivssingle_summary_forestplot_fdr.png', sigphenos_only=F,facet_order=c(c('Multi vs. Single')), height = 6, width=4)
```

```{r}
#This repeats with the non-transformed (includng no LGE logged)
multi_vs_single_lm_results_notransform <- map(continuous_param_list, ~linear_logistic_regression_modeller(multi_vs_single_tb_notransform, .,'Multi','Single', covars=c('age','sex','bmi','FieldStrength'))) %>% bind_rows()  %>%
  mutate(comparison = str_c(group1, ' vs. ', group2)) %>%
  left_join(select(label_tb,Variable, Label),by=c('pheno'='Variable')) %>%
  mutate(estimate = - estimate) %>% #Invert the estimate to reflect group1 vs. group2 
  select(-term) %>%
  mutate(Label_short = case_when(
    pheno == 'wallthkmax' ~ 'maxLVWT',
    pheno == 'lvmassi' ~ 'LVMi',
    pheno == 'lge_total' ~ 'LGE',
    pheno == 'ecvfwhole' ~ 'ECVF',
    pheno == 'lvef' ~ 'LVEF',
    pheno == 'lvs_ellmean' ~ '-LongStrain',
    pheno == 'lvs_eccmean' ~ '-CircStrain',
    pheno == 'lvs_errmean' ~ 'RadStrain'
  )) %>%
    mutate(Label_short = str_c(Label_short, ' (', comparison,')'))

write_tsv(multi_vs_single_lm_results_notransform,'genetic_subset_analysis/RareVar/3_output/plots/1_Endophenotype/1_LM/MultiVariant/allmulti_vs_single_lm_notransform_results.tsv')
```

```{r}
#Wilcoxon rank sum test on nontransformed tibble results for those which are MTC-significant

multi_vs_single_wilcoxon_results <- map(
  continuous_param_list[names(continuous_param_list) %in% c('lge_total')],
  ~plot_tester(input_tb=multi_vs_single_tb_notransform,.,group1_name='Multi', group2_name='Single', output_path='genetic_subset_analysis/RareVar/3_output/plots/1_Endophenotype/2_Wilcoxon/MultiVariant/')
  ) %>% 
    bind_rows()
```

These 26 individuals can be sub-categorised into

<!-- 1) 1x homozygote for MYBPC3 P (frameshift) + heterozygote for VUS  -->
<!-- 2) 1x homozygotes for (MYBPC3) P/LP  with MYBPC3 stop-gain -->
Above 2 are defunct and false positive.

3) 2x homozygotes for MYL3 VUS

4) 3x double heterozygotes for 2x different P/LP variants (1x LP MYH7 + LP TNNI3; 1X)
5) 19x double heterozygotes for 1x P/LP variant and 1x VUS variant
6) 2x double heterozygotes for 2x VUS variants

```{r multi_dosage_subcategorisation}

cat1 <- multi_vs_single_tb_notransform %>%
  filter(HCR_IDs %in% filter(homozygotes, Curation1 %in% c('P','LP'), dosage1==2, dosage2==1)$HCR_ID)

cat2 <- multi_vs_single_tb_notransform %>%
  filter(HCR_IDs %in% filter(homozygotes, Curation1 %in% c('P','LP'), dosage1==2, is.na(dosage2))$HCR_ID
  )
cat3 <- multi_vs_single_tb_notransform %>%
  filter(HCR_IDs %in% filter(homozygotes, Curation1 %in% c('VUS','VUS_FP'), dosage1==2, is.na(dosage2))$HCR_ID)

cat4 <- multi_vs_single_tb_notransform %>%
  filter(HCR_IDs %in% filter(all_merged_overlapvars_tb,Curation1 %in% c('P','LP'),Curation2 %in% c('P','LP'),dosage1==1, dosage2==1)$HCR_ID)

cat5 <- multi_vs_single_tb_notransform %>%
  filter(
    HCR_IDs %in% filter(all_merged_overlapvars_tb,Curation1 %in% c('P','LP'),Curation2 %in% c('VUS','VUS_FP'),dosage1==1, dosage2==1)$HCR_ID |
    HCR_IDs %in% filter(all_merged_overlapvars_tb,Curation2 %in% c('P','LP'),Curation1 %in% c('VUS','VUS_FP'),dosage1==1, dosage2==1)$HCR_ID)

cat6 <- multi_vs_single_tb_notransform %>%
  filter(HCR_IDs %in% filter(all_merged_overlapvars_tb,Curation1 %in% c('VUS','VUS_FP'),Curation2 %in% c('VUS','VUS_FP'),dosage1==1, dosage2==1)$HCR_ID)

print(str_glue('Category 1 (Homozygous P/LP + Heterozygous): {nrow(cat1)} individuals\n Category 2 (Homozygous P/LP): {nrow(cat2)} individuals\n Category 3 (Homozygous VUS): {nrow(cat3)} individuals\n Category 4 (Double Heterozygous P/LP): {nrow(cat4)} individuals\n Category 5 (Heterozygous P/LP + Heterozygous VUS): {nrow(cat5)} individuals\n Category 6 (Double Heterozygous VUS): {nrow(cat6)} individuals'))

all_cats <- c(cat1$HCR_IDs, cat2$HCR_IDs, cat3$HCR_IDs, cat4$HCR_IDs, cat5$HCR_IDs, cat6$HCR_IDs)

#Which individuals are not categorised?
uncategorised_ids <- setdiff(multi_dosage_ids, all_cats)
print(uncategorised_ids) #Should be none

rm(homozygotes, multi_het, multi_vs_single_lm_results, multi_vs_single_wilcoxon_results, multi_dosage_ids, multi_vs_single_tb, multi_vs_single_tb_notransform)

rm(list=ls(pattern='cat'))

```
